### 2016-09-29 MSF Groupwork - NBA Hot Hand.Rmd
Created blank file

---
title: "MSF Groupwork"
author: "Justin"
date: "29 September 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2016-09-29 dataframe.R
m

@@ -24,5 +24,5 @@ PlayerName <- df[,20] #name of the player making the shot
 PlayerID <- df[,21] #id of the player making the shot
 
 summary(df)
-
+  
 

### 2016-09-30 MSFGroupwork-NBAHotHand.Rmd
resolving file name issue
test 2
Added report structure

--- 2016-09-29 MSF Groupwork - NBA Hot Hand.Rmd
@@
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+Report structure
+Introduction
+I   What is your theme or research question?
+Theory
+I   What do you expect to find?
+Methods
+I   Describe the dataset:
+I   Strengths
+I   Limitations
+Analysis
+Discussion
+I   What results did you find?
+I   Why is this interesting?
+I   What would be the next steps?

### 2016-09-30 MSFGroupwork-NBAHotHand.Rmd
updated formt+ info on dataset

@@
 ---
-title: "MSF Groupwork"
-author: "Justin"
-date: "29 September 2016"
+title: "MSF_Groupwork-NBA_shotlog"
+author: "Team_9"
 output: html_document
@@
 ```
+Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Widodo
 
-## R Markdown
+#Report structure
 
-Report structure
-Introduction
+###Introduction
 I   What is your theme or research question?
-Theory
+
+###Theory
 I   What do you expect to find?
-Methods
+
+###Methods
 I   Describe the dataset:
+
+Data on shots taken during the 2014-2015 season, who took the shot, where on the floor was the shot taken from, who was the nearest defender, how far away was the nearest defender, time on the shot clock, and much more.
+The column titles are generally self-explanatory.
+Useful for evaluating who the best shooter is, who the best defender is, the hot-hand hypothesis, etc.
+Scraped from NBA's REST API.  ##TO BE UPDATED FURTHER
+
 I   Strengths
+
 I   Limitations
-Analysis
-Discussion
+
+###Analysis
+
+###Discussion
 I   What results did you find?
+
 I   Why is this interesting?
+
 I   What would be the next steps?

### 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
Add topic options based on today's discussions.
started cleaning up data (win/loss column)

@@
 ---
+title: "MSF_Groupwork-NBA_shotlog"
+author: "Team_9"
+output: html_document  ---
   title: "MSF_Groupwork-NBA_shotlog"
@@
 I   What is your theme or research question?
+a. Hot Hand Theory  
+b. Defender Proximity Analysis  
+c. Shot Clock Pressure  
+d. Close Game Effect - Small Margin affecting the scoring rate  
+e. Home Advantage (Alex is working in this point)  
+f. Fatigue Effect (Quarter Heat Rate)  
+g. Shooting Percentage  
+h. Fatigue Effect (Number of Rest Day)  
+i.
 
@@
 I   What do you expect to find?
+a.
+
+###Methods
+I   Describe the dataset:
+
+  Data on shots taken during the 2014-2015 season, who took the shot, where on the floor was the shot taken from, who was the nearest defender, how far away was the nearest defender, time on the shot clock, and much more.
+The column titles are generally self-explanatory.
+Useful for evaluating who the best shooter is, who the best defender is, the hot-hand hypothesis, etc.
+Scraped from NBA's REST API.  ##TO BE UPDATED FURTHER
+
+I   Strengths
+
+I   Limitations
+
+###Analysis
+
+###Data clean-up
+
+```{r}
+df<- read.csv("shot_logs.csv")  #assigning the dataframe to 'df' variables
+attach(df) #telling R that this is the dataframe we'll be working with (eliminating the need to use df$ for variables)
+
+```
+
+```{r}
+
+df[,W] <- as.character(df[,W])
+
+for (i in 1:nrow(df)) {
+  df[i,] <- as.character(ky[, i])
+  if (df[i,"W"]="W") df[i,"W"]<- 1
+  if (df[i,"W"]="L") df[i,"W"]<- 0
+}
+
+for (i in 1:(ncol(ky) - 1)){
+  ky[, i] <- as.character(ky[, i])
+  for (j in 1:nrow(ky)){
+    if (ky[j, i] == "Strongly Agree") ky[j, i] <- 5
+    if (ky[j, i] == "Agree") ky[j, i] <- 4
+    if (ky[j, i] == "Neither Agree nor Disagree") ky[j, i] <- 3
+    if (ky[j, i] == "Disagree") ky[j, i] <- 2
+    if (ky[j, i] == "Strongly Disagree") ky[j, i] <- 1
+  }
+  ky[, i] <- as.numeric(ky[, i])
+}
+
+```
+
+###Discussion
+I   What results did you find?
+
+I   Why is this interesting?
+
+I   What would be the next steps?
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
+Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Widodo
+
+#Report structure
+
+###Introduction
+I   What is your theme or research question?
+a. Hot Hand Theory
+b. Defender Proximity Analysis
+c. Shot Clock Pressure
+d. Close Game Effect - Small Margin affecting the scoring rate
+e. Home Advantage
+f. Fatigue Effect (Quarter Heat Rate)
+g. Shooting Percentage
+h. Fatigue Effect (Number of Rest Day)
+i.
+
+###Theory
+I   What do you expect to find?
+a.
 

### 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
working on w/L criteria

@@
 
-for (i in 1:(ncol(ky) - 1)){
-  ky[, i] <- as.character(ky[, i])
-  for (j in 1:nrow(ky)){
-    if (ky[j, i] == "Strongly Agree") ky[j, i] <- 5
-    if (ky[j, i] == "Agree") ky[j, i] <- 4
-    if (ky[j, i] == "Neither Agree nor Disagree") ky[j, i] <- 3
-    if (ky[j, i] == "Disagree") ky[j, i] <- 2
-    if (ky[j, i] == "Strongly Disagree") ky[j, i] <- 1
-  }
-  ky[, i] <- as.numeric(ky[, i])
-}
+#for (i in 1:(ncol(ky) - 1)){
+#  ky[, i] <- as.character(ky[, i])
+ # for (j in 1:nrow(ky)){
+  #  if (ky[j, i] == "Strongly Agree") ky[j, i] <- 5
+   # if (ky[j, i] == "Agree") ky[j, i] <- 4
+    #if (ky[j, i] == "Neither Agree nor Disagree") ky[j, i] <- 3
+#    if (ky[j, i] == "Disagree") ky[j, i] <- 2
+#    if (ky[j, i] == "Strongly Disagree") ky[j, i] <- 1
+#  }
+#  ky[, i] <- as.numeric(ky[, i])
+#}
 

### 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
replaced w/l with 1/0

--- 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
@@
 ```
-
+### Home Advantage analysis
 ```{r}
-
-df[,W] <- as.character(df[,W])
+# replacing W/L in column "W" with 1/0 (1 for win, 0 for loss)
+df[,"W"] <- as.character(df[,"W"])
 
 for (i in 1:nrow(df)) {
-  df[i,] <- as.character(ky[, i])
-  if (df[i,"W"]="W") df[i,"W"]<- 1
-  if (df[i,"W"]="L") df[i,"W"]<- 0
+  if (df[i,"W"]=="W") df[i,"W"]<- 1
+  if (df[i,"W"]=="L") df[i,"W"]<- 0
 }
 
-for (i in 1:(ncol(ky) - 1)){
-  ky[, i] <- as.character(ky[, i])
-  for (j in 1:nrow(ky)){
-    if (ky[j, i] == "Strongly Agree") ky[j, i] <- 5
-    if (ky[j, i] == "Agree") ky[j, i] <- 4
-    if (ky[j, i] == "Neither Agree nor Disagree") ky[j, i] <- 3
-    if (ky[j, i] == "Disagree") ky[j, i] <- 2
-    if (ky[j, i] == "Strongly Disagree") ky[j, i] <- 1
-  }
-  ky[, i] <- as.numeric(ky[, i])
-}
+df[,"W"] <- as.numeric(df[,"W"])
 

### 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
Added columns for names of HOME and AWAY teams

@@
 
+##Installing packages
+```{r}
+install.packages('reshape')
+
+```
+
 ###Data clean-up
@@
 df[,"W"] <- as.numeric(df[,"W"])
+```
 
+```{r}
+#adding columns for names of 'home' and 'away' teams
+df[,"MATCHUP"] <- as.character(df[,"MATCHUP"])
+
+for (i in 1:nrow(df)) {
+  df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
+  df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7]
+}
+```
+
+```{r}
+#combining columns for the analysis
+library(reshape)
 ```

### 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
creating a new DF for win/loss analysis

@@
 for (i in 1:nrow(df)) {
-  df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
-  df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7]
+  if (LOCATION[i]=="H") {df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
+  df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7]}
+  if (LOCATION[i]=="A") {df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7] 
+  df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]}
 }
+
+df$BLANK<-"" #for some reason cast function does not work with last column in dataframe, so I have created this BLANK column as I need to use AWAY_TEAM colum (it was a last column). 
 ```
@@
 library(reshape)
+dfwl<- cast(df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + W~ LOCATION)
+
+summary(dfwl)
 ```

### 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
completed home/away initial analysis

--- 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
@@
 
+#we need to change the format of some of the columns to 'Character'
+
 ```
@@
   if (df[i,"W"]=="W") df[i,"W"]<- 1
-  if (df[i,"W"]=="L") df[i,"W"]<- 0
-}
+  if (df[i,"W"]=="L") df[i,"W"]<- 0}
 
@@
 for (i in 1:nrow(df)) {
-  df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
+  if (LOCATION[i]=="H") {df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
   df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7]
-}
+  if (W[i]==1) df$WINNER[i] <- "HOME" 
+  if (W[i]==0) df$WINNER[i] <- "AWAY"}
+  if (LOCATION[i]=="A") {df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7] 
+  df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
+  if (W[i]==0) df$WINNER[i] <- "HOME"  
+  if (W[i]==1) df$WINNER[i] <- "AWAY"}}
+
+df$BLANK<-"" #for some reason cast function does not work with last column in dataframe, so I have created this BLANK column as I need to use AWAY_TEAM colum (it was a last column). 
 ```
@@
 library(reshape)
+dfwl<- cast(df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + WINNER ~ LOCATION)
+dfwl[,"WINNER"] <- as.character(dfwl[,"WINNER"])
+
+for (i in 1:nrow(dfwl)) { 
+if (dfwl$WINNER[i]=="HOME") dfwl$WINNER2[i]<-1
+if (dfwl$WINNER[i]=="AWAY") dfwl$WINNER2[i]<-2}
+```
+
+```{r}
+dfwl2<-cast(dfwl, HOME_TEAM ~ WINNER)
+dfwl2<- dfwl2[,c(1,3,2)]
+dfwl2$WIN_PCT<- round(dfwl2$HOME/(dfwl2$HOME+dfwl2$AWAY),2)
+
+for (i in 1:nrow(dfwl2)) {
+if(dfwl2$WIN_PCT[i]>0.5) dfwl2$HOMEADV[i]<- "YES" else dfwl2$HOMEADV[i]<- "NO"}
+
+dfwl2<- dfwl2[order(dfwl2$WIN_PCT),]; dfwl2
+
+library(ggplot2)
+ggplot(dfwl2, aes(x=HOME_TEAM,y=WIN_PCT,col=HOMEADV))+ geom_point()
 ```

### 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
added some codes for defender proximity analysis (correlate between defender distance with points made)

--- 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
@@
 a. Hot Hand Theory  
-b. Defender Proximity Analysis  
+b. Defender Proximity Analysis (Reza)
 c. Shot Clock Pressure  
@@
 
+###Defender Proximity Analysis
+```{r}
+#Data Preparation
+#Change the value of shot result column to numeric
+df[, "SHOT_RESULT"] <- as.character(df[, "SHOT_RESULT"])
+for (j in 1:nrow(df)){
+  df[j, "SHOT_RESULT"] <- ifelse (df[j, "SHOT_RESULT"] == "made", 1, 0)
+}
+
+#set the value to numeric
+df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
+df[, "PTS_TYPE"] <- as.integer(df[, "PTS_TYPE"])
+df[, "CLOSE_DEF_DIS"] <- as.numeric(df[, "CLOSE_DEF_DIST"])
+
+#calculate correlation between defender distance and shot result
+cor(df[, "SHOT_RESULT"], df[, "CLOSE_DEF_DIS"])
+
+#calculate correlation between defender distance and points made (2 or 3 points)
+cor(df[, "PTS_TYPE"], df[, "CLOSE_DEF_DIS"])
+
+#plot the diagram
+library(ggplot2)
+ggplot(data = df, aes(x = PTS_TYPE, y = CLOSE_DEF_DIST, group = PTS_TYPE)) + geom_boxplot() + stat_summary(fun.y = mean, geom = "point", shape = 5, size = 1) +ggtitle("Points Scored based on Distance from Defender")
+```
+
 ###Discussion

### 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
add notes for point Shot Clock Presure by Alex
I just started on this topic
I'll analyse this topic
win lose analysis

--- 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
@@
 I   What is your theme or research question?
-a. Hot Hand Theory  
+a. Hot Hand Theory (Artemis)
 b. Defender Proximity Analysis (Reza)
-c. Shot Clock Pressure  
+c. Shot Clock Pressure  (Alex)
 d. Close Game Effect - Small Margin affecting the scoring rate  
@@
 f. Fatigue Effect (Quarter Heat Rate)  
-g. Shooting Percentage  
+g. Shooting Percentage (James)
 h. Fatigue Effect (Number of Rest Day)  
@@
 
-##Installing packages
+##Installing packages and calling libraries
 ```{r}
 install.packages('reshape')
+library(reshape)
+library(ggplot2)
 
@@
 ```{r}
-#adding columns for names of 'home' and 'away' teams
+#adding columns for names of 'home' and 'away' teams. In addition, WINNER column tells which team won
 df[,"MATCHUP"] <- as.character(df[,"MATCHUP"])
+df[,"W"] <- as.character(df[,"W"])
 
 for (i in 1:nrow(df)) {
-  df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
+  if (LOCATION[i]=="H") {df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
   df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7]
-}
+  if (W[i]=="W") df$WINNER[i] <- "HOME" 
+  if (W[i]=="L") df$WINNER[i] <- "AWAY"}
+  if (LOCATION[i]=="A") {df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7] 
+  df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
+  if (W[i]=="L") df$WINNER[i] <- "HOME"  
+  if (W[i]=="W") df$WINNER[i] <- "AWAY"}}
+
+df$BLANK<-"" #for some reason cast function does not work with last column in dataframe, so I have created this BLANK column as I need to use AWAY_TEAM colum (it was a last column). 
 ```
@@
 ```{r}
-#combining columns for the analysis
-library(reshape)
+aa<- cast(df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + WINNER ~ W)
+
+table(aa$WINNER) [2]
+#home wins:
+hh<-sum(aa$WINNER=="HOME")
+#Away wins:
+aw<-sum(aa$WINNER=="AWAY")
+#percentage of home team winning: 
+round(hh/(hh+aw),2)
+```
+
+```{r}
+head(aa)
+aa1<- cast(aa, HOME_TEAM  ~ WINNER); names(aa1)[1]<-"TEAM"
+aa1$WIN_PCT_HOME<-round(aa1$HOME/(aa1$HOME+aa1$AWAY),2)
+
+aa2<- cast(aa, AWAY_TEAM ~ WINNER); names(aa2)[1]<-"TEAM"
+aa2$WIN_PCT_AWAY<-round(aa2$HOME/(aa2$HOME+aa2$AWAY),2)
+
+aa3<- merge (aa1,aa2, by ="TEAM")
+for (i in 1:nrow(aa3)) {if(aa3$WIN_PCT_HOME[i]>aa3$WIN_PCT_AWAY[i]) aa3$HOME_ADV[i]<-"YES" else aa3$HOME_ADV[i]<-"NO"}
+head(aa3)
+
+ggplot(aa3,aes(x=TEAM, y= WIN_PCT_HOME, col= HOME_ADV))+ geom_point()
+#TO BE UPDATED
 ```

### 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
calculate top 5 players with most 2 points and 3 points

--- 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
@@
 b. Defender Proximity Analysis (Reza)
-c. Shot Clock Pressure  
+c. Shot Clock Pressure  (Alex)
 d. Close Game Effect - Small Margin affecting the scoring rate  
@@
 
-#set the value to numeric
+#set the value to numeric or integer accordingly
 df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
@@
 
+#select data with only scored result
+#Point type (2 or 3 points) and distance
+point_type <- c()
+distance <- c()
+for (z in 1:nrow(df)){
+  if(df[z, "SHOT_RESULT"] == 1 & df[z, "PTS_TYPE"] >= 1){
+    point_type <- c(point_type, df[z, "PTS_TYPE"])
+    distance <- c(distance, df[z, "CLOSE_DEF_DIST"])
+  }
+}
+
+#combine point type and distance
+valid_distance = data.frame(point_type, distance)
+
+#calculate correlation between defender distance and points made (2 or 3 points)
+cor(point_type, distance)
+
+#plot the diagram for this correlation
+library(ggplot2)
+ggplot(data = valid_distance, aes(x = point_type, y = distance, group = point_type)) + 
+  geom_boxplot() + stat_summary(fun.y = mean, geom = "point", shape = 5, size = 1) + 
+  ggtitle("Points Scored based on Distance from Defender")
+
 #calculate correlation between defender distance and shot result
@@
 
-#calculate correlation between defender distance and points made (2 or 3 points)
-cor(df[, "PTS_TYPE"], df[, "CLOSE_DEF_DIS"])
+#--------------------------------------------------------------------------------#
 
-#plot the diagram
+#further analysis
+#select distance data from 2 points shot scored and the players who scored it
+distance2 <- c()
+players2 <- c()
+for (z in 1:nrow(df)){
+  if(df[z, "SHOT_RESULT"] == 1 & df[z, "PTS_TYPE"] == 2){
+    distance2 <- c(distance2, df[z, "CLOSE_DEF_DIST"])
+    players2 <- c(players2, as.character(df[z, "player_name"]))
+  }
+}
+
+#combine 2 point players and defender distance
+player2point <- data.frame(players2, distance2)
+summary(player2point)
+
+#count players who scored 2 points
+counter <- as.data.frame(table(players2))
+
+#count 5 top players
+top5Players2 <- counter[rev(order(counter$Freq)), "players2"][1:5]
+top5Counter <- counter[rev(order(counter$Freq)), "Freq"][1:5]
+
+top5_2points <- data.frame(top5players2, top5Counter)
+
+#plot the graph
 library(ggplot2)
-ggplot(data = df, aes(x = PTS_TYPE, y = CLOSE_DEF_DIST, group = PTS_TYPE)) + geom_boxplot() + stat_summary(fun.y = mean, geom = "point", shape = 5, size = 1) +ggtitle("Points Scored based on Distance from Defender")
+ggplot(data = top5_2points, aes(x = top5Players2)) + geom_bar(aes(x = top5players2, y = top5Counter), stat = "identity")
+
+#----------------------------------------------------------------#
+
+#select distance data from 3 points shot scored and the players who scored it
+distance3 <- c()
+players3 <- c()
+for (z in 1:nrow(df)){
+  if(df[z, "SHOT_RESULT"] == 1 & df[z, "PTS_TYPE"] == 3){
+    distance3 <- c(distance3, df[z, "CLOSE_DEF_DIST"])
+    players3 <- c(players3, as.character(df[z, "player_name"]))
+  }
+}
+
+#combine 3 point players and defender distance
+player3point <- data.frame(players3, distance3)
+summary(player3point)
+
+#count players who scored 3 points
+counter3 <- as.data.frame(table(players3))
+
+#count 5 top players
+top5Players3 <- counter3[rev(order(counter3$Freq)), "players3"][1:5]
+top5Counter3 <- counter3[rev(order(counter3$Freq)), "Freq"][1:5]
+
+top5_3points <- data.frame(top5Players3, top5Counter3)
+
+#plot the graph
+library(ggplot2)
+ggplot(data = top5_3points, aes(x = top5Players3)) + geom_bar(aes(x = top5Players3, y = top5Counter3), stat = "identity")
+
 ```

### 2016-10-09 MSFGroupwork-NBAHotHand.Rmd

@@
 I   What is your theme or research question?
-a. Hot Hand Theory  
+a. Hot Hand Theory (Artemis)
 b. Defender Proximity Analysis (Reza)
@@
 ```{r}
-#adding columns for names of 'home' and 'away' teams
+#adding columns for names of 'home' and 'away' teams. In addition, WINNER column tells which team won
 df[,"MATCHUP"] <- as.character(df[,"MATCHUP"])
+df[,"W"] <- as.character(df[,"W"])
 
@@
   df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7]
-  if (W[i]==1) df$WINNER[i] <- "HOME" 
-  if (W[i]==0) df$WINNER[i] <- "AWAY"}
+  if (W[i]=="W") df$WINNER[i] <- "HOME" 
+  if (W[i]=="L") df$WINNER[i] <- "AWAY"}
   if (LOCATION[i]=="A") {df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7] 
   df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
-  if (W[i]==0) df$WINNER[i] <- "HOME"  
-  if (W[i]==1) df$WINNER[i] <- "AWAY"}}
+  if (W[i]=="L") df$WINNER[i] <- "HOME"  
+  if (W[i]=="W") df$WINNER[i] <- "AWAY"}}
 
@@
 ```{r}
-#combining columns for the analysis
-library(reshape)
-dfwl<- cast(df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + WINNER ~ LOCATION)
-dfwl[,"WINNER"] <- as.character(dfwl[,"WINNER"])
+aa<- cast(df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + WINNER ~ W)
 
-for (i in 1:nrow(dfwl)) { 
-if (dfwl$WINNER[i]=="HOME") dfwl$WINNER2[i]<-1
-if (dfwl$WINNER[i]=="AWAY") dfwl$WINNER2[i]<-2}
+table(aa$WINNER) [2]
+#home wins:
+hh<-sum(aa$WINNER=="HOME")
+#Away wins:
+aw<-sum(aa$WINNER=="AWAY")
+#percentage of home team winning: 
+round(hh/(hh+aw),2)
 ```
@@
 ```{r}
-dfwl2<-cast(dfwl, HOME_TEAM ~ WINNER)
-dfwl2<- dfwl2[,c(1,3,2)]
-dfwl2$WIN_PCT<- round(dfwl2$HOME/(dfwl2$HOME+dfwl2$AWAY),2)
+head(aa)
+aa1<- cast(aa, HOME_TEAM  ~ WINNER); names(aa1)[1]<-"TEAM"
+aa1$WIN_PCT_HOME<-round(aa1$HOME/(aa1$HOME+aa1$AWAY),2)
 
-for (i in 1:nrow(dfwl2)) {
-if(dfwl2$WIN_PCT[i]>0.5) dfwl2$HOMEADV[i]<- "YES" else dfwl2$HOMEADV[i]<- "NO"}
+aa2<- cast(aa, AWAY_TEAM ~ WINNER); names(aa2)[1]<-"TEAM"
+aa2$WIN_PCT_AWAY<-round(aa2$HOME/(aa2$HOME+aa2$AWAY),2)
 
-dfwl2<- dfwl2[order(dfwl2$WIN_PCT),]; dfwl2
+aa3<- merge (aa1,aa2, by ="TEAM")
+for (i in 1:nrow(aa3)) {if(aa3$WIN_PCT_HOME[i]>aa3$WIN_PCT_AWAY[i]) aa3$HOME_ADV[i]<-"YES" else aa3$HOME_ADV[i]<-"NO"}
+head(aa3)
 
-library(ggplot2)
-ggplot(dfwl2, aes(x=HOME_TEAM,y=WIN_PCT,col=HOMEADV))+ geom_point()
+ggplot(aa3,aes(x=TEAM, y= WIN_PCT_HOME, col= HOME_ADV))+ geom_point()
+#TO BE UPDATED
 ```

### 2016-10-09 MSFGroupwork-NBAHotHand.Rmd
little code styling revision

@@
 
-#further analysis
+#Further analysis
 #select distance data from 2 points shot scored and the players who scored it
@@
 library(ggplot2)
-ggplot(data = top5_2points, aes(x = top5Players2)) + geom_bar(aes(x = top5players2, y = top5Counter), stat = "identity")
+ggplot(data = top5_2points, aes(x = top5Players2)) +
+  geom_bar(aes(x = top5players2, y = top5Counter), stat = "identity")
 
@@
 library(ggplot2)
-ggplot(data = top5_3points, aes(x = top5Players3)) + geom_bar(aes(x = top5Players3, y = top5Counter3), stat = "identity")
+ggplot(data = top5_3points, aes(x = top5Players3)) +
+  geom_bar(aes(x = top5Players3, y = top5Counter3), stat = "identity")
 

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
1st part - Hot Hand Theory

@@
 ```
+###Hot Hand Theory
+
+```{r}
+#Set the value to character
+df[, "SHOT_RESULT"] <- as.character(df[, "SHOT_RESULT"])
+
+#Define 1 for baskets made and 0 for baskets missed. We need a vector to use it to our function below.
+shotresult <- c()
+for(i in 1:nrow(df)) {
+  if(shots[i] == 'missed') shotresult <- c(shotresult, 0)
+  else if(shots[i] == 'made') shotresult <- c(shotresult, 1)
+}
+
+shotresult
+
+#Set the value to numeric/integer
+df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
+df[ ,'SHOT_NUMBER'] <- as.integer(df[ , 'SHOT_NUMBER'])
+
+#Create new column with shot results(1 for baskets made and 0 for baskets missed)
+df[ ,'SHOTRESULT_0_1'] <- shotresult
+
+#Define the length of a shooting streak to be the number of consecutive baskets made until a miss occurs
+#Create a function that gives us a vector with the number of streaks
+number_of_streaks <- function(x) {
+  x <- c(0, x, 0)
+  which_zero <- which(x == 0)
+  streak <- diff(which_zero) - 1
+  streak
+}
+
+number_of_streaks(shotresult)
+#To be continued (probabilities- indipendence, plots, interpret results for specific players(those with higher shot_numbers))
+```
 

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
summary of points code updated

@@
 ```
+
+#SUMMARY OF POINTS
+```{r}
+ptssum<- cast(df, GAME_ID1+ player_id + player_name ~ PTS); head(ptssum)
+#creating "total points per player, per game" column
+ptssum$totalpts<- ptssum[,"2"]*2 + ptssum[,"3"]*3
+
+pttsum2<-ptssum[order(ptssum$totalpts,decreasing=TRUE),]
+head(pttsum2,10)
+
+```
+
 ### Home Advantage analysis

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
struggling with the graph(home win loss)

@@
 df$BLANK<-"" #for some reason cast function does not work with last column in dataframe, so I have created this BLANK column as I need to use AWAY_TEAM colum (it was a last column).
+
+write.csv(df,"shot_logs1010.csv")
 ```
@@
 
-table(aa$WINNER) [2]
+table(aa$WINNER)
 #home wins:
@@
 aa2<- cast(aa, AWAY_TEAM ~ WINNER); names(aa2)[1]<-"TEAM"
-aa2$WIN_PCT_AWAY<-round(aa2$HOME/(aa2$HOME+aa2$AWAY),2)
+aa2$WIN_PCT_AWAY<-round(aa2$AWAY/(aa2$HOME+aa2$AWAY),2)
 
@@
 
-ggplot(aa3,aes(x=TEAM, y= WIN_PCT_HOME, col= HOME_ADV))+ geom_point()
+#aa3$WIN_PCT_HOME <- with(aa3, reorder(WIN_PCT_HOME, WIN_PCT_HOME, function(aa3) - length(aa3)))
+#ggplot(aa3) + geom_point(aes(WIN_PCT_HOME))
+
+aa3<-aa3[order(aa3$WIN_PCT_HOME,decreasing=TRUE),]
+head(aa3)
+
+#home_pct <- table(aa3$WIN_PCT_HOME)
+#home_pct_level <- names(home_pct)[order(home_pct)]
+#aa3$WIN_PCT_HOME2 <- factor(aa3$WIN_PCT_HOME, levels = home_pct_level)
+head(aa3)
+
+ggplot(aa3,aes(x=TEAM,y= WIN_PCT_HOME,colour = "HOME")) +geom_point() + geom_point(aes(y= WIN_PCT_AWAY, colour ="AWAY"))
+
+#ggplot(aa3,aes(x=TEAM))+ geom_point(aes(y= WIN_PCT_HOME,colour = "HOME")) + geom_point(aes(y= WIN_PCT_AWAY, colour ="AWAY"))
 #TO BE UPDATED

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
changes
part 2 - Hot Hand Theory

--- 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
@@
 attach(df) #telling R that this is the dataframe we'll be working with (eliminating the need to use df$ for variables)
-
+df1 <- read.csv('shot_logs1010.csv')
 #we need to change the format of some of the columns to 'Character'
@@
 ```{r}
-ptssum<- cast(df, GAME_ID1+ player_id + player_name ~ PTS); head(ptssum)
+ptssum<- cast(df1, GAME_ID1+ player_id + player_name  ~ PTS); head(ptssum)
 #creating "total points per player, per game" column
@@
 number_of_streaks(shotresult)
+
+player_and_game <- function(x, y) {
+   player <- df[df[ ,'player_id'] == x & df[ ,'GAME_ID1'] == y, 'FGM']
+   number_of_streaks(as.vector(player))
+}
+player_and_game(202681, 21400681)
+
+ourtable <- head(pttsum2,10)
+
+list_streaks <- list()
+
+for(i in ourtable$player_id) {
+   streak_per_player <- player_and_game(i, ourtable[ourtable[ ,'player_id'] == i, 'GAME_ID1'])
+   print(streak_per_player)
+}
+
 #To be continued (probabilities- indipendence, plots, interpret results for specific players(those with higher shot_numbers))

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
FG% analysis (not done yet)

--- 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
@@
 ```
-
-#SUMMARY OF POINTS
-```{r}
-ptssum<- cast(df, GAME_ID1+ player_id + player_name ~ PTS); head(ptssum)
-#creating "total points per player, per game" column
-ptssum$totalpts<- ptssum[,"2"]*2 + ptssum[,"3"]*3
-
-pttsum2<-ptssum[order(ptssum$totalpts,decreasing=TRUE),]
-head(pttsum2,10)
-
-```
-
 ### Home Advantage analysis
@@
 #TO BE UPDATED
+```
+
+# Shooting Percentage Analysis (James)
+
+```{r}
+# Replace "made" and "missed" with 1 and 0 respectively
+
+df[, "SHOT_RESULT"] <- as.character(df[, "SHOT_RESULT"])
+
+for (i in 1:nrow(df)) {
+  if (df[i, "SHOT_RESULT"] == "made") df[i, "SHOT_RESULT"] <- 1
+  if (df[i, "SHOT_RESULT"] == "missed") df[i, "SHOT_RESULT"] <- 0
+}
+
+df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
+
 ```
@@
 ```
-###Hot Hand Theory
-
-```{r}
-#Set the value to character
-df[, "SHOT_RESULT"] <- as.character(df[, "SHOT_RESULT"])
-
-#Define 1 for baskets made and 0 for baskets missed. We need a vector to use it to our function below.
-shotresult <- c()
-for(i in 1:nrow(df)) {
-  if(shots[i] == 'missed') shotresult <- c(shotresult, 0)
-  else if(shots[i] == 'made') shotresult <- c(shotresult, 1)
-}
-
-shotresult
-
-#Set the value to numeric/integer
-df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
-df[ ,'SHOT_NUMBER'] <- as.integer(df[ , 'SHOT_NUMBER'])
-
-#Create new column with shot results(1 for baskets made and 0 for baskets missed)
-df[ ,'SHOTRESULT_0_1'] <- shotresult
-
-#Define the length of a shooting streak to be the number of consecutive baskets made until a miss occurs
-#Create a function that gives us a vector with the number of streaks
-number_of_streaks <- function(x) {
-  x <- c(0, x, 0)
-  which_zero <- which(x == 0)
-  streak <- diff(which_zero) - 1
-  streak
-}
-
-number_of_streaks(shotresult)
-#To be continued (probabilities- indipendence, plots, interpret results for specific players(those with higher shot_numbers))
-```
 

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
...

@@
 df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
+
+```
+
+```{r}
+# Find shooting percentage of each match
+
+# Trying to reshape...
+
+library(reshape)
+game <- cast(data = df, GAME_ID1 ~ SHOT_RESULT)
+View(game)
+
+#
+
+Team <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
+FG_Percentage <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
+for(i in 21400001:21400908){
+  j = i - 21400001 +1 # 
+  game <- df[, df[, "GAME_ID1"] == i]
+  home <- game[, game["LOCATION"] == "H"]
+  home.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / nrow(home)
+  away <- game[, game["LOCATION"] == "A"]
+  away.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / away(home)
+  FG_Percentage[j] <- home.p
+  FG_Percentage[j + 1] = away.p
+  Team[j] = "H"
+  Team[j + 1] = "A"
+}
 

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd

--- 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
@@
 #TO BE UPDATED
+```
+
+# Shooting Percentage Analysis (James)
+
+```{r}
+# Replace "made" and "missed" with 1 and 0 respectively
+
+df[, "SHOT_RESULT"] <- as.character(df[, "SHOT_RESULT"])
+
+for (i in 1:nrow(df)) {
+  if (df[i, "SHOT_RESULT"] == "made") df[i, "SHOT_RESULT"] <- 1
+  if (df[i, "SHOT_RESULT"] == "missed") df[i, "SHOT_RESULT"] <- 0
+}
+
+df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
+
+```
+
+```{r}
+# Find shooting percentage of each match
+
+# Trying to reshape...
+
+library(reshape)
+game <- cast(data = df, GAME_ID1 ~ SHOT_RESULT)
+View(game)
+
+#
+
+Team <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
+FG_Percentage <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
+for(i in 21400001:21400908){
+  j = i - 21400001 +1 # 
+  game <- df[, df[, "GAME_ID1"] == i]
+  home <- game[, game["LOCATION"] == "H"]
+  home.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / nrow(home)
+  away <- game[, game["LOCATION"] == "A"]
+  away.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / away(home)
+  FG_Percentage[j] <- home.p
+  FG_Percentage[j + 1] = away.p
+  Team[j] = "H"
+  Team[j + 1] = "A"
+}
+
 ```

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
j
graph for win loss updated

@@
 ```{r}
-aa<- cast(df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + WINNER ~ W)
+aa<- cast(df1, GAME_ID1 + HOME_TEAM + AWAY_TEAM + WINNER ~ W)
 
@@
 aa2<- cast(aa, AWAY_TEAM ~ WINNER); names(aa2)[1]<-"TEAM"
-aa2$WIN_PCT_AWAY<-round(aa2$HOME/(aa2$HOME+aa2$AWAY),2)
+aa2$WIN_PCT_AWAY<-round(aa2$AWAY/(aa2$HOME+aa2$AWAY),2)
 
 aa3<- merge (aa1,aa2, by ="TEAM")
+
 for (i in 1:nrow(aa3)) {if(aa3$WIN_PCT_HOME[i]>aa3$WIN_PCT_AWAY[i]) aa3$HOME_ADV[i]<-"YES" else aa3$HOME_ADV[i]<-"NO"}
@@
 
-ggplot(aa3,aes(x=TEAM, y= WIN_PCT_HOME, col= HOME_ADV))+ geom_point()
+aa3<- aa3[order(aa3$WIN_PCT_HOME,decreasing=TRUE),]
+rownames(aa4)<-NULL
+
+aa3$TEAM <- factor(aa5$TEAM, levels = aa5$TEAM[order(-aa5$WIN_PCT_HOME)])
+
+ggplot(aa3,aes(x=TEAM)) + geom_point(aes(y= WIN_PCT_HOME, col= "HOME"))+ geom_point(aes( y= WIN_PCT_AWAY, col= "AWAY")) 
 #TO BE UPDATED
-```
-
-# Shooting Percentage Analysis (James)
-
-```{r}
-# Replace "made" and "missed" with 1 and 0 respectively
-
-df[, "SHOT_RESULT"] <- as.character(df[, "SHOT_RESULT"])
-
-for (i in 1:nrow(df)) {
-  if (df[i, "SHOT_RESULT"] == "made") df[i, "SHOT_RESULT"] <- 1
-  if (df[i, "SHOT_RESULT"] == "missed") df[i, "SHOT_RESULT"] <- 0
-}
-
-df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
-
-```
-
-```{r}
-# Find shooting percentage of each match
-
-# Trying to reshape...
-
-library(reshape)
-game <- cast(data = df, GAME_ID1 ~ SHOT_RESULT)
-View(game)
-
-#
-
-Team <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
-FG_Percentage <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
-for(i in 21400001:21400908){
-  j = i - 21400001 +1 # 
-  game <- df[, df[, "GAME_ID1"] == i]
-  home <- game[, game["LOCATION"] == "H"]
-  home.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / nrow(home)
-  away <- game[, game["LOCATION"] == "A"]
-  away.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / away(home)
-  FG_Percentage[j] <- home.p
-  FG_Percentage[j + 1] = away.p
-  Team[j] = "H"
-  Team[j + 1] = "A"
-}
-
 ```

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
graph for win loss

--- 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
@@
 
-ggplot(aa3,aes(x=TEAM, y= WIN_PCT_HOME, col= HOME_ADV))+ geom_point()
+aa3<- aa3[order(aa3$WIN_PCT_HOME,decreasing=TRUE),]
+rownames(aa4)<-NULL
+
+aa3$TEAM <- factor(aa5$TEAM, levels = aa5$TEAM[order(-aa5$WIN_PCT_HOME)])
+
+ggplot(aa3,aes(x=TEAM)) + geom_point(aes(y= WIN_PCT_HOME, col= "HOME"))+ geom_point(aes( y= WIN_PCT_AWAY, col= "AWAY")) 
 #TO BE UPDATED  
@@
 ```
+
+#
 

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
completed graph for win/loss
new code- hand theory

@@
 
+```{r}
+#Summary of total points
+ptssum<- cast(df1, GAME_ID1+ player_id + player_name  ~ PTS); head(ptssum)
+#creating "total points per player, per game" column
+ptssum$totalpts<- ptssum[,"2"]*2 + ptssum[,"3"]*3
+
+pttsum2<-ptssum[order(ptssum$totalpts,decreasing=TRUE),]
+head(pttsum2,10)
+```
+
+```{r}
+###Hot Hand Theory
+
+#Set the value to character
+df[, "SHOT_RESULT"] <- as.character(df[, "SHOT_RESULT"])
+
+#Define 1 for baskets made and 0 for baskets missed. We need a vector to use it to our function below.
+shotresult <- c()
+for(i in 1:nrow(df)) {
+  if(shots[i] == 'missed') shotresult <- c(shotresult, 0)
+  else if(shots[i] == 'made') shotresult <- c(shotresult, 1)
+}
+
+shotresult
+
+#Set the value to numeric/integer
+df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
+df[ ,'SHOT_NUMBER'] <- as.integer(df[ , 'SHOT_NUMBER'])
+
+#Define the length of a shooting streak to be the number of consecutive baskets made until a miss occurs
+#Create a function that gives us a vector with the number of streaks
+number_of_streaks <- function(x) {
+  x <- c(0, x, 0)
+  which_zero <- which(x == 0)
+  streak <- diff(which_zero) - 1
+  streak
+}
+
+number_of_streaks(shotresult)
+
+player_and_game <- function(x, y) {
+   player <- df[df[ ,'player_id'] == x & df[ ,'GAME_ID1'] == y, 'FGM']
+   number_of_streaks(as.vector(player))
+}
+player_and_game(202681, 21400681)
+
+ourtable <- head(pttsum2,10)
+
+list_per_player <- list()
+j <- 1
+for(i in ourtable$player_id) {
+   streak_per_player[[j]] <- player_and_game(i, ourtable[ourtable[ ,'player_id'] == i, 'GAME_ID1'])
+   j <- j + 1
+}
+streak_per_player
+
 ###Discussion

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
Installing and calling is seperated

@@
 
-##Installing packages and calling libraries
+## Installing libraries
+
 ```{r}
 install.packages('reshape')
+
+```
+
+## Calling libraries
+```{r}
 library(reshape)
@@
 ```
+
 ### Home Advantage analysis
+
 ```{r}

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
new tries - hot hand theory

--- 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
@@
 
+#Create a function that gives us a vector of streaks for a specific player and game
 player_and_game <- function(x, y) {
@@
 
-list_per_player <- list()
+#Create a list that gives us the streaks of each player and game (10 best performances)
+streak_per_player <- list()
 j <- 1
@@
 streak_per_player
+
+#Create a data frame for the strikes of each one of the 10 players (cannot create a data.frame including all of them, because each list has different length)
+player1_df <- data.frame('streaks' = streak_per_player[[1]])
+player2_df <- data.frame('streaks' = streak_per_player[[2]])
+player3_df <- data.frame('streaks' = streak_per_player[[3]])
+player4_df <- data.frame('streaks' = streak_per_player[[4]])
+player5_df <- data.frame('streaks' = streak_per_player[[5]])
+player6_df <- data.frame('streaks' = streak_per_player[[6]])
+player7_df <- data.frame('streaks' = streak_per_player[[7]])
+player8_df <- data.frame('streaks' = streak_per_player[[8]])
+player9_df <- data.frame('streaks' = streak_per_player[[9]])
+player10_df <- data.frame('streaks' = streak_per_player[[10]])
+ourtable
+
+#Visualisation for 10 players
+ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = 'Kyrie Irving')
+ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = 'Kyrie Irving')
+ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = 'Kyrie Irving')
+ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = 'Kyrie Irving')
+ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = 'Kyrie Irving')
+ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = 'Kyrie Irving')
+ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = 'Kyrie Irving')
 

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
calculate distance

--- 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
@@
 distance <- c()
+shot_dist <- c()
 for (z in 1:nrow(df)){
@@
     distance <- c(distance, df[z, "CLOSE_DEF_DIST"])
+    shot_dist <- c(shot_dist, df[z, "SHOT_DIST"])
   }
@@
 #combine point type and distance
-valid_distance = data.frame(point_type, distance)
+valid_distance = data.frame(point_type, distance, shot_dist)
+dist_group <- c(5, 10, 15, 20, 25, Inf)
+valid_distance[, "Range Group"] <- c(findInterval(valid_distance$distance, dist_group))
 
@@
 cor(point_type, distance)
+
+#calculate correlation between defender distance and shot distance (2 or 3 points)
+cor(distance, shot_dist)
 
@@
   ggtitle("Points Scored based on Distance from Defender")
+
+ggplot(data = valid_distance, aes(x = distance, y = shot_dist)) + geom_point()
+
+ggplot(data = valid_distance, aes(x = distance) + geom_bar())
 
@@
 player2point <- data.frame(players2, distance2)
+
 summary(player2point)
+
+#calculate mean distance from 2-pointers
+avg_dist2 <- aggregate(distance2 ~ players2, player2point, mean)
 
@@
 top5_2points <- data.frame(top5players2, top5Counter)
+
+#count players who scored 2 points
+counter <- as.data.frame(table(players2))
 

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
Found FGA and FG% for each individual player

--- 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
@@
 
-# Trying to reshape...
+# Reshape the dataframe
 
-library(reshape)
-game <- cast(data = df, GAME_ID1 ~ SHOT_RESULT)
-View(game)
+game <- cast(data = df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + LOCATION + player_name + W ~ FGM)
+
+# Field Goal Attempts
+
+game[, "FGA"] <- NA
+for(i in 1:nrow(game)){
+  game[i, "FGA"] <- game[i, "1"] + game[i, "0"]
+}
+
+# find Feild Goal % of each indivudual player in that match
+
+game[, "FG%"] <- NA 
+for (i in 1:nrow(game)){
+  game[i, "FG%"] <- round(((game[i, "1"]/game[i, "FGA"]) * 100), digits = 2)
+}
 
@@
 
-Team <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
-FG_Percentage <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
-for(i in 21400001:21400908){
-  j = i - 21400001 +1 # 
-  game <- df[, df[, "GAME_ID1"] == i]
-  home <- game[, game["LOCATION"] == "H"]
-  home.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / nrow(home)
-  away <- game[, game["LOCATION"] == "A"]
-  away.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / away(home)
-  FG_Percentage[j] <- home.p
-  FG_Percentage[j + 1] = away.p
-  Team[j] = "H"
-  Team[j + 1] = "A"
-}
+#Team <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
+#FG_Percentage <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
+#for(i in 21400001:21400908){
+#  j = i - 21400001 +1 # 
+#  game <- df[, df[, "GAME_ID1"] == i]
+#  home <- game[, game["LOCATION"] == "H"]
+#  home.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / nrow(home)
+#  away <- game[, game["LOCATION"] == "A"]
+#  away.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / away(home)
+#  FG_Percentage[j] <- home.p
+#  FG_Percentage[j + 1] = away.p
+#  Team[j] = "H"
+#  Team[j + 1] = "A"
+#}
 

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
Added anaysis on accuracy

@@
 
+#Close Game Effect (Justin)
+
+```{r}
+# create new variable to seperate overtime
+
+df[, "OVERTIME"] <- as.numeric()
+
+for (i in 1:nrow(df)) {
+  if (df[i, "PERIOD"] >= 5) df[i, "OVERTIME"] <- 1
+  if (df[i, "PERIOD"] <= 4) df[i, "OVERTIME"] <- 0
+}
+
+# accuracy 
+acc_ot <- as.numeric
+acc_ot <- sum(df$OVERTIME == 1 & df$FGM == 1)/sum(df$OVERTIME == 1)
+
+acc_qrt1 <- as.numeric
+acc_qrt1 <- sum(df$PERIOD == 1 & df$FGM == 1)/sum(df$PERIOD == 1)
+
+acc_qrt2 <- as.numeric
+acc_qrt2 <- sum(df$PERIOD == 2 & df$FGM == 1)/sum(df$PERIOD == 2)
+
+acc_qrt3 <- as.numeric
+acc_qrt3 <- sum(df$PERIOD == 3 & df$FGM == 1)/sum(df$PERIOD == 3)
+
+acc_qrt4 <- as.numeric
+acc_qrt4 <- sum(df$PERIOD == 4 & df$FGM == 1)/sum(df$PERIOD == 4)
+
+acc_qrt5 <- as.numeric
+acc_qrt5 <- sum(df$PERIOD == 5 & df$FGM == 1)/sum(df$PERIOD == 5)
+
+acc_qrt6 <- as.numeric
+acc_qrt6 <- sum(df$PERIOD == 6 & df$FGM == 1)/sum(df$PERIOD == 6)
+
+acc_qrt7 <- as.numeric
+acc_qrt7 <- sum(df$PERIOD == 7 & df$FGM == 1)/sum(df$PERIOD == 7)
+
+acc_qrt1_4 <- as.numeric
+acc_qrt1_4 <- sum(df$PERIOD <= 4 & df$FGM == 1) / sum(df$PERIOD <= 4)
+
+t.test(df$FGM[df$PERIOD <= 4], df$FGM[df$PERIOD >= 5])
+t.test(df$FGM[df$PERIOD == 1], df$FGM[df$PERIOD == 2])
+t.test(df$FGM[df$PERIOD == 2], df$FGM[df$PERIOD == 3])
+t.test(df$FGM[df$PERIOD == 3], df$FGM[df$PERIOD == 4])
+t.test(df$FGM[df$PERIOD == 1], df$FGM[df$PERIOD == 3])
+t.test(df$FGM[df$PERIOD == 2], df$FGM[df$PERIOD == 4])
+
+# accuracy for Shot Clock
+
+  # dealing with missing values
+df[, "SHOT_CLOCK_NONA"] <- df[, "SHOT_CLOCK"]
+is.na(df$SHOT_CLOCK_NONA) <- 99
+
+acc_sec24 <- as.numeric
+acc_sec24 <- sum(df$SHOT_CLOCK > 23 & df$FGM == 1) / sum(df$SHOT_CLOCK > 23)
+
+```
+
 # Shooting Percentage Analysis (James)
@@
 
-# Reshape the dataframe
+# Trying to reshape...
 
-game <- cast(data = df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + LOCATION + player_name + W ~ FGM)
-
-# Field Goal Attempts
-
-game[, "FGA"] <- NA
-for(i in 1:nrow(game)){
-  game[i, "FGA"] <- game[i, "1"] + game[i, "0"]
-}
-
-# find Feild Goal % of each indivudual player in that match
-
-game[, "FG%"] <- NA 
-for (i in 1:nrow(game)){
-  game[i, "FG%"] <- round(((game[i, "1"]/game[i, "FGA"]) * 100), digits = 2)
-}
+library(reshape)
+game <- cast(data = df, GAME_ID1 ~ SHOT_RESULT)
+View(game)
 
@@
 
-#Team <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
-#FG_Percentage <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
-#for(i in 21400001:21400908){
-#  j = i - 21400001 +1 # 
-#  game <- df[, df[, "GAME_ID1"] == i]
-#  home <- game[, game["LOCATION"] == "H"]
-#  home.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / nrow(home)
-#  away <- game[, game["LOCATION"] == "A"]
-#  away.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / away(home)
-#  FG_Percentage[j] <- home.p
-#  FG_Percentage[j + 1] = away.p
-#  Team[j] = "H"
-#  Team[j + 1] = "A"
-#}
+Team <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
+FG_Percentage <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
+for(i in 21400001:21400908){
+  j = i - 21400001 +1 # 
+  game <- df[, df[, "GAME_ID1"] == i]
+  home <- game[, game["LOCATION"] == "H"]
+  home.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / nrow(home)
+  away <- game[, game["LOCATION"] == "A"]
+  away.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / away(home)
+  FG_Percentage[j] <- home.p
+  FG_Percentage[j + 1] = away.p
+  Team[j] = "H"
+  Team[j + 1] = "A"
+}
 

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
PART 3 - HOT HAND THEORY

--- 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
@@
 distance <- c()
-shot_dist <- c()
 for (z in 1:nrow(df)){
@@
     distance <- c(distance, df[z, "CLOSE_DEF_DIST"])
-    shot_dist <- c(shot_dist, df[z, "SHOT_DIST"])
   }
@@
 #combine point type and distance
-valid_distance = data.frame(point_type, distance, shot_dist)
-dist_group <- c(5, 10, 15, 20, 25, Inf)
-valid_distance[, "Range Group"] <- c(findInterval(valid_distance$distance, dist_group))
+valid_distance = data.frame(point_type, distance)
 
@@
 cor(point_type, distance)
-
-#calculate correlation between defender distance and shot distance (2 or 3 points)
-cor(distance, shot_dist)
 
@@
   ggtitle("Points Scored based on Distance from Defender")
-
-ggplot(data = valid_distance, aes(x = distance, y = shot_dist)) + geom_point()
-
-ggplot(data = valid_distance, aes(x = distance) + geom_bar())
 
@@
 player2point <- data.frame(players2, distance2)
-
 summary(player2point)
-
-#calculate mean distance from 2-pointers
-avg_dist2 <- aggregate(distance2 ~ players2, player2point, mean)
 
@@
 top5_2points <- data.frame(top5players2, top5Counter)
-
-#count players who scored 2 points
-counter <- as.data.frame(table(players2))
 
@@
 }
-player_and_game(202681, 21400681)
 
+#Set a new data.frame for the 10 players with best performances(higher points)
 ourtable <- head(pttsum2,10)
 
-list_per_player <- list()
+#Create a list that gives us the streaks of each player and game (10 best performances)
+streak_per_player <- list()
 j <- 1
@@
 streak_per_player
+
+#Create a data frame for the strikes of each one of the 10 players (cannot create a data.frame including all of them, because each list has different length)
+player1_df <- data.frame('streaks' = streak_per_player[[1]])
+player2_df <- data.frame('streaks' = streak_per_player[[2]])
+player3_df <- data.frame('streaks' = streak_per_player[[3]])
+player4_df <- data.frame('streaks' = streak_per_player[[4]])
+player5_df <- data.frame('streaks' = streak_per_player[[5]])
+player6_df <- data.frame('streaks' = streak_per_player[[6]])
+player7_df <- data.frame('streaks' = streak_per_player[[7]])
+player8_df <- data.frame('streaks' = streak_per_player[[8]])
+player9_df <- data.frame('streaks' = streak_per_player[[9]])
+player10_df <- data.frame('streaks' = streak_per_player[[10]])
+ourtable
+
+#Visualisation for 10 players - distribution of each players' streaklengths
+ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[1,'player_name'])
+ggplot(data = player2_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[2,'player_name'])
+ggplot(data = player3_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[3,'player_name'])
+ggplot(data = player4_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[4,'player_name'])
+ggplot(data = player5_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[5,'player_name'])
+ggplot(data = player6_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[6,'player_name'])
+ggplot(data = player7_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[7,'player_name'])
+ggplot(data = player8_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[8,'player_name'])
+ggplot(data = player9_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[9,'player_name'])
+ggplot(data = player10_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[10,'player_name'])
+
+#We can see that the distribution of all players is right skewed and that all of them (especially player3) have some long shooting streaks.
+To prove/disprove the hot hand theory we need to examine the indipendence of the shots.
+If each shot that a player takes is independent of the next shot, then a missed/made shot will not affect the propability of succesfulness of the next shot.
+So, if consecutive shots are independent, we disprove hot hand theory, while if consecutive shots are dependent, we prove the hot hand theory.
+#Simulation
+shot_results <- c('made', 'missed')
+simul_shooter <- sample(shot_results, size = nrow(player1_df), replace = TRUE, prob = c(0.45, 0.55))
+simul_shooter
+```
 

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
calculate shooting distance vs accuracy

--- 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
@@
 df[, "PTS_TYPE"] <- as.integer(df[, "PTS_TYPE"])
-df[, "CLOSE_DEF_DIS"] <- as.numeric(df[, "CLOSE_DEF_DIST"])
+df[, "CLOSE_DEF_DIST"] <- as.character(df[, "CLOSE_DEF_DIST"])
+
+#Calculate accuracy against defender distance
+#Group the defender distance level
+distance_level <- as.data.frame(table(factor(ClosestDefenderDistance)))
+head(distance_level)
+
+#Calculate how many times players scored in every defender distance
+for (i in 1:nrow(distance_level)){
+  distance_level[i, "SCORED"] <- sum(df[df[, "CLOSE_DEF_DIST"] == distance_level[i, "Var1"], "FGM"])
+}
+
+#Calculate shooting accuracy for each distance
+distance_level$Accuracy <- round(distance_level$SCORED / distance_level$Freq, 2)
+
+#Plotting the result
+library(ggplot2)
+ggplot(data = distance_level, aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
+  geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy))
+
+#Calculate accuracy against shooting distance
+#Group the shooting distance level
+shooting_distance_level <- as.data.frame(table(factor(ShotDistance)))
+head(shooting_distance_level)
+
+#Calculate how many times players scored in every shooting distance
+for (i in 1:nrow(shooting_distance_level)){
+  shooting_distance_level[i, "SCORED"] <- sum(df[df[, "SHOT_DIST"] == shooting_distance_level[i, "Var1"], "FGM"])
+}
+
+#Calculate shooting accuracy for each distance
+shooting_distance_level$Accuracy <- round(shooting_distance_level$SCORED / shooting_distance_level$Freq, 2)
+
+#Plotting the result
+library(ggplot2)
+ggplot(data = shooting_distance_level, aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) +
+  geom_point(aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy))
 
@@
 #combine point type and distance
+
+valid_distance = data.frame(point_type, distance, shot_dist)
+dist_group <- c(3, 6, 9, 12, 15, Inf)
+valid_distance[, "Range_Group"] <- c(findInterval(valid_distance$distance, dist_group))
+
 valid_distance = data.frame(point_type, distance)
@@
   ggtitle("Points Scored based on Distance from Defender")
+
+ggplot(data = valid_distance, aes(x = Range_Group)) + geom_bar()
+
+ggplot(data = valid_distance, aes(x = distance) + geom_bar())
 
@@
 
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
 ###Discussion
@@
 I   What would be the next steps?
----
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Widodo
+Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo
 

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
Added Accuracy by seconds left on Shot clock

--- 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
@@
   # dealing with missing values
-df[, "SHOT_CLOCK_NONA"] <- df[, "SHOT_CLOCK"]
-is.na(df$SHOT_CLOCK_NONA) <- 99
+df[is.na(df)] <- 99
 
-acc_sec24 <- as.numeric
-acc_sec24 <- sum(df$SHOT_CLOCK > 23 & df$FGM == 1) / sum(df$SHOT_CLOCK > 23)
+acc <- data.frame()
+acc[, "SECONDS_LEFT"] <- c(1:23)
+
+#accuracy for every second
+for (i in 1:23){
+  acc[i, "Accuracy"] <- (sum(df$SHOT_CLOCK > i & df$SHOT_CLOCK <= i+1 & df$SHOT_CLOCK != 99 & df$FGM == 1)/ sum(df$SHOT_CLOCK > i & df$SHOT_CLOCK <= i+1 & df$SHOT_CLOCK != 99))}
+
+ggplot(acc, aes(x = SECONDS_LEFT, y = Accuracy)) + geom_point() + geom_smooth()
 

### 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
fixed mistake - hot hand theory

--- 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
@@
 
-## Installing libraries
-
+##Installing packages and calling libraries
 ```{r}
 install.packages('reshape')
-
-```
-
-## Calling libraries
-```{r}
 library(reshape)
@@
 ```
-
 ### Home Advantage analysis
-
 ```{r}
@@
 #TO BE UPDATED  
-```
-
-#Close Game Effect (Justin)
-
-```{r}
-# create new variable to seperate overtime
-
-df[, "OVERTIME"] <- as.numeric()
-
-for (i in 1:nrow(df)) {
-  if (df[i, "PERIOD"] >= 5) df[i, "OVERTIME"] <- 1
-  if (df[i, "PERIOD"] <= 4) df[i, "OVERTIME"] <- 0
-}
-
-# accuracy 
-acc_ot <- as.numeric
-acc_ot <- sum(df$OVERTIME == 1 & df$FGM == 1)/sum(df$OVERTIME == 1)
-
-acc_qrt1 <- as.numeric
-acc_qrt1 <- sum(df$PERIOD == 1 & df$FGM == 1)/sum(df$PERIOD == 1)
-
-acc_qrt2 <- as.numeric
-acc_qrt2 <- sum(df$PERIOD == 2 & df$FGM == 1)/sum(df$PERIOD == 2)
-
-acc_qrt3 <- as.numeric
-acc_qrt3 <- sum(df$PERIOD == 3 & df$FGM == 1)/sum(df$PERIOD == 3)
-
-acc_qrt4 <- as.numeric
-acc_qrt4 <- sum(df$PERIOD == 4 & df$FGM == 1)/sum(df$PERIOD == 4)
-
-acc_qrt5 <- as.numeric
-acc_qrt5 <- sum(df$PERIOD == 5 & df$FGM == 1)/sum(df$PERIOD == 5)
-
-acc_qrt6 <- as.numeric
-acc_qrt6 <- sum(df$PERIOD == 6 & df$FGM == 1)/sum(df$PERIOD == 6)
-
-acc_qrt7 <- as.numeric
-acc_qrt7 <- sum(df$PERIOD == 7 & df$FGM == 1)/sum(df$PERIOD == 7)
-
-acc_qrt1_4 <- as.numeric
-acc_qrt1_4 <- sum(df$PERIOD <= 4 & df$FGM == 1) / sum(df$PERIOD <= 4)
-
-t.test(df$FGM[df$PERIOD <= 4], df$FGM[df$PERIOD >= 5])
-t.test(df$FGM[df$PERIOD == 1], df$FGM[df$PERIOD == 2])
-t.test(df$FGM[df$PERIOD == 2], df$FGM[df$PERIOD == 3])
-t.test(df$FGM[df$PERIOD == 3], df$FGM[df$PERIOD == 4])
-t.test(df$FGM[df$PERIOD == 1], df$FGM[df$PERIOD == 3])
-t.test(df$FGM[df$PERIOD == 2], df$FGM[df$PERIOD == 4])
-
-# accuracy for Shot Clock
-
-  # dealing with missing values
-df[is.na(df)] <- 99
-
-acc <- data.frame()
-acc[, "SECONDS_LEFT"] <- c(1:23)
-
-#accuracy for every second
-for (i in 1:23){
-  acc[i, "Accuracy"] <- (sum(df$SHOT_CLOCK > i & df$SHOT_CLOCK <= i+1 & df$SHOT_CLOCK != 99 & df$FGM == 1)/ sum(df$SHOT_CLOCK > i & df$SHOT_CLOCK <= i+1 & df$SHOT_CLOCK != 99))}
-
-ggplot(acc, aes(x = SECONDS_LEFT, y = Accuracy)) + geom_point() + geom_smooth()
-
 ```
@@
 
-shotresult
-
 #Set the value to numeric/integer
@@
 streak_per_player
+
+#Create a data frame for the strikes of each one of the 10 players (cannot create a data.frame including all of them, because each list has different length)
+player1_df <- data.frame('streaks' = streak_per_player[[1]])
+player2_df <- data.frame('streaks' = streak_per_player[[2]])
+player3_df <- data.frame('streaks' = streak_per_player[[3]])
+player4_df <- data.frame('streaks' = streak_per_player[[4]])
+player5_df <- data.frame('streaks' = streak_per_player[[5]])
+player6_df <- data.frame('streaks' = streak_per_player[[6]])
+player7_df <- data.frame('streaks' = streak_per_player[[7]])
+player8_df <- data.frame('streaks' = streak_per_player[[8]])
+player9_df <- data.frame('streaks' = streak_per_player[[9]])
+player10_df <- data.frame('streaks' = streak_per_player[[10]])
+ourtable
+
+#Visualisation for 10 players - distribution of each players' streaklengths
+ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[1,'player_name'])
+ggplot(data = player2_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[2,'player_name'])
+ggplot(data = player3_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[3,'player_name'])
+ggplot(data = player4_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[4,'player_name'])
+ggplot(data = player5_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[5,'player_name'])
+ggplot(data = player6_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[6,'player_name'])
+ggplot(data = player7_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[7,'player_name'])
+ggplot(data = player8_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[8,'player_name'])
+ggplot(data = player9_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[9,'player_name'])
+ggplot(data = player10_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[10,'player_name'])
+
+#We can see that the distribution of all players is right skewed and that all of them (especially player3) have some long shooting streaks.
+To prove/disprove the hot hand theory we need to examine the indipendence of the shots.
+If each shot that a player takes is independent of the next shot, then a missed/made shot will not affect the propability of succesfulness of the next shot.
+So, if consecutive shots are independent, we disprove hot hand theory, while if consecutive shots are dependent, we prove the hot hand theory.
+#Simulation
+shot_results <- c('made', 'missed')
+simul_shooter <- sample(shot_results, size = nrow(player1_df), replace = TRUE, prob = c(0.45, 0.55))
+simul_shooter
+
+simul_shooter <- as.list(simul_shooter)
+for(i in 1:length(simul_shooter)) {
+   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
+   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0
+
+}
+simul_shooter
+number_of_streaks(simul_shooter)
+```
 

### 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
minor changes

--- 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
@@
 }
-
-#
-
-#Team <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
-#FG_Percentage <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
-#for(i in 21400001:21400908){
-#  j = i - 21400001 +1 # 
-#  game <- df[, df[, "GAME_ID1"] == i]
-#  home <- game[, game["LOCATION"] == "H"]
-#  home.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / nrow(home)
-#  away <- game[, game["LOCATION"] == "A"]
-#  away.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / away(home)
-#  FG_Percentage[j] <- home.p
-#  FG_Percentage[j + 1] = away.p
-#  Team[j] = "H"
-#  Team[j + 1] = "A"
-#}
-
 ```

### 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
Impact of Super Star Players

@@
 
-#
+# Find shooting percentage of each match
 
 ```{r}
-# Find shooting percentage of each match
-
 # Reshape the dataframe
@@
 
-# find Feild Goal % of each indivudual player in that match
+# Find Feild Goal Percentage of each indivudual player in that match
 
-game[, "FG%"] <- NA 
+game[, "FG.percentage"] <- NA 
 for (i in 1:nrow(game)){
-  game[i, "FG%"] <- round(((game[i, "1"]/game[i, "FGA"]) * 100), digits = 2)
+  game[i, "FG.percentage"] <- round(((game[i, "1"]/game[i, "FGA"]) * 100), digits = 2)
 }
+```
+
+# Impact of Super Star Players
+
+```{r}
+
+# Pick out 8 of the best players in that regular season and create a new subset. Analyse how shooting percentage of these players can influence the result of a match.
+
+super.star.list <- c("lebron james", "james harden", "russell westbrook", "stephen curry", "anthony davis", "blake griffin", "klay thompson", "lamarcus aldridge") # players picked out
+super.star <- game[game$player_name %in% super.star.list, c("player_name", "W", "FGA", "FG.percentage")] # create a subset for these players
+
+ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
+ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W))
+
+# Replace "W" and "L" with 1 and 0 respectively, 
+super.star[, "W"] <- as.character(super.star[, "W"])
+for (i in 1:nrow(super.star)) {
+  if (super.star[i, "W"] == "W") super.star[i, "W"] <- 1
+  if (super.star[i, "W"] == "L") super.star[i, "W"] <- 0
+}
+super.star[, "W"] <- as.numeric(super.star[, "W"])
+
+# Correlation test
+t.test(super.star[, "W"], super.star[, "FGA"]) # this makes no sense...
+
+# Linear regression
+summary(lm(W ~ player_name + FGA, data = super.star))
+summary(lm(W ~ player_name + FG.percentage, data = super.star))
+
 ```

### 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
hot hand theory - continue

--- 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
@@
 
-#Set the value to character
-df[, "SHOT_RESULT"] <- as.character(df[, "SHOT_RESULT"])
-
-#Define 1 for baskets made and 0 for baskets missed. We need a vector to use it to our function below.
-shotresult <- c()
-for(i in 1:nrow(df)) {
-  if(shots[i] == 'missed') shotresult <- c(shotresult, 0)
-  else if(shots[i] == 'made') shotresult <- c(shotresult, 1)
-}
-
-shotresult
-
-#Set the value to numeric/integer
-df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
-df[ ,'SHOT_NUMBER'] <- as.integer(df[ , 'SHOT_NUMBER'])
-
+#From 'FGM' column we have that 1 represents a basket made and 0 represents a basket missed
 #Define the length of a shooting streak to be the number of consecutive baskets made until a miss occurs
-#Create a function that gives us a vector with the number of streaks
+#Create a function that gives us a vector with the number of total streaks(from all players and all games)
 number_of_streaks <- function(x) {
@@
 
-number_of_streaks(shotresult)
+number_of_streaks(df$FGM)
 
+#Set a new data frame for the 5 players with best performances(higher points)
+best_players_df <- head(pttsum2,5)
+
+#Create a function that gives us a vector of streaks for a specific player and game
 player_and_game <- function(x, y) {
@@
 }
-player_and_game(202681, 21400681)
+#For the first player and game try:
+#player_and_game('202681', '21400681')
 
-ourtable <- head(pttsum2,10)
-
-list_per_player <- list()
+#Create a list that gives us the streaks of each player and game (5 best performances)
+streak_per_player <- list()
 j <- 1
-for(i in ourtable$player_id) {
-   streak_per_player[[j]] <- player_and_game(i, ourtable[ourtable[ ,'player_id'] == i, 'GAME_ID1'])
+for(i in best_players_df$player_id) {
+   streak_per_player[[j]] <- player_and_game(i, best_players_df[ourtable[ ,'player_id'] == i, 'GAME_ID1'])
    j <- j + 1
@@
 streak_per_player
+
+#Create a data frame for the strikes of each one of the 10 players (cannot create a data frame including all of them, because each list has different length)
+player1_df <- data.frame('streaks' = streak_per_player[[1]])
+player2_df <- data.frame('streaks' = streak_per_player[[2]])
+player3_df <- data.frame('streaks' = streak_per_player[[3]])
+player4_df <- data.frame('streaks' = streak_per_player[[4]])
+player5_df <- data.frame('streaks' = streak_per_player[[5]])
+
+#Show distribution for each player's streak lengths
+ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[1,'player_name'])
+ggplot(data = player2_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[2,'player_name'])
+ggplot(data = player3_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[3,'player_name'])
+ggplot(data = player4_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[4,'player_name'])
+ggplot(data = player5_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[5,'player_name'])
+
+#Create a boxplot for each player's streak lengths
+boxplot(player1_df$streaks) 
+summary(player1_df$streaks) #Check that boxplot is valid
+IQR(player1_df$streaks) #Check that boxplot is valid
+#The typical length of a streak is 0 (median = 0) - The Interquartile Range is 1 - Streak lengths of 3 and 6 are unusually high compared to the rest of the distribution
+boxplot(player2_df$streaks)
+summary(player2_df$streaks)
+IQR(player2_df$streaks)
+#The typical length of a streak is 1 (median = 1) - The Interquartile Range is 2 - There are some high streak lengths (max = 4)
+boxplot(player3_df$streaks)
+summary(player3_df$streaks)
+IQR(player3_df$streaks)
+#The typical length of a streak is 0 (median = 0) - The Interquartile Range is 0 - Streak lengths of 3 and 13 (!!!) are unusually high compared to the rest of the distribution
+boxplot(player4_df$streaks)
+summary(player4_df$streaks)
+IQR(player4_df$streaks)
+#The typical length of a streak is 1 (median = 1) - The Interquartile Range is 2 - Streak length of 6 is unusually high compared to the rest of the distribution
+boxplot(player5_df$streaks)
+summary(player5_df$streaks)
+IQR(player5_df$streaks)
+#The typical length of a streak is 1 (median = 1) - The Interquartile Range is 2 - There are some high streak lengths (max = 5)
+
+#We can see that the distribution of all players is right skewed and that all of them (especially player3) have some long shooting streaks.
+To prove/disprove the hot hand theory we need to examine the indipendence of the shots.
+If each shot that a player takes is independent of the next shot, then a missed/made shot will not affect the propability that the next shot will be missed/made.
+So, if consecutive shots are independent, we disprove hot hand theory, while if consecutive shots are dependent, we prove the hot hand theory.
+#Simulation - create a vector of shots('made', 'missed') for a sample shooter
+set.seed(23)
+shot_results <- c('made', 'missed')
+simul_shooter <- sample(shot_results, size = nrow(player1_df), replace = TRUE, prob = c(0.45, 0.55))
+simul_shooter
+
+#Set 1 for baskets made and 0 for baskets missed
+simul_shooter <- as.list(simul_shooter)
+for(i in 1:length(simul_shooter)) {
+   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
+   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0
+}
+simul_shooter
+
+#Use previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
+streaks_of_simulshooter <- number_of_streaks(simul_shooter)
+simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
+
+#Comparisons of data between sample shooter and real shooters
+summary(streaks_of_simulshooter)
+summary(player1_df$streaks)
+
+#Compare barplots of sample shooter and real shooter1
+#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
+ggplot(data = simulshooter_df, aes(x = streaks)) + geom_bar() 
+ggplot(data = player1_df, aes(x = streaks)) + geom_bar()
+#
+```
 

### 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
df2 read csv updated. and working on regression

@@
 h. Fatigue Effect (Number of Rest Day)  
-i.
+i. Regression Analysis (Alex)
 
@@
 
+df2<- read.csv("shot_logs1010.csv")
+
 #we need to change the format of some of the columns to 'Character'
+#get rid off 'attach()'
+#get rid off column X in shot logs1010 file
 
@@
 
+```{r}
+#Regression Analysis
+colnames(df2)
+
+linreg_fgm <-lm(FGM ~ GAME_ID1 + MATCHUP+LOCATION+W+FINAL_MARGIN+SHOT_NUMBER+ PERIOD+GAME_CLOCK+  SHOT_CLOCK+  DRIBBLES+ TOUCH_TIME+SHOT_DIST+PTS_TYPE+ SHOT_RESULT+ CLOSEST_DEFENDER + CLOSEST_DEFENDER_PLAYER_ID + CLOSE_DEF_DIST +PTS+player_name+ player_id+HOME_TEAM+AWAY_TEAM+WINNER, df2 )
+
+summary(linreg_fgm)
+
+linreg_fgm2 <-lm(FGM ~ GAME_ID1 + MATCHUP+LOCATION+W+FINAL_MARGIN+SHOT_NUMBER+ PERIOD+GAME_CLOCK+  SHOT_CLOCK+  DRIBBLES+ TOUCH_TIME+SHOT_DIST+PTS_TYPE+ SHOT_RESULT+ CLOSEST_DEFENDER + CLOSEST_DEFENDER_PLAYER_ID + CLOSE_DEF_DIST +PTS+player_name+ player_id+HOME_TEAM+AWAY_TEAM+WINNER, df2 )
+
+summary(linreg_fgm)
+
+```
+
 ###Discussion

### 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
continued work on accuracy

--- 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
@@
 
+# accuracy for Top 10 scoring players
+
 # accuracy for Shot Clock
@@
 ggplot(acc, aes(x = SECONDS_LEFT, y = Accuracy)) + geom_point() + geom_smooth()
+
+df[, "ACC_P_SEC"] <- as.numeric()
+
+#Assign accuracy per second to main dataframe
+for (i in 1:nrow(df)){
+  for (j in 1:23){
+    if (df$SHOT_CLOCK[i] >j & df$SHOT_CLOCK[i] <= j+1 & df$SHOT_CLOCK[i] != 99) df[i, "ACC_P_SEC"] <- acc[j, "Accuracy"]
+  }
+}
+
+cov(df$ACC_P_SEC, df$SHOT_CLOCK)
+cor.test(df$ACC_P_SEC, df$SHOT_CLOCK)
 
@@
 
+#
+
+```{r}
 # Find shooting percentage of each match
 
-```{r}
-# Reshape the dataframe
+# Trying to reshape...
 
-game <- cast(data = df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + LOCATION + player_name + W ~ FGM)
+library(reshape)
+game <- cast(data = df, GAME_ID1 ~ SHOT_RESULT)
+View(game)
 
-# Field Goal Attempts
+#
 
-game[, "FGA"] <- NA
-for(i in 1:nrow(game)){
-  game[i, "FGA"] <- game[i, "1"] + game[i, "0"]
+Team <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
+FG_Percentage <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
+for(i in 21400001:21400908){
+  j = i - 21400001 +1 # 
+  game <- df[, df[, "GAME_ID1"] == i]
+  home <- game[, game["LOCATION"] == "H"]
+  home.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / nrow(home)
+  away <- game[, game["LOCATION"] == "A"]
+  away.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / away(home)
+  FG_Percentage[j] <- home.p
+  FG_Percentage[j + 1] = away.p
+  Team[j] = "H"
+  Team[j + 1] = "A"
 }
-
-# Find Feild Goal Percentage of each indivudual player in that match
-
-game[, "FG.percentage"] <- NA 
-for (i in 1:nrow(game)){
-  game[i, "FG.percentage"] <- round(((game[i, "1"]/game[i, "FGA"]) * 100), digits = 2)
-}
-```
-
-# Impact of Super Star Players
-
-```{r}
-
-# Pick out 8 of the best players in that regular season and create a new subset. Analyse how shooting percentage of these players can influence the result of a match.
-
-super.star.list <- c("lebron james", "james harden", "russell westbrook", "stephen curry", "anthony davis", "blake griffin", "klay thompson", "lamarcus aldridge") # players picked out
-super.star <- game[game$player_name %in% super.star.list, c("player_name", "W", "FGA", "FG.percentage")] # create a subset for these players
-
-ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
-ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W))
-
-# Replace "W" and "L" with 1 and 0 respectively, 
-super.star[, "W"] <- as.character(super.star[, "W"])
-for (i in 1:nrow(super.star)) {
-  if (super.star[i, "W"] == "W") super.star[i, "W"] <- 1
-  if (super.star[i, "W"] == "L") super.star[i, "W"] <- 0
-}
-super.star[, "W"] <- as.numeric(super.star[, "W"])
-
-# Correlation test
-t.test(super.star[, "W"], super.star[, "FGA"]) # this makes no sense...
-
-# Linear regression
-summary(lm(W ~ player_name + FGA, data = super.star))
-summary(lm(W ~ player_name + FG.percentage, data = super.star))
 
@@
 
+shotresult
+
 #Set the value to numeric/integer
@@
 
-#Create a function that gives us a vector of streaks for a specific player and game
 player_and_game <- function(x, y) {
@@
 }
+player_and_game(202681, 21400681)
 
-#Set a new data.frame for the 10 players with best performances(higher points)
 ourtable <- head(pttsum2,10)
 
-#Create a list that gives us the streaks of each player and game (10 best performances)
-streak_per_player <- list()
+list_per_player <- list()
 j <- 1
@@
 streak_per_player
-
-#Create a data frame for the strikes of each one of the 10 players (cannot create a data.frame including all of them, because each list has different length)
-player1_df <- data.frame('streaks' = streak_per_player[[1]])
-player2_df <- data.frame('streaks' = streak_per_player[[2]])
-player3_df <- data.frame('streaks' = streak_per_player[[3]])
-player4_df <- data.frame('streaks' = streak_per_player[[4]])
-player5_df <- data.frame('streaks' = streak_per_player[[5]])
-player6_df <- data.frame('streaks' = streak_per_player[[6]])
-player7_df <- data.frame('streaks' = streak_per_player[[7]])
-player8_df <- data.frame('streaks' = streak_per_player[[8]])
-player9_df <- data.frame('streaks' = streak_per_player[[9]])
-player10_df <- data.frame('streaks' = streak_per_player[[10]])
-ourtable
-
-#Visualisation for 10 players - distribution of each players' streaklengths
-ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[1,'player_name'])
-ggplot(data = player2_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[2,'player_name'])
-ggplot(data = player3_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[3,'player_name'])
-ggplot(data = player4_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[4,'player_name'])
-ggplot(data = player5_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[5,'player_name'])
-ggplot(data = player6_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[6,'player_name'])
-ggplot(data = player7_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[7,'player_name'])
-ggplot(data = player8_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[8,'player_name'])
-ggplot(data = player9_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[9,'player_name'])
-ggplot(data = player10_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[10,'player_name'])
-
-#We can see that the distribution of all players is right skewed and that all of them (especially player3) have some long shooting streaks.
-To prove/disprove the hot hand theory we need to examine the indipendence of the shots.
-If each shot that a player takes is independent of the next shot, then a missed/made shot will not affect the propability of succesfulness of the next shot.
-So, if consecutive shots are independent, we disprove hot hand theory, while if consecutive shots are dependent, we prove the hot hand theory.
-#Simulation
-shot_results <- c('made', 'missed')
-simul_shooter <- sample(shot_results, size = nrow(player1_df), replace = TRUE, prob = c(0.45, 0.55))
-simul_shooter
-
-simul_shooter <- as.list(simul_shooter)
-for(i in 1:length(simul_shooter)) {
-   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
-   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0
-
-}
-simul_shooter
-number_of_streaks(simul_shooter)
-```
 

### 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
add notes

@@
 author: "Team_9"
-output: html_document  ---
-  title: "MSF_Groupwork-NBA_shotlog"
-author: "Team_9"
-output: html_document 
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 
-###Introduction
+##Introduction
 I   What is your theme or research question?
@@
 
-###Theory
+##Theory
 I   What do you expect to find?
@@
 
-###Methods
+##Methods
 I   Describe the dataset:
@@
 Useful for evaluating who the best shooter is, who the best defender is, the hot-hand hypothesis, etc.
-Scraped from NBA's REST API.  ##TO BE UPDATED FURTHER
+Scraped from NBA's REST API.  It has 128,609 rows and 22 columns with 16 MB data size. ##TO BE UPDATED FURTHER
 
@@
 
+    The dataset comprise of comprehensive observational data to start to work with.
+
 I   Limitations
+    The dataset consists only for 1 year NBA season with regular match (82 matches per team for both home and away matches). The playoff and other matches during the year is not included.
 
-###Analysis
+##Analysis
 
-## Installing libraries
+### Installing libraries
 
 ```{r}
-install.packages('reshape')
+#install.packages('reshape')
 
@@
 
-## Calling libraries
+### Calling libraries
 ```{r}
@@
 
-# accuracy for Top 10 scoring players
-
 # accuracy for Shot Clock
@@
 ggplot(acc, aes(x = SECONDS_LEFT, y = Accuracy)) + geom_point() + geom_smooth()
-
-df[, "ACC_P_SEC"] <- as.numeric()
-
-#Assign accuracy per second to main dataframe
-for (i in 1:nrow(df)){
-  for (j in 1:23){
-    if (df$SHOT_CLOCK[i] >j & df$SHOT_CLOCK[i] <= j+1 & df$SHOT_CLOCK[i] != 99) df[i, "ACC_P_SEC"] <- acc[j, "Accuracy"]
-  }
-}
-
-cov(df$ACC_P_SEC, df$SHOT_CLOCK)
-cor.test(df$ACC_P_SEC, df$SHOT_CLOCK)
 
@@
 
-#
+# Find shooting percentage of each match
 
 ```{r}
-# Find shooting percentage of each match
+# Reshape the dataframe
 
-# Trying to reshape...
+game <- cast(data = df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + LOCATION + player_name + W ~ FGM)
 
-library(reshape)
-game <- cast(data = df, GAME_ID1 ~ SHOT_RESULT)
-View(game)
+# Field Goal Attempts
 
-#
+game[, "FGA"] <- NA
+for(i in 1:nrow(game)){
+  game[i, "FGA"] <- game[i, "1"] + game[i, "0"]
+}
 
-Team <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
-FG_Percentage <- rep(NA, (lastGamesID - firstGamesID + 1) * 2)
-for(i in 21400001:21400908){
-  j = i - 21400001 +1 # 
-  game <- df[, df[, "GAME_ID1"] == i]
-  home <- game[, game["LOCATION"] == "H"]
-  home.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / nrow(home)
-  away <- game[, game["LOCATION"] == "A"]
-  away.p <- count(home[, home[, "SHOT_RESULT"] == 1]) / away(home)
-  FG_Percentage[j] <- home.p
-  FG_Percentage[j + 1] = away.p
-  Team[j] = "H"
-  Team[j + 1] = "A"
+# Find Feild Goal Percentage of each indivudual player in that match
+
+game[, "FG.percentage"] <- NA 
+for (i in 1:nrow(game)){
+  game[i, "FG.percentage"] <- round(((game[i, "1"]/game[i, "FGA"]) * 100), digits = 2)
 }
+```
+
+# Impact of Super Star Players
+
+```{r}
+
+# Pick out 8 of the best players in that regular season and create a new subset. Analyse how shooting percentage of these players can influence the result of a match.
+
+super.star.list <- c("lebron james", "james harden", "russell westbrook", "stephen curry", "anthony davis", "blake griffin", "klay thompson", "lamarcus aldridge") # players picked out
+super.star <- game[game$player_name %in% super.star.list, c("player_name", "W", "FGA", "FG.percentage")] # create a subset for these players
+
+ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
+ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W))
+
+# Replace "W" and "L" with 1 and 0 respectively, 
+super.star[, "W"] <- as.character(super.star[, "W"])
+for (i in 1:nrow(super.star)) {
+  if (super.star[i, "W"] == "W") super.star[i, "W"] <- 1
+  if (super.star[i, "W"] == "L") super.star[i, "W"] <- 0
+}
+super.star[, "W"] <- as.numeric(super.star[, "W"])
+
+# Correlation test
+t.test(super.star[, "W"], super.star[, "FGA"]) # this makes no sense...
+
+# Linear regression
+summary(lm(W ~ player_name + FGA, data = super.star))
+summary(lm(W ~ player_name + FG.percentage, data = super.star))
 
@@
 
-###Discussion
-I   What results did you find?
-
-I   Why is this interesting?
-
-I   What would be the next steps?
----
-
-```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)

### 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
completed regression analysis. Updated Intro/summary

--- 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
@@
 ```
-Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Widodo
+Report produced by: Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Widodo  
+Date submitted: 16/10/2016
 
@@
 
-###Introduction
-I   What is your theme or research question?
+###Introduction/Summary
+
+The purpose of this report is to analyse a chosen dataset and to produce meaningful results using statistical and mathematical analysis.
+These analyses have been conducted using RStudio software.
+This report will focused on analysis of a number of areas, demonstrate how the analysis has been conducted, discusses theory used, demonstrate and evaluate produced results.
+In addition this report will highlight potential areas of future research.
+
+The chosen dataset for this report is a summary of shots made during NBA season 2015-2016. The following areas of research have been selected:
+
+#TO BE UPDATED when we know final areas
 a. Hot Hand Theory (Artemis)
@@
 
-df2<- read.csv("shot_logs1010.csv")
+df1<- read.csv("shot_logs1010.csv")
 
@@
 #Regression Analysis
-colnames(df2)
 
-linreg_fgm <-lm(FGM ~ GAME_ID1 + MATCHUP+LOCATION+W+FINAL_MARGIN+SHOT_NUMBER+ PERIOD+GAME_CLOCK+  SHOT_CLOCK+  DRIBBLES+ TOUCH_TIME+SHOT_DIST+PTS_TYPE+ SHOT_RESULT+ CLOSEST_DEFENDER + CLOSEST_DEFENDER_PLAYER_ID + CLOSE_DEF_DIST +PTS+player_name+ player_id+HOME_TEAM+AWAY_TEAM+WINNER, df2 )
+df1$GAME_CLOCK <-as.numeric(df1$GAME_CLOCK)
+
+df1$AWAY_TEAM <- as.character(df1$AWAY_TEAM)
+
+linreg_fgm <-lm(FGM ~ LOCATION+W+FINAL_MARGIN+SHOT_NUMBER+ PERIOD+GAME_CLOCK+  SHOT_CLOCK+  DRIBBLES+ TOUCH_TIME+SHOT_DIST+PTS_TYPE+ CLOSEST_DEFENDER_PLAYER_ID + CLOSE_DEF_DIST+ player_id+ WINNER, df1 )
 
@@
 
-linreg_fgm2 <-lm(FGM ~ GAME_ID1 + MATCHUP+LOCATION+W+FINAL_MARGIN+SHOT_NUMBER+ PERIOD+GAME_CLOCK+  SHOT_CLOCK+  DRIBBLES+ TOUCH_TIME+SHOT_DIST+PTS_TYPE+ SHOT_RESULT+ CLOSEST_DEFENDER + CLOSEST_DEFENDER_PLAYER_ID + CLOSE_DEF_DIST +PTS+player_name+ player_id+HOME_TEAM+AWAY_TEAM+WINNER, df2 )
+linreg_fgm2 <-lm(FGM ~ FINAL_MARGIN+SHOT_NUMBER+ PERIOD +  SHOT_CLOCK+  DRIBBLES+ TOUCH_TIME+SHOT_DIST+PTS_TYPE+ CLOSE_DEF_DIST+ player_id, df1 )
 
-summary(linreg_fgm)
+summary(linreg_fgm2)
 

### 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
add explanantion for dataset description

@@
 ```
-Report produced by: Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Widodo  
-Date submitted: 16/10/2016
+Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Widodo
 
@@
 
-###Introduction/Summary
-
-The purpose of this report is to analyse a chosen dataset and to produce meaningful results using statistical and mathematical analysis.
-These analyses have been conducted using RStudio software.
-This report will focused on analysis of a number of areas, demonstrate how the analysis has been conducted, discusses theory used, demonstrate and evaluate produced results.
-In addition this report will highlight potential areas of future research.
-
-The chosen dataset for this report is a summary of shots made during NBA season 2015-2016. The following areas of research have been selected:
-
-#TO BE UPDATED when we know final areas
+##Introduction
+###What is your theme or research question?
     a. Hot Hand Theory (Artemis)
@@
 
-###Theory
-I   What do you expect to find?
+##Theory
+###What do you expect to find?
     a.
 
-###Methods
+##Methods
 I   Describe the dataset:
 
-  Data on shots taken during the 2014-2015 season, who took the shot, where on the floor was the shot taken from, who was the nearest defender, how far away was the nearest defender, time on the shot clock, and much more.
+    Data on shots taken during the NBA 2014-2015 season for regular matches throughout the year.
+This comprehensive data comprise of 128,609 rows and 22 columns with 16 MB data size with only several value in a column is missing (shot clock).
 The column titles are generally self-explanatory.
-Useful for evaluating who the best shooter is, who the best defender is, the hot-hand hypothesis, etc.
-Scraped from NBA's REST API.  ##TO BE UPDATED FURTHER
+It has the following data:
+    1.  Match ID  
+    2.  Match date  
+    3.  Which team plays as home and away  
+    4.  Match result  
+    5.  Final score difference  
+    6.  Every shots per game for each player  
+    7.  Shot result whether it is success or not  
+    8.  Shot attempt for 2 points or 3 points  
+    9.  What is the game clock  
+    10. How much time left for a shot (shot clock)  
+    11. Who took the shot  
+    12. Shot distance from basket  
+    13. Who was the nearest defender  
+    14. Distance from the nearest defender  
+    The dataset is useful for evaluating who the best shooter is, who the best defender is, the hot-hand hypothesis, affect of home playing team, close game effect, shooting accuracy from best players, and defender distance effect on shooting accuracy.
+Data was taken from NBA's REST API provided by [Kaggle](https://www.kaggle.com/dansbecker/nba-shot-logs) and acquired at 30 September 2016.
 
@@
 #calculate correlation between defender distance and points made (2 or 3 points)
-cor(point_type, distance)
+cor.test(point_type, distance)
 
@@
 #calculate correlation between defender distance and shot result
-cor(df[, "SHOT_RESULT"], df[, "CLOSE_DEF_DIS"])
+cor.test(df[, "SHOT_RESULT"], df[, "CLOSE_DEF_DIS"])
 
@@
 
+#Calculate accuracy against defender distance
+#Group the defender distance level
+
+distance_level <- as.data.frame(table(factor(ClosestDefenderDistance)))
+head(distance_level)
+
+#Calculate how many times players scored in every defender distance
+for (i in 1:nrow(distance_level)){
+  distance_level[i, "SCORED"] <- sum(df[df[, "CLOSE_DEF_DIST"] == distance_level[i, "Var1"], "FGM"])
+}
+
+#Calculate shooting accuracy for each distance
+distance_level$Accuracy <- round(distance_level$SCORED / distance_level$Freq, 2)
+
+#Plotting the result
+library(ggplot2)
+ggplot(data = distance_level, aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
+  geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy))
+
+#----------------------------------------------------------------#
+
+#Calculate accuracy against shooting distance
+#Group the shooting distance level
+shooting_distance_level <- as.data.frame(table(factor(ShotDistance)))
+head(shooting_distance_level)
+
+#Calculate how many times players scored in every shooting distance
+for (i in 1:nrow(shooting_distance_level)){
+  shooting_distance_level[i, "SCORED"] <- sum(df[df[, "SHOT_DIST"] == shooting_distance_level[i, "Var1"], "FGM"])
+}
+
+#Calculate shooting accuracy for each distance
+shooting_distance_level$Accuracy <- round(shooting_distance_level$SCORED / shooting_distance_level$Freq, 2)
+
+#Plotting the result
+library(ggplot2)
+ggplot(data = shooting_distance_level, aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) +
+  geom_point(aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy))
 ```
@@
 #Regression Analysis
+colnames(df2)
 
-df1$GAME_CLOCK <-as.numeric(df1$GAME_CLOCK)
-
-df1$AWAY_TEAM <- as.character(df1$AWAY_TEAM)
-
-linreg_fgm <-lm(FGM ~ LOCATION+W+FINAL_MARGIN+SHOT_NUMBER+ PERIOD+GAME_CLOCK+  SHOT_CLOCK+  DRIBBLES+ TOUCH_TIME+SHOT_DIST+PTS_TYPE+ CLOSEST_DEFENDER_PLAYER_ID + CLOSE_DEF_DIST+ player_id+ WINNER, df1 )
+linreg_fgm <-lm(FGM ~ GAME_ID1 + MATCHUP+LOCATION+W+FINAL_MARGIN+SHOT_NUMBER+ PERIOD+GAME_CLOCK+  SHOT_CLOCK+  DRIBBLES+ TOUCH_TIME+SHOT_DIST+PTS_TYPE+ SHOT_RESULT+ CLOSEST_DEFENDER + CLOSEST_DEFENDER_PLAYER_ID + CLOSE_DEF_DIST +PTS+player_name+ player_id+HOME_TEAM+AWAY_TEAM+WINNER, df2 )
 
@@
 
-linreg_fgm2 <-lm(FGM ~ FINAL_MARGIN+SHOT_NUMBER+ PERIOD +  SHOT_CLOCK+  DRIBBLES+ TOUCH_TIME+SHOT_DIST+PTS_TYPE+ CLOSE_DEF_DIST+ player_id, df1 )
+linreg_fgm2 <-lm(FGM ~ GAME_ID1 + MATCHUP+LOCATION+W+FINAL_MARGIN+SHOT_NUMBER+ PERIOD+GAME_CLOCK+  SHOT_CLOCK+  DRIBBLES+ TOUCH_TIME+SHOT_DIST+PTS_TYPE+ SHOT_RESULT+ CLOSEST_DEFENDER + CLOSEST_DEFENDER_PLAYER_ID + CLOSE_DEF_DIST +PTS+player_name+ player_id+HOME_TEAM+AWAY_TEAM+WINNER, df2 )
 
-summary(linreg_fgm2)
+summary(linreg_fgm)
 

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
enhance plot for defender distance vs accuracy

--- 2016-10-11 MSFGroupwork-NBAHotHand.Rmd
@@
 ```
-Report produced by: Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Widodo  
+Report produced by: Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo  
 Date submitted: 16/10/2016
@@
 #Report structure
+
+##Introduction
+###What is your theme or research question?
+    a. Hot Hand Theory (Artemis)
+    b. Defender Proximity Analysis (Reza)
+    c. Shot Clock Pressure  (Alex)
+    d. Close Game Effect - Small Margin affecting the scoring rate 
+    e. Home Advantage (Alex is working in this point)  
+    f. Fatigue Effect (Quarter Heat Rate)  
+    g. Shooting Percentage (James)
+    h. Fatigue Effect (Number of Rest Day)  
+    i. Regression Analysis (Alex)
 
@@
 
-#TO BE UPDATED when we know final areas
+#Report structure
+
+##Introduction
+I   What is your theme or research question?
 a. Hot Hand Theory (Artemis)
@@
 
-###Methods
-I   Describe the dataset:
+##Methods
+###Dataset Description:
 
-  Data on shots taken during the 2014-2015 season, who took the shot, where on the floor was the shot taken from, who was the nearest defender, how far away was the nearest defender, time on the shot clock, and much more.
+    Data on shots taken during the NBA 2014-2015 season for regular matches throughout the year.
+The data comprise of 128,609 rows and 22 columns with 16 MB data size with several value in some column is missing.
 The column titles are generally self-explanatory.
-Useful for evaluating who the best shooter is, who the best defender is, the hot-hand hypothesis, etc.
-Scraped from NBA's REST API.  ##TO BE UPDATED FURTHER
+It has the following data:
+    1.  Match ID  
+    2.  Match date  
+    3.  Which team plays as home and away  
+    4.  Match result  
+    5.  Final score difference  
+    6.  Every shots attempt per game for each player  
+    7.  Shot result whether it is success or not  
+    8.  Type of shoot attempt (2 points or 3 points)  
+    9.  Quarter number data (1 to 4)  
+    10. Game clock for each quarter  
+    11. Time left for a shot (shot clock)  
+    12. Who took the shot  
+    13. Shot distance from basket  
+    14. Who was the nearest defender  
+    15. Distance from the nearest defender
 
-I   Strengths
+  The dataset is useful for evaluating who the best shooter is, who the best defender is, the hot-hand hypothesis, affect of home playing team, close game effect, shooting accuracy from best players, and defender distance effect on shooting accuracy.
+Data was taken from NBA's REST API provided by [Kaggle](https://www.kaggle.com/dansbecker/nba-shot-logs) and acquired at 30 September 2016.
+>>>>>>> ffa732d5e8e2f51ae1f136d3572f36becfcf2f39
 
-I   Limitations
+###Dataset Strengths
 
-###Analysis
+    The dataset comprise of comprehensive observational data to start to work with.
+In general, the dataset only have small percentage of missing value (0.19%) so we do not need to remove or assign many data points to default values.
+And since the column name is not ambiguous it really helps us to understand dataset from the beginning.
 
-## Installing libraries
+###Dataset Limitations  
+  There are several constraints we need to mentioned in order to provide same understandings for this group work analysis as follows:  
+    1. The dataset consists only for 1 year NBA season (2014-2015)  
+    2. The dataset comprise only with regular match (82 matches per team for both home and away matches)  
+    3. The playoff and other matches during the year is not included
+    4. The dataset does not include free throw shot and results
+
+##Analysis
+
+### Installing libraries
 
 ```{r}
-install.packages('reshape')
+#install.packages('reshape')
 
@@
 acc <- data.frame()
-acc[, "SECONDS_LEFT"] <- c(1:23)
 
@@
 
+acc[, "SECONDS_LEFT"] <- c(1:23)
+
 ggplot(acc, aes(x = SECONDS_LEFT, y = Accuracy)) + geom_point() + geom_smooth()
+
+df[, "ACC_P_SEC"] <- as.numeric()
+
+#Assign accuracy per second to main dataframe
+for (i in 1:nrow(df)){
+  for (j in 1:23){
+    if (df$SHOT_CLOCK[i] >j & df$SHOT_CLOCK[i] <= j+1 & df$SHOT_CLOCK[i] != 99) df[i, "ACC_P_SEC"] <- acc[j, "Accuracy"]
+  }
+}
+
+cov(df$ACC_P_SEC, df$SHOT_CLOCK)
+cor.test(df$ACC_P_SEC, df$SHOT_CLOCK)
 
@@
 
-game <- cast(data = df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + LOCATION + player_name + W ~ FGM)
+game <- cast(data = df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + LOCATION + player_name + W + FINAL_MARGIN~ FGM)
 
@@
 ```{r}
-
 # Pick out 8 of the best players in that regular season and create a new subset. Analyse how shooting percentage of these players can influence the result of a match.
@@
 super.star.list <- c("lebron james", "james harden", "russell westbrook", "stephen curry", "anthony davis", "blake griffin", "klay thompson", "lamarcus aldridge") # players picked out
-super.star <- game[game$player_name %in% super.star.list, c("player_name", "W", "FGA", "FG.percentage")] # create a subset for these players
-
-ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
-ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W))
+super.star <- game[game$player_name %in% super.star.list, c("player_name", "W", "LOCATION", "FINAL_MARGIN", "FGA", "FG.percentage")] # create a subset for these players
 
@@
 
+ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
+ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W))
+ggplot(super.star) + geom_point(aes(x = FINAL_MARGIN, y = FGA, colour = player_name))
+ggplot(super.star) + geom_boxplot(aes(x = LOCATION, y = FINAL_MARGIN))
+
+# t.test(super.star$FINAL_MARGIN[super.star$LOCATION == "H"], super.star$FINAL_MARGIN[super.star$LOCATION == "A"])
+
 # Correlation test
-t.test(super.star[, "W"], super.star[, "FGA"]) # this makes no sense...
+cor.test(super.star$FINAL_MARGIN, super.star$FGA)
+cor.test(super.star$FINAL_MARGIN, super.star$FG.percentage) # cor = 0.272
 
 # Linear regression
-summary(lm(W ~ player_name + FGA, data = super.star))
-summary(lm(W ~ player_name + FG.percentage, data = super.star))
+summary(lm(FINAL_MARGIN ~ FGA + LOCATION, data = super.star)) # Final margin ~ FGA and location
+summary(lm(FINAL_MARGIN ~ FG.percentage + LOCATION, data = super.star)) # Final margin ~ FG% and location
+summary(lm(FINAL_MARGIN ~ FG.percentage + FGA + LOCATION, data= super.star))
+
+summary(lm(FINAL_MARGIN ~ LOCATION, data = super.star)) # Final margin ~ FGA and location
 
@@
 
-###Defender Proximity Analysis
+##Defender Proximity Analysis  
+###Closest Defender Distance vs Shot Accuracy
+####Introduction  
+This analysis wants to calculate whether is there any significant effect from closest defender distance from a shooter with the shot outcome.
+The hypothesis is the further a defender from the shooter, then it will be more likely to increase the chance of scoring.
+This is because the shooter is practically in "free" condition without any physical interference.
+
+###Methods  
+The steps to calculate this measurement are as follows:  
+  1. Assign shot result to determine whether the shot is succeed (1) or missed (0)  
+  2. Select distinct of **all** closest defender distance  
+  3. Count the occurence fOr every distance  
+  4. Calculate how many succesfull shot scored from every distance point  
+  5. Calculate accuracy by dividing successfull shot scored with distance occurence for every distance point  
+  6. Plot the result
 ```{r}
+#Calculate accuracy against defender distance
+#Group the defender distance level
+distance_level <- as.data.frame(table(factor(ClosestDefenderDistance)))
+head(distance_level)
+
+#Calculate how many times players scored in every defender distance
+for (i in 1:nrow(distance_level)){
+  distance_level[i, "SCORED"] <- sum(df[df[, "CLOSE_DEF_DIST"] == distance_level[i, "Var1"], "FGM"])
+}
+
+#Calculate shooting accuracy for each distance
+distance_level$Accuracy <- round(distance_level$SCORED / distance_level$Freq, 4)
+
+#Plotting the result
+library(ggplot2)
+library(scales)
+ggplot(data = distance_level, aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
+  geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
+  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
+  scale_x_discrete("Closest Defender Distance", breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24)) +
+  ylab("Shooting Accuracy (%)") +
+  ggtitle("Closest Defender Distance vs Shooting Accuracy")
+```
+###Plot Result
+  The explanation from the plot above are as follows:   
+  1. If the defender stands really close to the shooter (0 to 2 feet), the accuracy varied between **42%** to **50%**  
+  2. If the defender distance range from 2 (two) to 4 (four) feet, the accuracy rate for this distance range from **44%** to **49%**  
+  3. If the defender distance range from 4 (four) to 6 (six) feet, the accuracy rate for this distance range from **41%** to **46%**
+  4. If the defender distance range from 6 (six) to 8 (eight) feet, the accuracy rate for this distance range from **37%** to **42%**
+  5. If the defender distance range from 8 (eight) to 10 (ten) feet, the accuracy rate for this distance range from **37%** to **54%**
+  6. For defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from **33%** to **60%**.  
+  7. For the distance greater than 16 feet, the data varies between **0%** (no shots scored from that distance) to **100%** (certain successfull shot from that distance)
+
+###Findings  
+  As the analysis shown, the accuracy rate of every player in the last season was somehow **unaffected** by the defender distance.
+For defender distance starting from 0 (zero) to For the distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance).
+It is also shown from the dataset that only small attempt of shots were made from further defender distance affecting the accuracy calculation result.
+This was arguably because further defender distance did not implied shooter distance from the basket is nearer to improve the accuracy.
+
+###Limitations  
+  As the dataset did not hold many data points for more than 20 feet defender distance, the accuracy analysis for these distances are limited.
+The analysis could be enhanced further by including additional season data.
+That way, there will be more data point for further distance and possibly would produce stronger pattern as the short distance shown.
+```{r}
+#Calculate accuracy against defender distance
+#Group the defender distance level
+
+distance_level <- as.data.frame(table(factor(ClosestDefenderDistance)))
+head(distance_level)
+
+#Calculate how many times players scored in every defender distance
+for (i in 1:nrow(distance_level)){
+  distance_level[i, "SCORED"] <- sum(df[df[, "CLOSE_DEF_DIST"] == distance_level[i, "Var1"], "FGM"])
+}
+
+#Calculate shooting accuracy for each distance
+distance_level$Accuracy <- round(distance_level$SCORED / distance_level$Freq, 4)
+
+#Plotting the result
+library(ggplot2)
+library(scales)
+ggplot(data = distance_level, aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
+  geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
+  scale_y_continuous(labels = percent) +
+  scale_x_discrete("Closest Defender Distance", breaks=c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24, Inf)) +
+  ylab("Shooting Accuracy (%)") +
+  ggtitle("Closest Defender Distance vs Shooting Accuracy")
+
+#----------------------------------------------------------------#
+
+#Calculate accuracy against shooting distance
+#Group the shooting distance level
+shooting_distance_level <- as.data.frame(table(factor(ShotDistance)))
+head(shooting_distance_level)
+
+#Calculate how many times players scored in every shooting distance
+for (i in 1:nrow(shooting_distance_level)){
+  shooting_distance_level[i, "SCORED"] <- sum(df[df[, "SHOT_DIST"] == shooting_distance_level[i, "Var1"], "FGM"])
+}
+
+#Calculate shooting accuracy for each distance
+shooting_distance_level$Accuracy <- round(shooting_distance_level$SCORED / shooting_distance_level$Freq, 2)
+
+#Plotting the result
+library(ggplot2)
+ggplot(data = shooting_distance_level, aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) +
+  geom_point(aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy))
+
 #Data Preparation
@@
 
+#Remove 0 from streak_per_player in order to show only the streaks achieved
+streak_per_player[[1]] <- streak_per_player[[1]][streak_per_player[[1]] != 0]
+for(i in 1:length(streak_per_player)) {
+  streak_per_player[[i]] <- streak_per_player[[i]][streak_per_player[[i]] != 0]
+}
+streak_per_player
+
 #Create a data frame for the strikes of each one of the 10 players (cannot create a data frame including all of them, because each list has different length)
@@
 IQR(player1_df$streaks) #Check that boxplot is valid
-#The typical length of a streak is 0 (median = 0) - The Interquartile Range is 1 - Streak lengths of 3 and 6 are unusually high compared to the rest of the distribution
+#The typical length of a streak is 1 (median = 1) - The Interquartile Range is 1 - Streak length of 6 is unusually high compared to the rest of the distribution
 boxplot(player2_df$streaks)
@@
 IQR(player2_df$streaks)
-#The typical length of a streak is 1 (median = 1) - The Interquartile Range is 2 - There are some high streak lengths (max = 4)
+#The typical length of a streak is 1.5 (median = 1.5) - The Interquartile Range is 1 - Streak length of 4 is unusually high compared to the rest of the distribution
 boxplot(player3_df$streaks)
@@
 IQR(player3_df$streaks)
-#The typical length of a streak is 0 (median = 0) - The Interquartile Range is 0 - Streak lengths of 3 and 13 (!!!) are unusually high compared to the rest of the distribution
+#The typical length of a streak is 8 (median = 8) - The Interquartile Range is 5 - There is a streak lengths of 13(!!!)
 boxplot(player4_df$streaks)
@@
 IQR(player4_df$streaks)
-#The typical length of a streak is 1 (median = 1) - The Interquartile Range is 2 - Streak length of 6 is unusually high compared to the rest of the distribution
+#The typical length of a streak is 2 (median = 2) - The Interquartile Range is 1.5 - Streak length of 6 is unusually high compared to the rest of the distribution
 boxplot(player5_df$streaks)
@@
 IQR(player5_df$streaks)
-#The typical length of a streak is 1 (median = 1) - The Interquartile Range is 2 - There are some high streak lengths (max = 5)
+#The typical length of a streak is 2 (median = 2) - The Interquartile Range is 2 - There are some high streak lengths (max = 5)
 
@@
 So, if consecutive shots are independent, we disprove hot hand theory, while if consecutive shots are dependent, we prove the hot hand theory.
-#Simulation - create a vector of shots('made', 'missed') for a sample shooter
-set.seed(23)
+
+#Find the shooting percentage for each of the 5 players
+j <- 1
+shooting_percentage <- list()
+for(i in best_players_df$player_id) {
+  shooting_percentage[j] <- sum(df[df[ ,'player_id'] == i, 'FGM'])/ nrow(df[df[ ,'player_id'] == i, ])
+  j <- j+1
+}
+shooting_percentage
+
+###PLAYER 1 VS SAMPLE PLAYER###
+#Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
+set.seed(21)
 shot_results <- c('made', 'missed')
-simul_shooter <- sample(shot_results, size = nrow(player1_df), replace = TRUE, prob = c(0.45, 0.55))
+simul_shooter <- sample(shot_results, size = length(streak_per_player[[1]]), replace = TRUE, prob = c(shooting_percentage[[1]], 1 - shooting_percentage[[1]]))
 simul_shooter
@@
 streaks_of_simulshooter <- number_of_streaks(simul_shooter)
-simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
+
+#Remove zeros
+true_streaks_of_simulshooter <- list()
+j <- 1
+for(i in 1:length(streaks_of_simulshooter)) {
+  if(streaks_of_simulshooter[[i]] != 0){
+    true_streaks_of_simulshooter[j] <- streaks_of_simulshooter[[i]]
+    j <- j + 1
+  }
+}
+true_streaks_of_simulshooter
+true_streaks_of_simulshooter <- unlist(true_streaks_of_simulshooter)
+
+#Create a data frame 
+simulshooter_df <- data.frame('streaks' = true_streaks_of_simulshooter)
+simulshooter_df
 
 #Comparisons of data between sample shooter and real shooters
-summary(streaks_of_simulshooter)
+summary(simulshooter_df$streaks)
 summary(player1_df$streaks)
@@
 ggplot(data = player1_df, aes(x = streaks)) + geom_bar()
-#
+#We can see that streak length distributions are quite similar, so we can conclude that Kyrie Irving doesn't have a hot hand.
+
+###PLAYER 2 VS SAMPLE PLAYER###
+#Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter2 
+set.seed(154)
+shot_results <- c('made', 'missed')
+simul_shooter <- sample(shot_results, size = length(streak_per_player[[2]]), replace = TRUE, prob = c(shooting_percentage[[2]], 1 - shooting_percentage[[2]]))
+simul_shooter
+
+#Set 1 for baskets made and 0 for baskets missed
+simul_shooter <- as.list(simul_shooter)
+for(i in 1:length(simul_shooter)) {
+   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
+   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0
+}
+simul_shooter
+
+#Use previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
+streaks_of_simulshooter <- number_of_streaks(simul_shooter)
+
+#Remove zeros
+true_streaks_of_simulshooter <- list()
+j <- 1
+for(i in 1:length(streaks_of_simulshooter)) {
+  if(streaks_of_simulshooter[[i]] != 0){
+    true_streaks_of_simulshooter[j] <- streaks_of_simulshooter[[i]]
+    j <- j + 1
+  }
+}
+true_streaks_of_simulshooter
+true_streaks_of_simulshooter <- unlist(true_streaks_of_simulshooter)
+
+#Create a data frame 
+simulshooter_df <- data.frame('streaks' = true_streaks_of_simulshooter)
+
+#Comparisons of data between sample shooter and real shooters
+summary(simulshooter_df$streaks)
+summary(player2_df$streaks)
+
+#Compare barplots of sample shooter and real shooter1
+#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
+ggplot(data = simulshooter_df, aes(x = streaks)) + geom_bar() 
+ggplot(data = player2_df, aes(x = streaks)) + geom_bar()
+#We can see that the sample shooter with the same shooting percentage can have extremely high streak lengths, which means that Mo Williams does not have a hot hand.
+
+###PLAYER 3###
+
+best_players_df
+#logistic regression
+
 ```
@@
 
-###Discussion
-I   What results did you find?
+##Discussion  
+###What results did you find?
 
-I   Why is this interesting?
+###Why is this interesting?
 
-I   What would be the next steps?
+###What would be the next steps?
 ---
@@
 ```{r setup, include=FALSE}
+
 knitr::opts_chunk$set(echo = TRUE)
 ```
-Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Widodo
+Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo
 

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
Updating text for home/away adv section

@@
 
-df$BLANK<-"" #for some reason cast function does not work with last column in dataframe, so I have created this BLANK column as I need to use AWAY_TEAM colum (it was a last column). 
 ```
@@
 round(hh/(hh+aw),2)
+round(aw/(hh+aw),2)
+
 ```
@@
 
-aa3$TEAM <- factor(aa5$TEAM, levels = aa5$TEAM[order(-aa5$WIN_PCT_HOME)])
+aa3$TEAM <- factor(aa3$TEAM, levels = aa3$TEAM[order(-aa3$WIN_PCT_HOME)])
 
 ggplot(aa3,aes(x=TEAM)) + geom_point(aes(y= WIN_PCT_HOME, col= "HOME"))+ geom_point(aes( y= WIN_PCT_AWAY, col= "AWAY"))
-#TO BE UPDATED  
+
 ```

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd

@@
 #install.packages('reshape')
-
+#install.packages('gridExtra')
 ```
@@
 library(ggplot2)
-
+library(gridExtra)
 ```
@@
 
-#Set a new data frame for the 5 players with best performances(higher points)
-best_players_df <- head(pttsum2,5)
+#Set a new data frame for the 3 players with best performances(higher points)
+best_players_df <- head(pttsum2,3)
 
@@
 
-#Create a data frame for the strikes of each one of the 10 players (cannot create a data frame including all of them, because each list has different length)
+#Create a data frame for the strikes of each one of the 3 players (cannot create a data frame including all of them, because each list has different length)
 player1_df <- data.frame('streaks' = streak_per_player[[1]])
@@
 player3_df <- data.frame('streaks' = streak_per_player[[3]])
-player4_df <- data.frame('streaks' = streak_per_player[[4]])
-player5_df <- data.frame('streaks' = streak_per_player[[5]])
 
 #Show distribution for each player's streak lengths
-ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[1,'player_name'])
-ggplot(data = player2_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[2,'player_name'])
-ggplot(data = player3_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[3,'player_name'])
-ggplot(data = player4_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[4,'player_name'])
-ggplot(data = player5_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[5,'player_name'])
+gg1 <- ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[1,'player_name'])
+gg2 <- ggplot(data = player2_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[2,'player_name'])
+gg3 <- ggplot(data = player3_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[3,'player_name'])
+grid.arrange(gg1, gg2, gg3, ncol = 3)
 
@@
 #The typical length of a streak is 8 (median = 8) - The Interquartile Range is 5 - There is a streak lengths of 13(!!!)
-boxplot(player4_df$streaks)
-summary(player4_df$streaks)
-IQR(player4_df$streaks)
-#The typical length of a streak is 2 (median = 2) - The Interquartile Range is 1.5 - Streak length of 6 is unusually high compared to the rest of the distribution
-boxplot(player5_df$streaks)
-summary(player5_df$streaks)
-IQR(player5_df$streaks)
-#The typical length of a streak is 2 (median = 2) - The Interquartile Range is 2 - There are some high streak lengths (max = 5)
 
@@
 
-#Find the shooting percentage for each of the 5 players
+#Find the shooting percentage for each of the 3 players
 j <- 1
@@
 #Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
-set.seed(21)
+set.seed(974)
 shot_results <- c('made', 'missed')
@@
 
-#Remove zeros
-true_streaks_of_simulshooter <- list()
-j <- 1
-for(i in 1:length(streaks_of_simulshooter)) {
-  if(streaks_of_simulshooter[[i]] != 0){
-    true_streaks_of_simulshooter[j] <- streaks_of_simulshooter[[i]]
-    j <- j + 1
-  }
-}
-true_streaks_of_simulshooter
-true_streaks_of_simulshooter <- unlist(true_streaks_of_simulshooter)
-
-#Create a data frame 
-simulshooter_df <- data.frame('streaks' = true_streaks_of_simulshooter)
-simulshooter_df
+#Create new data frames for the graphs 
+simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
+realshooter_df <- data.frame('streaks' = streak_per_player[[1]])
 
@@
 summary(simulshooter_df$streaks)
-summary(player1_df$streaks)
+summary(realshooter_df$streaks)
 
@@
 #If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
-ggplot(data = simulshooter_df, aes(x = streaks)) + geom_bar() 
-ggplot(data = player1_df, aes(x = streaks)) + geom_bar()
+g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar() + labs(title = 'sample shooter')
+g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar() + labs(title = 'Kyrie Irving')
+grid.arrange(g1, g2, ncol = 2)
 #We can see that streak length distributions are quite similar, so we can conclude that Kyrie Irving doesn't have a hot hand.
@@
 ###PLAYER 2 VS SAMPLE PLAYER###
-#Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter2 
-set.seed(154)
+#Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
+set.seed(572)
 shot_results <- c('made', 'missed')
@@
 
-#Remove zeros
-true_streaks_of_simulshooter <- list()
-j <- 1
-for(i in 1:length(streaks_of_simulshooter)) {
-  if(streaks_of_simulshooter[[i]] != 0){
-    true_streaks_of_simulshooter[j] <- streaks_of_simulshooter[[i]]
-    j <- j + 1
-  }
-}
-true_streaks_of_simulshooter
-true_streaks_of_simulshooter <- unlist(true_streaks_of_simulshooter)
-
-#Create a data frame 
-simulshooter_df <- data.frame('streaks' = true_streaks_of_simulshooter)
+#Create new data frames for the graphs 
+simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
+realshooter_df <- data.frame('streaks' = streak_per_player[[2]])
 
@@
 summary(simulshooter_df$streaks)
-summary(player2_df$streaks)
+summary(realshooter_df$streaks)
 
@@
 #If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
-ggplot(data = simulshooter_df, aes(x = streaks)) + geom_bar() 
-ggplot(data = player2_df, aes(x = streaks)) + geom_bar()
-#We can see that the sample shooter with the same shooting percentage can have extremely high streak lengths, which means that Mo Williams does not have a hot hand.
+g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar() + labs(title = 'sample shooter')
+g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar() + labs(title = 'Mo Williams')
+grid.arrange(g1, g2, ncol = 2)
 
-###PLAYER 3###
+###PLAYER 3 VS SAMPLE PLAYER###
+#Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
+set.seed(57)
+shot_results <- c('made', 'missed')
+simul_shooter <- sample(shot_results, size = length(streak_per_player[[3]]), replace = TRUE, prob = c(shooting_percentage[[3]], 1 - shooting_percentage[[3]]))
+simul_shooter
+
+#Set 1 for baskets made and 0 for baskets missed
+simul_shooter <- as.list(simul_shooter)
+for(i in 1:length(simul_shooter)) {
+   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
+   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0
+}
+simul_shooter
+
+#Use previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
+streaks_of_simulshooter <- number_of_streaks(simul_shooter)
+
+#Create new data frames for the graphs 
+simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
+realshooter_df <- data.frame('streaks' = streak_per_player[[3]])
+
+#Comparisons of data between sample shooter and real shooters
+summary(simulshooter_df$streaks)
+summary(realshooter_df$streaks)
+
+#Compare barplots of sample shooter and real shooter1
+#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
+g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar() + labs(title = 'sample shooter')
+g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar() + labs(title = 'Klay Thompson')
+grid.arrange(g1, g2, ncol = 2)
+#Klay Thompson is an exception, had 13 streaks in a row! We cannot find a similar distribution in our sample, so we can assume that he is likely to have a hot hand.
 
 best_players_df
-#logistic regression
 

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
additional explanation for defender proximity

--- 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
@@
 ####Introduction  
-This analysis wants to calculate whether is there any significant effect from closest defender distance from a shooter with the shot outcome.
+This analysis wants to calculate whether there was any significant effect from closest defender distance from a shooter with the shot outcome.
 The hypothesis is the further a defender from the shooter, then it will be more likely to increase the chance of scoring.
@@
   ggtitle("Closest Defender Distance vs Shooting Accuracy")
+
 ```
@@
 
-###Findings  
-  As the analysis shown, the accuracy rate of every player in the last season was somehow **unaffected** by the defender distance.
-For defender distance starting from 0 (zero) to For the distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance).
+  Next analysis will be calculated from correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:
+```{r}
+distance_level$Var1 <- as.numeric(distance_level$Var1)
+distance_level$Var1 <- distance_level$Var1 * 0.1 - 0.1
+
+distanceZerotoAll <- c()
+accuracyDist1 <- c()
+for (i in 0:nrow(distance_level)){
+  distanceZerotoAll <- c(distanceZerotoAll, distance_level[i, "Var1"])
+  accuracyDist1 <- c(accuracyDist1, distance_level[i, "Accuracy"])
+}
+
+dfZerotoAll <- data.frame(distanceZerotoAll, accuracyDist1)
+cor.test(dfZerotoFive$distanceZerotoAll, dfZerotoFive$accuracyDist1)
+```
+###Correlation Test Result  
+  The correlation test result produced slightly high number of correlation between defender distance with shooting accuracy for all shot distance attempt in the last seasons's NBA (0.59).
+If the findings made only based on this number only, there would be obvious that further defender distance will **automatically** increase shooting accuracy.
+However, after the analysis went into smaller piece of distance, ther result will be difference.
+For example, for distance range between 0 (zero) to 8 feet, the correlation measurement will be as follows:
+
+```{r}
+distanceZerotoEight <- c()
+accuracyDist2 <- c()
+for (i in 0:80){
+  distanceZerotoEight <- c(distanceZerotoEight, distance_level[i, "Var1"])
+  accuracyDist2 <- c(accuracyDist2, distance_level[i, "Accuracy"])
+}
+
+dfZerotoEight <- data.frame(distanceZerotoEight, accuracyDist2)
+cor.test(dfZerotoEight$distanceZerotoEight, dfZerotoEight$accuracyDist2)
+```
+  Therefore,
+
+###Discussion  
+  As the analysis shown, the accuracy rate of every player in the last season was somehow **affected** by the defender distance for certain range.
+For defender distance starting from 0 (zero) to 8 (eight) feet For the distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance).
 It is also shown from the dataset that only small attempt of shots were made from further defender distance affecting the accuracy calculation result.
@@
 That way, there will be more data point for further distance and possibly would produce stronger pattern as the short distance shown.
+
+###Shooting Distance vs Shot Accuracy
+####Introduction  
+This analysis wants to calculate whether is there any significant effect from closest defender distance from a shooter with the shot outcome.
+The hypothesis is the further a defender from the shooter, then it will be more likely to increase the chance of scoring.
+This is because the shooter is practically in "free" condition without any physical interference.
+
+###Methods  
+The steps to calculate this measurement are as follows:  
+  1. Assign shot result to determine whether the shot is succeed (1) or missed (0)  
+  2. Select distinct of **all** closest defender distance  
+  3. Count the occurence fOr every distance  
+  4. Calculate how many succesfull shot scored from every distance point  
+  5. Calculate accuracy by dividing successfull shot scored with distance occurence for every distance point  
+  6. Plot the result
 ```{r}
-#Calculate accuracy against defender distance
-#Group the defender distance level
-
-distance_level <- as.data.frame(table(factor(ClosestDefenderDistance)))
-head(distance_level)
-
-#Calculate how many times players scored in every defender distance
-for (i in 1:nrow(distance_level)){
-  distance_level[i, "SCORED"] <- sum(df[df[, "CLOSE_DEF_DIST"] == distance_level[i, "Var1"], "FGM"])
-}
-
-#Calculate shooting accuracy for each distance
-distance_level$Accuracy <- round(distance_level$SCORED / distance_level$Freq, 4)
-
-#Plotting the result
-library(ggplot2)
-library(scales)
-ggplot(data = distance_level, aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
-  geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
-  scale_y_continuous(labels = percent) +
-  scale_x_discrete("Closest Defender Distance", breaks=c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24, Inf)) +
-  ylab("Shooting Accuracy (%)") +
-  ggtitle("Closest Defender Distance vs Shooting Accuracy")
-
-#----------------------------------------------------------------#
-
 #Calculate accuracy against shooting distance

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
more code-hot hand

--- 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
@@
 set.seed(57)
-shot_results <- c('made', 'missed')
+shot_results <- c('made', 'missed')javascript:;
 simul_shooter <- sample(shot_results, size = length(streak_per_player[[3]]), replace = TRUE, prob = c(shooting_percentage[[3]], 1 - shooting_percentage[[3]]))

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
text

@@
 set.seed(57)
-shot_results <- c('made', 'missed')javascript:;
+shot_results <- c('made', 'missed')
 simul_shooter <- sample(shot_results, size = length(streak_per_player[[3]]), replace = TRUE, prob = c(shooting_percentage[[3]], 1 - shooting_percentage[[3]]))
@@
 grid.arrange(g1, g2, ncol = 2)
-#Klay Thompson is an exception, had 13 streaks in a row! We cannot find a similar distribution in our sample, so we can assume that he is likely to have a hot hand.
-
-best_players_df
-
+#Klay Thompson is an exception, had 13 streaks in a row!
+We cannot find a similar distribution in our sample, so we can assume that he is likely to have a hot hand.
+Fearther statistical analysis (logistic regression) should be done in order to reach a conclusion.
 ```

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
working on rest days analysis

--- 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
@@
 #install.packages('reshape')
-#install.packages('gridExtra')
+
 ```
@@
 library(ggplot2)
-library(gridExtra)
+
 ```
@@
 df1<- read.csv("shot_logs1010.csv")
-
-#we need to change the format of some of the columns to 'Character'
-#get rid off 'attach()'
-#get rid off column X in shot logs1010 file
-
-```
-
-### Home Advantage analysis
 
@@
 
+#adding a column 'DATE'
+df1[,"MATCHUP"] <- as.character(df1[,"MATCHUP"])
+for (i in 1:nrow(df1)) df1$DATE[i] <- as.character(as.Date(unlist(strsplit(df1$MATCHUP[i]," -"))[1], format= "%B %d, %Y"))
+
+restdays<-cast(df1, DATE+ HOME_TEAM + AWAY_TEAM + WINNER ~ FGM)
 ```
+
+#we need to change the format of some of the columns to 'Character'
+#get rid off 'attach()'
+#get rid off column X in shot logs1010 file
+
+```
+
+## Home Advantage analysis
+###Introduction
+
+This area of anlysis focuses on investigation of whether playing at 'home' gives the team advantage comparing to playing 'away'.
+
+Literature review demonstrates that
+
+The NULL Hypothesis is:
 
@@
 
-#Set a new data frame for the 3 players with best performances(higher points)
-best_players_df <- head(pttsum2,3)
+#Set a new data frame for the 5 players with best performances(higher points)
+best_players_df <- head(pttsum2,5)
 
@@
 
-#Create a data frame for the strikes of each one of the 3 players (cannot create a data frame including all of them, because each list has different length)
+#Create a data frame for the strikes of each one of the 10 players (cannot create a data frame including all of them, because each list has different length)
 player1_df <- data.frame('streaks' = streak_per_player[[1]])
@@
 player3_df <- data.frame('streaks' = streak_per_player[[3]])
+player4_df <- data.frame('streaks' = streak_per_player[[4]])
+player5_df <- data.frame('streaks' = streak_per_player[[5]])
 
 #Show distribution for each player's streak lengths
-gg1 <- ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[1,'player_name'])
-gg2 <- ggplot(data = player2_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[2,'player_name'])
-gg3 <- ggplot(data = player3_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[3,'player_name'])
-grid.arrange(gg1, gg2, gg3, ncol = 3)
+ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[1,'player_name'])
+ggplot(data = player2_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[2,'player_name'])
+ggplot(data = player3_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[3,'player_name'])
+ggplot(data = player4_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[4,'player_name'])
+ggplot(data = player5_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[5,'player_name'])
 
@@
 #The typical length of a streak is 8 (median = 8) - The Interquartile Range is 5 - There is a streak lengths of 13(!!!)
+boxplot(player4_df$streaks)
+summary(player4_df$streaks)
+IQR(player4_df$streaks)
+#The typical length of a streak is 2 (median = 2) - The Interquartile Range is 1.5 - Streak length of 6 is unusually high compared to the rest of the distribution
+boxplot(player5_df$streaks)
+summary(player5_df$streaks)
+IQR(player5_df$streaks)
+#The typical length of a streak is 2 (median = 2) - The Interquartile Range is 2 - There are some high streak lengths (max = 5)
 
@@
 
-#Find the shooting percentage for each of the 3 players
+#Find the shooting percentage for each of the 5 players
 j <- 1
@@
 #Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
-set.seed(974)
+set.seed(21)
 shot_results <- c('made', 'missed')
@@
 
-#Create new data frames for the graphs 
-simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
-realshooter_df <- data.frame('streaks' = streak_per_player[[1]])
+#Remove zeros
+true_streaks_of_simulshooter <- list()
+j <- 1
+for(i in 1:length(streaks_of_simulshooter)) {
+  if(streaks_of_simulshooter[[i]] != 0){
+    true_streaks_of_simulshooter[j] <- streaks_of_simulshooter[[i]]
+    j <- j + 1
+  }
+}
+true_streaks_of_simulshooter
+true_streaks_of_simulshooter <- unlist(true_streaks_of_simulshooter)
+
+#Create a data frame 
+simulshooter_df <- data.frame('streaks' = true_streaks_of_simulshooter)
+simulshooter_df
 
@@
 summary(simulshooter_df$streaks)
-summary(realshooter_df$streaks)
+summary(player1_df$streaks)
 
@@
 #If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
-g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar() + labs(title = 'sample shooter')
-g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar() + labs(title = 'Kyrie Irving')
-grid.arrange(g1, g2, ncol = 2)
+ggplot(data = simulshooter_df, aes(x = streaks)) + geom_bar() 
+ggplot(data = player1_df, aes(x = streaks)) + geom_bar()
 #We can see that streak length distributions are quite similar, so we can conclude that Kyrie Irving doesn't have a hot hand.
@@
 ###PLAYER 2 VS SAMPLE PLAYER###
-#Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
-set.seed(572)
+#Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter2 
+set.seed(154)
 shot_results <- c('made', 'missed')
@@
 
-#Create new data frames for the graphs 
-simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
-realshooter_df <- data.frame('streaks' = streak_per_player[[2]])
+#Remove zeros
+true_streaks_of_simulshooter <- list()
+j <- 1
+for(i in 1:length(streaks_of_simulshooter)) {
+  if(streaks_of_simulshooter[[i]] != 0){
+    true_streaks_of_simulshooter[j] <- streaks_of_simulshooter[[i]]
+    j <- j + 1
+  }
+}
+true_streaks_of_simulshooter
+true_streaks_of_simulshooter <- unlist(true_streaks_of_simulshooter)
+
+#Create a data frame 
+simulshooter_df <- data.frame('streaks' = true_streaks_of_simulshooter)
 
@@
 summary(simulshooter_df$streaks)
-summary(realshooter_df$streaks)
+summary(player2_df$streaks)
 
@@
 #If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
-g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar() + labs(title = 'sample shooter')
-g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar() + labs(title = 'Mo Williams')
-grid.arrange(g1, g2, ncol = 2)
+ggplot(data = simulshooter_df, aes(x = streaks)) + geom_bar() 
+ggplot(data = player2_df, aes(x = streaks)) + geom_bar()
+#We can see that the sample shooter with the same shooting percentage can have extremely high streak lengths, which means that Mo Williams does not have a hot hand.
 
-###PLAYER 3 VS SAMPLE PLAYER###
-#Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
-set.seed(57)
-shot_results <- c('made', 'missed')javascript:;
-simul_shooter <- sample(shot_results, size = length(streak_per_player[[3]]), replace = TRUE, prob = c(shooting_percentage[[3]], 1 - shooting_percentage[[3]]))
-simul_shooter
-
-#Set 1 for baskets made and 0 for baskets missed
-simul_shooter <- as.list(simul_shooter)
-for(i in 1:length(simul_shooter)) {
-   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
-   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0
-}
-simul_shooter
-
-#Use previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
-streaks_of_simulshooter <- number_of_streaks(simul_shooter)
-
-#Create new data frames for the graphs 
-simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
-realshooter_df <- data.frame('streaks' = streak_per_player[[3]])
-
-#Comparisons of data between sample shooter and real shooters
-summary(simulshooter_df$streaks)
-summary(realshooter_df$streaks)
-
-#Compare barplots of sample shooter and real shooter1
-#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
-g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar() + labs(title = 'sample shooter')
-g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar() + labs(title = 'Klay Thompson')
-grid.arrange(g1, g2, ncol = 2)
-#Klay Thompson is an exception, had 13 streaks in a row! We cannot find a similar distribution in our sample, so we can assume that he is likely to have a hot hand.
+###PLAYER 3###
 
 best_players_df
+#logistic regression
 

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd

--- 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
@@
 
-#we need to change the format of some of the columns to 'Character'
-#get rid off 'attach()'
-#get rid off column X in shot logs1010 file
-
-```
-
-### Home Advantage analysis
-
 ```{r}

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
add analysis for shooting distance accuracy

@@
 
+#we need to change the format of some of the columns to 'Character'
+#get rid off 'attach()'
+#get rid off column X in shot logs1010 file
+
+```
+
+### Home Advantage analysis
+
 ```{r}
@@
 
-#adding a column 'DATE'
-df1[,"MATCHUP"] <- as.character(df1[,"MATCHUP"])
-for (i in 1:nrow(df1)) df1$DATE[i] <- as.character(as.Date(unlist(strsplit(df1$MATCHUP[i]," -"))[1], format= "%B %d, %Y"))
-
-restdays<-cast(df1, DATE+ HOME_TEAM + AWAY_TEAM + WINNER ~ FGM)
 ```
-
-#we need to change the format of some of the columns to 'Character'
-#get rid off 'attach()'
-#get rid off column X in shot logs1010 file
-
-```
-
-## Home Advantage analysis
-###Introduction
-
-This area of anlysis focuses on investigation of whether playing at 'home' gives the team advantage comparing to playing 'away'.
-
-Literature review demonstrates that
-
-The NULL Hypothesis is:
 
@@
 ####Introduction  
-This analysis wants to calculate whether there was any significant effect from closest defender distance from a shooter with the shot outcome.
+This analysis would calculate whether there was any significant effect from closest defender distance from a shooter with the shot outcome.
 The hypothesis is the further a defender from the shooter, then it will be more likely to increase the chance of scoring.
@@
   4. Calculate how many succesfull shot scored from every distance point  
-  5. Calculate accuracy by dividing successfull shot scored with distance occurence for every distance point  
+  5. Calculate accuracy by dividing successfull shot scored with distance occurence for every defender distance point  
   6. Plot the result
@@
   scale_x_discrete("Closest Defender Distance", breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24)) +
+  geom_smooth() +
   ylab("Shooting Accuracy (%)") +
@@
   1. If the defender stands really close to the shooter (0 to 2 feet), the accuracy varied between **42%** to **50%**  
-  2. If the defender distance range from 2 (two) to 4 (four) feet, the accuracy rate for this distance range from **44%** to **49%**  
-  3. If the defender distance range from 4 (four) to 6 (six) feet, the accuracy rate for this distance range from **41%** to **46%**
-  4. If the defender distance range from 6 (six) to 8 (eight) feet, the accuracy rate for this distance range from **37%** to **42%**
-  5. If the defender distance range from 8 (eight) to 10 (ten) feet, the accuracy rate for this distance range from **37%** to **54%**
+  2. If the defender distance range from 2 to 4 feet, the accuracy rate for this distance range from **44%** to **49%**  
+  3. If the defender distance range from 4 to 6 feet, the accuracy rate for this distance range from **41%** to **46%**
+  4. If the defender distance range from 6 to 8 feet, the accuracy rate for this distance range from **37%** to **42%**
+  5. If the defender distance range from 8 to 10 feet, the accuracy rate for this distance range from **37%** to **54%**
   6. For defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from **33%** to **60%**.  
@@
 ```{r}
+#change distance level to numeric type for correlation test function
 distance_level$Var1 <- as.numeric(distance_level$Var1)
-distance_level$Var1 <- distance_level$Var1 * 0.1 - 0.1
+distance_level$Var1 <- distance_level$Var1 * 0.1 - 0.1 #to keep the distance point as it is
 
+#populate data points from data frame
 distanceZerotoAll <- c()
@@
 
+#combine data frame
 dfZerotoAll <- data.frame(distanceZerotoAll, accuracyDist1)
+
+#calculate correlation
 cor.test(dfZerotoFive$distanceZerotoAll, dfZerotoFive$accuracyDist1)
@@
 ```{r}
+#populate data from data frame
 distanceZerotoEight <- c()
@@
 
+#combine data frame
 dfZerotoEight <- data.frame(distanceZerotoEight, accuracyDist2)
+
+#calculate correlation
 cor.test(dfZerotoEight$distanceZerotoEight, dfZerotoEight$accuracyDist2)
 ```
-  Therefore,
+  As the analysis shown, the correlation measurement between 0 (zero) to 8 feet would be negative (-0.40). This result was different with the previous calculation done for all data points (0.59).
 
@@
 It is also shown from the dataset that only small attempt of shots were made from further defender distance affecting the accuracy calculation result.
-This was arguably because further defender distance did not implied shooter distance from the basket is nearer to improve the accuracy.
+However, from the correlation perspective, the result would be different if the analysis included all of the data points versus only included short distance data points.
+This was arguably because further defender distance did not implied shooter distance from the basket is nearer to improve the accuracy and also there were not many players to perform shooting attempt automatically although their opponent was standing far from the shooter
 
@@
 ####Introduction  
-This analysis wants to calculate whether is there any significant effect from closest defender distance from a shooter with the shot outcome.
-The hypothesis is the further a defender from the shooter, then it will be more likely to increase the chance of scoring.
-This is because the shooter is practically in "free" condition without any physical interference.
+This analysis would calculate whether was there any significant effect from shooting distance from basket with the shot outcome.
+The hypothesis is the further a shooter from the basket, then the success rate would be decreased.
+This was because when the shooter was practically in closer shooting range, the shooter shooting's result would be more accurate.
 
@@
   1. Assign shot result to determine whether the shot is succeed (1) or missed (0)  
-  2. Select distinct of **all** closest defender distance  
+  2. Select distinct of **all** shooting distance  
   3. Count the occurence fOr every distance  
   4. Calculate how many succesfull shot scored from every distance point  
-  5. Calculate accuracy by dividing successfull shot scored with distance occurence for every distance point  
+  5. Calculate accuracy by dividing successfull shot scored with particular distance occurence for every shooting distance point  
   6. Plot the result
@@
 #Calculate shooting accuracy for each distance
-shooting_distance_level$Accuracy <- round(shooting_distance_level$SCORED / shooting_distance_level$Freq, 2)
+shooting_distance_level$Accuracy <- round(shooting_distance_level$SCORED / shooting_distance_level$Freq, 4)
 
@@
 library(ggplot2)
+library(scales)
 ggplot(data = shooting_distance_level, aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) +
-  geom_point(aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy))
-
-#Data Preparation
-#Change the value of shot result column to numeric
-df[, "SHOT_RESULT"] <- as.character(df[, "SHOT_RESULT"])
-for (j in 1:nrow(df)){
-  df[j, "SHOT_RESULT"] <- ifelse (df[j, "SHOT_RESULT"] == "made", 1, 0)
-}
-
-#set the value to numeric or integer accordingly
-df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
-df[, "PTS_TYPE"] <- as.integer(df[, "PTS_TYPE"])
-df[, "CLOSE_DEF_DIS"] <- as.numeric(df[, "CLOSE_DEF_DIST"])
-
-#select data with only scored result
-#Point type (2 or 3 points) and distance
-point_type <- c()
-distance <- c()
-for (z in 1:nrow(df)){
-  if(df[z, "SHOT_RESULT"] == 1 & df[z, "PTS_TYPE"] >= 1){
-    point_type <- c(point_type, df[z, "PTS_TYPE"])
-    distance <- c(distance, df[z, "CLOSE_DEF_DIST"])
-  }
-}
-
-#combine point type and distance
-valid_distance = data.frame(point_type, distance)
-
-#calculate correlation between defender distance and points made (2 or 3 points)
-cor(point_type, distance)
-
-#plot the diagram for this correlation
-library(ggplot2)
-ggplot(data = valid_distance, aes(x = point_type, y = distance, group = point_type)) + 
-  geom_boxplot() + stat_summary(fun.y = mean, geom = "point", shape = 5, size = 1) + 
-  ggtitle("Points Scored based on Distance from Defender")
-
-#calculate correlation between defender distance and shot result
-cor(df[, "SHOT_RESULT"], df[, "CLOSE_DEF_DIS"])
-
-#--------------------------------------------------------------------------------#
-
-#Further analysis
-#select distance data from 2 points shot scored and the players who scored it
-distance2 <- c()
-players2 <- c()
-for (z in 1:nrow(df)){
-  if(df[z, "SHOT_RESULT"] == 1 & df[z, "PTS_TYPE"] == 2){
-    distance2 <- c(distance2, df[z, "CLOSE_DEF_DIST"])
-    players2 <- c(players2, as.character(df[z, "player_name"]))
-  }
-}
-
-#combine 2 point players and defender distance
-player2point <- data.frame(players2, distance2)
-summary(player2point)
-
-#count players who scored 2 points
-counter <- as.data.frame(table(players2))
-
-#count 5 top players
-top5Players2 <- counter[rev(order(counter$Freq)), "players2"][1:5]
-top5Counter <- counter[rev(order(counter$Freq)), "Freq"][1:5]
-
-top5_2points <- data.frame(top5players2, top5Counter)
-
-#plot the graph
-library(ggplot2)
-ggplot(data = top5_2points, aes(x = top5Players2)) +
-  geom_bar(aes(x = top5players2, y = top5Counter), stat = "identity")
-
-#----------------------------------------------------------------#
-
-#select distance data from 3 points shot scored and the players who scored it
-distance3 <- c()
-players3 <- c()
-for (z in 1:nrow(df)){
-  if(df[z, "SHOT_RESULT"] == 1 & df[z, "PTS_TYPE"] == 3){
-    distance3 <- c(distance3, df[z, "CLOSE_DEF_DIST"])
-    players3 <- c(players3, as.character(df[z, "player_name"]))
-  }
-}
-
-#combine 3 point players and defender distance
-player3point <- data.frame(players3, distance3)
-summary(player3point)
-
-#count players who scored 3 points
-counter3 <- as.data.frame(table(players3))
-
-#count 5 top players
-top5Players3 <- counter3[rev(order(counter3$Freq)), "players3"][1:5]
-top5Counter3 <- counter3[rev(order(counter3$Freq)), "Freq"][1:5]
-
-top5_3points <- data.frame(top5Players3, top5Counter3)
-
-#plot the graph
-library(ggplot2)
-ggplot(data = top5_3points, aes(x = top5Players3)) +
-  geom_bar(aes(x = top5Players3, y = top5Counter3), stat = "identity")
+  geom_point(aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) +
+  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
+  scale_x_discrete("Shooting Distance", breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 32, 34, 36, 38, 40, 42, 44, 46, 48)) +
+  geom_smooth() +
+  ylab("Shooting Accuracy (%)") +
+  ggtitle("Shooting Distance vs Shooting Accuracy")
 
@@
 
+###Plot Result
+  The explanation from the plot above are as follows:   
+  1.  If the shooter stands really close to the basket (0 to 2 feet), the accuracy varied between **62%** to **68%**  
+  2.  If the shooting distance range from 2 to 4 feet, the accuracy rate for this distance range from **61%** to **67%**  
+  3.  If the shooting distance range from 4 to 6 feet, the accuracy rate for this distance range from **44%** to **58%**  
+  4.  If the shooting distance range from 6 to 8 feet, the accuracy rate for this distance range from **38%** to **51%**  
+  5.  If the shooting distance range from 8 to 10 feet, the accuracy rate for this distance range from **35%** to **42%**  
+  6.  If the shooting distance range from 10 to 12 feet, the accuracy rate for this distance range from **34%** to **42%**  
+  7.  If the shooting distance range from 12 to 14 feet, the accuracy rate for this distance range from **37%** to **47%**  
+  8.  If the shooting distance range from 14 to 16 feet, the accuracy rate for this distance range from **33%** to **51%**  
+  9.  If the shooting distance range from 16 to 18 feet, the accuracy rate for this distance range from **37%** to **48%**  
+  10. If the shooting distance range from 18 to 24 feet, the accuracy rate for this distance range from **33%** to **43%**  
+  11. If the shooting distance range from 24 to 28 feet, the accuracy rate for this distance range from **18%** to **38%**  
+  12. For shooting distance more than 28 feet, the accuracy varies greater than previous distance range, starting from **0%** (no shots scored from that distance) to **50%**
+
+  Next analysis will be calculated from correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:  
+```{r}
+#change distance level to numeric type for correlation test function
+shooting_distance_level$Var1 <- as.numeric(shooting_distance_level$Var1)
+shooting_distance_level$Var1 <- shooting_distance_level$Var1 * 0.1 - 0.1 #to maintain distance data point as it is
+
+#populate data for every shooting distance
+shootingDistanceZerotoAll <- c()
+shootingAccuracyDist1 <- c()
+for (i in 0:nrow(shooting_distance_level)){
+  shootingDistanceZerotoAll <- c(shootingDistanceZerotoAll, shooting_distance_level[i, "Var1"])
+  shootingAccuracyDist1 <- c(shootingAccuracyDist1, shooting_distance_level[i, "Accuracy"])
+}
+
+dfShootingZerotoAll <- data.frame(shootingDistanceZerotoAll, shootingAccuracyDist1)
+cor.test(dfShootingZerotoAll$shootingDistanceZerotoAll, dfShootingZerotoAll$shootingAccuracyDist1)
+```
+###Correlation Test Result  
+  The correlation test result shown strong negative correlation between shooting distance and shooting accuracy (-0.87).
+This meant that for all data points, the further a shooter from the basket, it was unlikely the shooter's attempt was success, regardless who was the shooter.
+However, if the correlation test taken for certain range distance based on the diagram, for example range from 10 feet to 20 feet, the correlation result would be different as follows:
+
+```{r}
+#populate data for shooting distance in range 10 to 24 feet
+shootingDistance10to20 <- c()
+shootingAccuracyDist2 <- c()
+for (i in 100:200){
+  shootingDistance10to20 <- c(shootingDistance10to20, shooting_distance_level[i, "Var1"])
+  shootingAccuracyDist2 <- c(shootingAccuracyDist2, shooting_distance_level[i, "Accuracy"])
+}
+
+dfShooting10to20 <- data.frame(shootingDistance10to20, shootingAccuracyDist2)
+cor.test(dfShooting10to20$shootingDistance10to20, dfShooting10to20$shootingAccuracyDist2)
+```
+  The second correlation test produced a small positive correlation between range 10 to 20 feet (0.11). This result was different with previous measurement (-0.87)
+
+###Discussion  
+  Based on the diagram and correlation, it can be concluded that although the shooting distance would directly affect shooting accuracy, but the relationship between these variables are not perfectly linear.
+This was shown in the diagram where there was a relatively **similar** shooting accuracy for shots taken from 10 feet until 20 feet, regardless the shooter's performance.
+And for the success rate in longer distances, it could be arguably inferred as the shooters tried to attempt in closing game time (buzzer beater) when the shot clock or game clock nearly reach zero.
+
+###Limitations  
+ Since the dataset mostly return 0% for shooting distance greater than 30 feet, it would better if the dataset could be enriched with previous season statistics, or even the last 10 seasons to exercise in more detail about the pattern for long distance shots attempt.
 ```{r}
@@
 grid.arrange(g1, g2, ncol = 2)
-#Klay Thompson is an exception, had 13 streaks in a row!
-We cannot find a similar distribution in our sample, so we can assume that he is likely to have a hot hand.
-Fearther statistical analysis (logistic regression) should be done in order to reach a conclusion.
+#Klay Thompson is an exception, had 13 streaks in a row! We cannot find a similar distribution in our sample, so we can assume that he is likely to have a hot hand.
+
+best_players_df
+
 ```

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
last part - hot hand

--- 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
@@
 
+##Hot Hand Theory
+###Introduction
+The analysis aims to show the effect of the 'hot hand theory' on NBA players.
+The analysis is based on the hypothesis that a shooter with a hot hand will have shots that are not independent of one another.
+If the shooter makes his first shot, the probability of making his next shot is higher.
+
+###Methods
+The steps to approximate this question are as follows:
+1. Calculate streaks achieved in all games and from all players
+2. Categorize streaks per player and game, focusing on the 3 players with the top 3 performances in terms of points
+3. Extract zeros and keep only streaks above 1
+4. Visualise streaks' length distribution - barrplots
+5. Visualise streaks' length via boxplot
+6. Calculate the shooting percentage of the 3 players 
+7. Create a sample independent shooter setting as shooting percentage the real player's shooting percentage
+8. Compare distributions of sample and real player
+
 ```{r}
-###Hot Hand Theory
 
@@
 streak_per_player
+```
+
+###Barplots
+```{r}
 
@@
 player3_df <- data.frame('streaks' = streak_per_player[[3]])
+best_players_df
+#Show distribution for each player's streak lengths
+gg1 <- ggplot(data = player1_df, aes(x = streaks)) + geom_bar(fill = 'brown') + labs(title = 'Kyrie Irnving')
+gg2 <- ggplot(data = player2_df, aes(x = streaks)) + geom_bar(fill = 'red') + labs(title = 'Mo Williams')
+gg3 <- ggplot(data = player3_df, aes(x = streaks)) + geom_bar(fill = 'blue') + labs(title = 'Klay Thompson')
+grid.arrange(gg1, gg2, gg3, ncol = 3)
+```
 
-#Show distribution for each player's streak lengths
-gg1 <- ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[1,'player_name'])
-gg2 <- ggplot(data = player2_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[2,'player_name'])
-gg3 <- ggplot(data = player3_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[3,'player_name'])
-grid.arrange(gg1, gg2, gg3, ncol = 3)
+The distribution of the three players is unimodal and left skewed. There are shown some extremely long shooting streaks made(i.e. Klay Thompson with 13 shots in a row).
 
+###Boxplots
+####Boxplot for Kyrie Irnving
+```{r}
 #Create a boxplot for each player's streak lengths 
@@
 IQR(player1_df$streaks) #Check that boxplot is valid
-#The typical length of a streak is 1 (median = 1) - The Interquartile Range is 1 - Streak length of 6 is unusually high compared to the rest of the distribution
+```
+
+1. The typical length of a streak is 1 (median = 1)
+2. The Interquartile Range is 1 
+3. Streak length of 6 is unusually high compared to the rest of the distribution
+
+####Boxplot for Mo Williams
+```{r}
 boxplot(player2_df$streaks)
@@
 IQR(player2_df$streaks)
-#The typical length of a streak is 1.5 (median = 1.5) - The Interquartile Range is 1 - Streak length of 4 is unusually high compared to the rest of the distribution
+```
+
+1. The typical length of a streak is 1.5 (median = 1.5)
+2. The Interquartile Range is 1
+3. Streak length of 4 is unusually high compared to the rest of the distribution
+
+####Boxplot for Klay Thompson
+```{r}
 boxplot(player3_df$streaks)
@@
 IQR(player3_df$streaks)
-#The typical length of a streak is 8 (median = 8) - The Interquartile Range is 5 - There is a streak lengths of 13(!!!)
+```
 
-#We can see that the distribution of all players is right skewed and that all of them (especially player3) have some long shooting streaks.
-To prove/disprove the hot hand theory we need to examine the indipendence of the shots.
-If each shot that a player takes is independent of the next shot, then a missed/made shot will not affect the propability that the next shot will be missed/made.
-So, if consecutive shots are independent, we disprove hot hand theory, while if consecutive shots are dependent, we prove the hot hand theory.
+1. The typical length of a streak is 8 (median = 8)
+2. The Interquartile Range is 5
+3. Streak length of 13 is unsually high compared to the rest of the distribution
 
+### Plots Result
+
+The distribution of all players is left skewed and all of them (especially player3) have some long shooting streaks.
+In order to prove/disprove the hot hand theory,  the examination of the indipendence of the shots is required.
+If each shot that a player takes is independent of the next shot, then the player has the same probability of hitting each shot regardless of the previous shot.
+If each shot that a player takes is dependent of the next shot, then the probability of making the next shot is higher - hot hand.
+Sampling an independent shooter having the same shooting percentage with the real one and comparing his distribution with real player's distribution gives further insight in the specific aera.
+
+```{r}
 #Find the shooting percentage for each of the 3 players
@@
 shooting_percentage
+```
 
+###Kyrie Irving
+```{r}
 ###PLAYER 1 VS SAMPLE PLAYER###
@@
 g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar() + labs(title = 'Kyrie Irving')
-grid.arrange(g1, g2, ncol = 2)
-#We can see that streak length distributions are quite similar, so we can conclude that Kyrie Irving doesn't have a hot hand.
+grid.arrange(g1, g2, ncol = 2
+```
 
+The streak length distributions are quite similar, so we can conclude that Kyrie Irving doesn't have a hot hand.
+
+### Mo Williams
+```{r}
 ###PLAYER 2 VS SAMPLE PLAYER###
@@
 grid.arrange(g1, g2, ncol = 2)
+```
 
+The streak length distributions are quite similar, so we can conclude that Mo Williams doesn't have a hot hand.
+
+###Klay Thompson
+```{r}
 ###PLAYER 3 VS SAMPLE PLAYER###
@@
 grid.arrange(g1, g2, ncol = 2)
-#Klay Thompson is an exception, had 13 streaks in a row!
-We cannot find a similar distribution in our sample, so we can assume that he is likely to have a hot hand.
-Fearther statistical analysis (logistic regression) should be done in order to reach a conclusion.
 ```
+
+Klay Thompson is an exception making 13 shots in a row.
+There is no such distribution in our sample, so it coulb be assumed that he is likely to have a hot hand.
+Further statistical analysis (logistic regression) should be done in order to reach a conclusion.
 

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
report modification

--- 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
@@
 ---
-
-```{r setup, include=FALSE}
-
-knitr::opts_chunk$set(echo = TRUE)
-```
-Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo
-
-#Report structure
-
-###Introduction
-I   What is your theme or research question?
-a. Hot Hand Theory
-b. Defender Proximity Analysis
-c. Shot Clock Pressure
-d. Close Game Effect - Small Margin affecting the scoring rate
-e. Home Advantage
-f. Fatigue Effect (Quarter Heat Rate)
-g. Shooting Percentage
-h. Fatigue Effect (Number of Rest Day)
-i.
-
-###Theory
-I   What do you expect to find?
-a.
-
-###Methods
-I   Describe the dataset:
-
-Data on shots taken during the 2014-2015 season, who took the shot, where on the floor was the shot taken from, who was the nearest defender, how far away was the nearest defender, time on the shot clock, and much more.
-The column titles are generally self-explanatory.
-Useful for evaluating who the best shooter is, who the best defender is, the hot-hand hypothesis, etc.
-Scraped from NBA's REST API.  ##TO BE UPDATED FURTHER
-
-I   Strengths
-
-I   Limitations
-
-###Analysis
-
-###Discussion
-I   What results did you find?
-
-I   Why is this interesting?
-
-I   What would be the next steps?

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
last part - hot hand
working on rest days

--- 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
@@
 restdays<-cast(df1, DATE+ HOME_TEAM + AWAY_TEAM + WINNER ~ FGM)
+
+for i in aa3$TEAM {
+  for (j in 1:nrow(restdays)) {aa3$DATES<- as.list(aa3$DATES)
+    if (restdays$HOME_TEAM[j]=="LAL" ) aa3$DATES[aa3$TEAM=="LAL"]<- c(aa3$DATES[aa3$TEAM=="LAL"], restdays$DATE[j])}
+}
+
+#| restdays$AWAY_TEAM[j])=="LAL"
+
 ```
@@
 ####Introduction  
-This analysis would calculate whether there was any significant effect from closest defender distance from a shooter with the shot outcome.
+This analysis wants to calculate whether there was any significant effect from closest defender distance from a shooter with the shot outcome.
 The hypothesis is the further a defender from the shooter, then it will be more likely to increase the chance of scoring.
@@
   4. Calculate how many succesfull shot scored from every distance point  
-  5. Calculate accuracy by dividing successfull shot scored with distance occurence for every defender distance point  
+  5. Calculate accuracy by dividing successfull shot scored with distance occurence for every distance point  
   6. Plot the result
@@
   scale_x_discrete("Closest Defender Distance", breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24)) +
-  geom_smooth() +
   ylab("Shooting Accuracy (%)") +
@@
   1. If the defender stands really close to the shooter (0 to 2 feet), the accuracy varied between **42%** to **50%**  
-  2. If the defender distance range from 2 to 4 feet, the accuracy rate for this distance range from **44%** to **49%**  
-  3. If the defender distance range from 4 to 6 feet, the accuracy rate for this distance range from **41%** to **46%**
-  4. If the defender distance range from 6 to 8 feet, the accuracy rate for this distance range from **37%** to **42%**
-  5. If the defender distance range from 8 to 10 feet, the accuracy rate for this distance range from **37%** to **54%**
+  2. If the defender distance range from 2 (two) to 4 (four) feet, the accuracy rate for this distance range from **44%** to **49%**  
+  3. If the defender distance range from 4 (four) to 6 (six) feet, the accuracy rate for this distance range from **41%** to **46%**
+  4. If the defender distance range from 6 (six) to 8 (eight) feet, the accuracy rate for this distance range from **37%** to **42%**
+  5. If the defender distance range from 8 (eight) to 10 (ten) feet, the accuracy rate for this distance range from **37%** to **54%**
   6. For defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from **33%** to **60%**.  
@@
 ```{r}
-#change distance level to numeric type for correlation test function
 distance_level$Var1 <- as.numeric(distance_level$Var1)
-distance_level$Var1 <- distance_level$Var1 * 0.1 - 0.1 #to keep the distance point as it is
+distance_level$Var1 <- distance_level$Var1 * 0.1 - 0.1
 
-#populate data points from data frame
 distanceZerotoAll <- c()
@@
 
-#combine data frame
 dfZerotoAll <- data.frame(distanceZerotoAll, accuracyDist1)
-
-#calculate correlation
 cor.test(dfZerotoFive$distanceZerotoAll, dfZerotoFive$accuracyDist1)
@@
 ```{r}
-#populate data from data frame
 distanceZerotoEight <- c()
@@
 
-#combine data frame
 dfZerotoEight <- data.frame(distanceZerotoEight, accuracyDist2)
-
-#calculate correlation
 cor.test(dfZerotoEight$distanceZerotoEight, dfZerotoEight$accuracyDist2)
 ```
-  As the analysis shown, the correlation measurement between 0 (zero) to 8 feet would be negative (-0.40). This result was different with the previous calculation done for all data points (0.59).
+  Therefore,
 
@@
 It is also shown from the dataset that only small attempt of shots were made from further defender distance affecting the accuracy calculation result.
-However, from the correlation perspective, the result would be different if the analysis included all of the data points versus only included short distance data points.
-This was arguably because further defender distance did not implied shooter distance from the basket is nearer to improve the accuracy and also there were not many players to perform shooting attempt automatically although their opponent was standing far from the shooter
+This was arguably because further defender distance did not implied shooter distance from the basket is nearer to improve the accuracy.
 
@@
 ####Introduction  
-This analysis would calculate whether was there any significant effect from shooting distance from basket with the shot outcome.
-The hypothesis is the further a shooter from the basket, then the success rate would be decreased.
-This was because when the shooter was practically in closer shooting range, the shooter shooting's result would be more accurate.
+This analysis wants to calculate whether is there any significant effect from closest defender distance from a shooter with the shot outcome.
+The hypothesis is the further a defender from the shooter, then it will be more likely to increase the chance of scoring.
+This is because the shooter is practically in "free" condition without any physical interference.
 
@@
   1. Assign shot result to determine whether the shot is succeed (1) or missed (0)  
-  2. Select distinct of **all** shooting distance  
+  2. Select distinct of **all** closest defender distance  
   3. Count the occurence fOr every distance  
   4. Calculate how many succesfull shot scored from every distance point  
-  5. Calculate accuracy by dividing successfull shot scored with particular distance occurence for every shooting distance point  
+  5. Calculate accuracy by dividing successfull shot scored with distance occurence for every distance point  
   6. Plot the result
@@
 #Calculate shooting accuracy for each distance
-shooting_distance_level$Accuracy <- round(shooting_distance_level$SCORED / shooting_distance_level$Freq, 4)
+shooting_distance_level$Accuracy <- round(shooting_distance_level$SCORED / shooting_distance_level$Freq, 2)
 
@@
 library(ggplot2)
-library(scales)
 ggplot(data = shooting_distance_level, aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) +
-  geom_point(aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) +
-  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
-  scale_x_discrete("Shooting Distance", breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 32, 34, 36, 38, 40, 42, 44, 46, 48)) +
-  geom_smooth() +
-  ylab("Shooting Accuracy (%)") +
-  ggtitle("Shooting Distance vs Shooting Accuracy")
+  geom_point(aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy))
+
+#Data Preparation
+#Change the value of shot result column to numeric
+df[, "SHOT_RESULT"] <- as.character(df[, "SHOT_RESULT"])
+for (j in 1:nrow(df)){
+  df[j, "SHOT_RESULT"] <- ifelse (df[j, "SHOT_RESULT"] == "made", 1, 0)
+}
+
+#set the value to numeric or integer accordingly
+df[, "SHOT_RESULT"] <- as.numeric(df[, "SHOT_RESULT"])
+df[, "PTS_TYPE"] <- as.integer(df[, "PTS_TYPE"])
+df[, "CLOSE_DEF_DIS"] <- as.numeric(df[, "CLOSE_DEF_DIST"])
+
+#select data with only scored result
+#Point type (2 or 3 points) and distance
+point_type <- c()
+distance <- c()
+for (z in 1:nrow(df)){
+  if(df[z, "SHOT_RESULT"] == 1 & df[z, "PTS_TYPE"] >= 1){
+    point_type <- c(point_type, df[z, "PTS_TYPE"])
+    distance <- c(distance, df[z, "CLOSE_DEF_DIST"])
+  }
+}
+
+#combine point type and distance
+valid_distance = data.frame(point_type, distance)
+
+#calculate correlation between defender distance and points made (2 or 3 points)
+cor(point_type, distance)
+
+#plot the diagram for this correlation
+library(ggplot2)
+ggplot(data = valid_distance, aes(x = point_type, y = distance, group = point_type)) + 
+  geom_boxplot() + stat_summary(fun.y = mean, geom = "point", shape = 5, size = 1) + 
+  ggtitle("Points Scored based on Distance from Defender")
+
+#calculate correlation between defender distance and shot result
+cor(df[, "SHOT_RESULT"], df[, "CLOSE_DEF_DIS"])
+
+#--------------------------------------------------------------------------------#
+
+#Further analysis
+#select distance data from 2 points shot scored and the players who scored it
+distance2 <- c()
+players2 <- c()
+for (z in 1:nrow(df)){
+  if(df[z, "SHOT_RESULT"] == 1 & df[z, "PTS_TYPE"] == 2){
+    distance2 <- c(distance2, df[z, "CLOSE_DEF_DIST"])
+    players2 <- c(players2, as.character(df[z, "player_name"]))
+  }
+}
+
+#combine 2 point players and defender distance
+player2point <- data.frame(players2, distance2)
+summary(player2point)
+
+#count players who scored 2 points
+counter <- as.data.frame(table(players2))
+
+#count 5 top players
+top5Players2 <- counter[rev(order(counter$Freq)), "players2"][1:5]
+top5Counter <- counter[rev(order(counter$Freq)), "Freq"][1:5]
+
+top5_2points <- data.frame(top5players2, top5Counter)
+
+#plot the graph
+library(ggplot2)
+ggplot(data = top5_2points, aes(x = top5Players2)) +
+  geom_bar(aes(x = top5players2, y = top5Counter), stat = "identity")
+
+#----------------------------------------------------------------#
+
+#select distance data from 3 points shot scored and the players who scored it
+distance3 <- c()
+players3 <- c()
+for (z in 1:nrow(df)){
+  if(df[z, "SHOT_RESULT"] == 1 & df[z, "PTS_TYPE"] == 3){
+    distance3 <- c(distance3, df[z, "CLOSE_DEF_DIST"])
+    players3 <- c(players3, as.character(df[z, "player_name"]))
+  }
+}
+
+#combine 3 point players and defender distance
+player3point <- data.frame(players3, distance3)
+summary(player3point)
+
+#count players who scored 3 points
+counter3 <- as.data.frame(table(players3))
+
+#count 5 top players
+top5Players3 <- counter3[rev(order(counter3$Freq)), "players3"][1:5]
+top5Counter3 <- counter3[rev(order(counter3$Freq)), "Freq"][1:5]
+
+top5_3points <- data.frame(top5Players3, top5Counter3)
+
+#plot the graph
+library(ggplot2)
+ggplot(data = top5_3points, aes(x = top5Players3)) +
+  geom_bar(aes(x = top5Players3, y = top5Counter3), stat = "identity")
 
@@
 
-###Plot Result
-  The explanation from the plot above are as follows:   
-  1.  If the shooter stands really close to the basket (0 to 2 feet), the accuracy varied between **62%** to **68%**  
-  2.  If the shooting distance range from 2 to 4 feet, the accuracy rate for this distance range from **61%** to **67%**  
-  3.  If the shooting distance range from 4 to 6 feet, the accuracy rate for this distance range from **44%** to **58%**  
-  4.  If the shooting distance range from 6 to 8 feet, the accuracy rate for this distance range from **38%** to **51%**  
-  5.  If the shooting distance range from 8 to 10 feet, the accuracy rate for this distance range from **35%** to **42%**  
-  6.  If the shooting distance range from 10 to 12 feet, the accuracy rate for this distance range from **34%** to **42%**  
-  7.  If the shooting distance range from 12 to 14 feet, the accuracy rate for this distance range from **37%** to **47%**  
-  8.  If the shooting distance range from 14 to 16 feet, the accuracy rate for this distance range from **33%** to **51%**  
-  9.  If the shooting distance range from 16 to 18 feet, the accuracy rate for this distance range from **37%** to **48%**  
-  10. If the shooting distance range from 18 to 24 feet, the accuracy rate for this distance range from **33%** to **43%**  
-  11. If the shooting distance range from 24 to 28 feet, the accuracy rate for this distance range from **18%** to **38%**  
-  12. For shooting distance more than 28 feet, the accuracy varies greater than previous distance range, starting from **0%** (no shots scored from that distance) to **50%**
-
-  Next analysis will be calculated from correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:  
-```{r}
-#change distance level to numeric type for correlation test function
-shooting_distance_level$Var1 <- as.numeric(shooting_distance_level$Var1)
-shooting_distance_level$Var1 <- shooting_distance_level$Var1 * 0.1 - 0.1 #to maintain distance data point as it is
-
-#populate data for every shooting distance
-shootingDistanceZerotoAll <- c()
-shootingAccuracyDist1 <- c()
-for (i in 0:nrow(shooting_distance_level)){
-  shootingDistanceZerotoAll <- c(shootingDistanceZerotoAll, shooting_distance_level[i, "Var1"])
-  shootingAccuracyDist1 <- c(shootingAccuracyDist1, shooting_distance_level[i, "Accuracy"])
-}
-
-dfShootingZerotoAll <- data.frame(shootingDistanceZerotoAll, shootingAccuracyDist1)
-cor.test(dfShootingZerotoAll$shootingDistanceZerotoAll, dfShootingZerotoAll$shootingAccuracyDist1)
-```
-###Correlation Test Result  
-  The correlation test result shown strong negative correlation between shooting distance and shooting accuracy (-0.87).
-This meant that for all data points, the further a shooter from the basket, it was unlikely the shooter's attempt was success, regardless who was the shooter.
-However, if the correlation test taken for certain range distance based on the diagram, for example range from 10 feet to 20 feet, the correlation result would be different as follows:
-
-```{r}
-#populate data for shooting distance in range 10 to 24 feet
-shootingDistance10to20 <- c()
-shootingAccuracyDist2 <- c()
-for (i in 100:200){
-  shootingDistance10to20 <- c(shootingDistance10to20, shooting_distance_level[i, "Var1"])
-  shootingAccuracyDist2 <- c(shootingAccuracyDist2, shooting_distance_level[i, "Accuracy"])
-}
-
-dfShooting10to20 <- data.frame(shootingDistance10to20, shootingAccuracyDist2)
-cor.test(dfShooting10to20$shootingDistance10to20, dfShooting10to20$shootingAccuracyDist2)
-```
-  The second correlation test produced a small positive correlation between range 10 to 20 feet (0.11). This result was different with previous measurement (-0.87)
-
-###Discussion  
-  Based on the diagram and correlation, it can be concluded that although the shooting distance would directly affect shooting accuracy, but the relationship between these variables are not perfectly linear.
-This was shown in the diagram where there was a relatively **similar** shooting accuracy for shots taken from 10 feet until 20 feet, regardless the shooter's performance.
-And for the success rate in longer distances, it could be arguably inferred as the shooters tried to attempt in closing game time (buzzer beater) when the shot clock or game clock nearly reach zero.
-
-###Limitations  
- Since the dataset mostly return 0% for shooting distance greater than 30 feet, it would better if the dataset could be enriched with previous season statistics, or even the last 10 seasons to exercise in more detail about the pattern for long distance shots attempt.
 ```{r}
@@
 ---
+
+```{r setup, include=FALSE}
+
+knitr::opts_chunk$set(echo = TRUE)
+```
+Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo
+
+#Report structure
+
+###Introduction
+I   What is your theme or research question?
+a. Hot Hand Theory
+b. Defender Proximity Analysis
+c. Shot Clock Pressure
+d. Close Game Effect - Small Margin affecting the scoring rate
+e. Home Advantage
+f. Fatigue Effect (Quarter Heat Rate)
+g. Shooting Percentage
+h. Fatigue Effect (Number of Rest Day)
+i.
+
+###Theory
+I   What do you expect to find?
+a.
+
+###Methods
+I   Describe the dataset:
+
+Data on shots taken during the 2014-2015 season, who took the shot, where on the floor was the shot taken from, who was the nearest defender, how far away was the nearest defender, time on the shot clock, and much more.
+The column titles are generally self-explanatory.
+Useful for evaluating who the best shooter is, who the best defender is, the hot-hand hypothesis, etc.
+Scraped from NBA's REST API.  ##TO BE UPDATED FURTHER
+
+I   Strengths
+
+I   Limitations
+
+###Analysis
+
+###Discussion
+I   What results did you find?
+
+I   Why is this interesting?
+
+I   What would be the next steps?

### 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
Rest days

--- 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
@@
 restdays<-cast(df1, DATE+ HOME_TEAM + AWAY_TEAM + WINNER ~ FGM)
+
+#this code is supposed to produce a list of dates when teams play at HOME in DATES column. for some reason it is not working.
+
+for (i in aa3$TEAM) {
+  for (j in 1:nrow(restdays)) {
+    aa3$DATES<- as.list(aa3$DATES)
+    if (restdays$HOME_TEAM[j]==i) aa3$DATES[aa3$TEAM==i]<- c(aa3$DATES[aa3$TEAM==i], restdays$DATE[j])}}
+
 ```

### 2016-10-13 MSFGroupwork-NBAHotHand.Rmd
added df for accuracy in general and df per player(not finished)

--- 2016-10-12 MSFGroupwork-NBAHotHand.Rmd
@@
 
-```{r}
 # replacing W/L in column "W" with 1/0 (1 for win, 0 for loss)
@@
 restdays<-cast(df1, DATE+ HOME_TEAM + AWAY_TEAM + WINNER ~ FGM)
-
-for i in aa3$TEAM {
-  for (j in 1:nrow(restdays)) {aa3$DATES<- as.list(aa3$DATES)
-    if (restdays$HOME_TEAM[j]=="LAL" ) aa3$DATES[aa3$TEAM=="LAL"]<- c(aa3$DATES[aa3$TEAM=="LAL"], restdays$DATE[j])}
-}
-
-#| restdays$AWAY_TEAM[j])=="LAL"
-
 ```
@@
 
-df[, "OVERTIME"] <- as.numeric()
+df[, "overtime"] <- as.numeric()
 
 for (i in 1:nrow(df)) {
-  if (df[i, "PERIOD"] >= 5) df[i, "OVERTIME"] <- 1
-  if (df[i, "PERIOD"] <= 4) df[i, "OVERTIME"] <- 0
+  if (df[i, "PERIOD"] >= 5) df[i, "overtime"] <- 1
+  if (df[i, "PERIOD"] <= 4) df[i, "overtime"] <- 0
 }
@@
 # accuracy
-acc_ot <- as.numeric
-acc_ot <- sum(df$OVERTIME == 1 & df$FGM == 1)/sum(df$OVERTIME == 1)
 
-acc_qrt1 <- as.numeric
-acc_qrt1 <- sum(df$PERIOD == 1 & df$FGM == 1)/sum(df$PERIOD == 1)
+df.acc <- data.frame()
 
-acc_qrt2 <- as.numeric
-acc_qrt2 <- sum(df$PERIOD == 2 & df$FGM == 1)/sum(df$PERIOD == 2)
+for (i in (1:7)){
+  df.acc[i, "accuracy"] <- sum(df$PERIOD == i & df$FGM == 1)/sum(df$PERIOD == i)
+  df.acc[i, "period"] <- i
+  if (df.acc$period[i] <= 4) df.acc[i, "overtime"] <- 0
+  if (df.acc$period[i] >= 5) df.acc[i, "overtime"] <- 1
+}
 
-acc_qrt3 <- as.numeric
-acc_qrt3 <- sum(df$PERIOD == 3 & df$FGM == 1)/sum(df$PERIOD == 3)
-
-acc_qrt4 <- as.numeric
-acc_qrt4 <- sum(df$PERIOD == 4 & df$FGM == 1)/sum(df$PERIOD == 4)
-
-acc_qrt5 <- as.numeric
-acc_qrt5 <- sum(df$PERIOD == 5 & df$FGM == 1)/sum(df$PERIOD == 5)
-
-acc_qrt6 <- as.numeric
-acc_qrt6 <- sum(df$PERIOD == 6 & df$FGM == 1)/sum(df$PERIOD == 6)
-
-acc_qrt7 <- as.numeric
-acc_qrt7 <- sum(df$PERIOD == 7 & df$FGM == 1)/sum(df$PERIOD == 7)
-
-acc_qrt1_4 <- as.numeric
-acc_qrt1_4 <- sum(df$PERIOD <= 4 & df$FGM == 1) / sum(df$PERIOD <= 4)
-
-t.test(df$FGM[df$PERIOD <= 4], df$FGM[df$PERIOD >= 5])
 t.test(df$FGM[df$PERIOD == 1], df$FGM[df$PERIOD == 2])
@@
 
-# accuracy for Top 10 scoring players
+# accuracy per player per period
+
+df.period <- cast(df, player_id + player_name + PERIOD ~ FGM)
+
+df.period[, "accuracy"] <- as.numeric()
+
+for (i in 1:nrow(df.period)){
+  df.period[i, "accuracy"] <- df.period[i, "1"] / sum(df.period[i, "1"], df.period[i, "0"] ) }
+
+# do analysis for top players per team in regards of accuracy per period
 
@@
 
-acc <- data.frame()
+acc <- data.frame(seconds.left = c(0:23))
 
 #accuracy for every second
-for (i in 1:23){
-  acc[i, "Accuracy"] <- (sum(df$SHOT_CLOCK > i & df$SHOT_CLOCK <= i+1 & df$SHOT_CLOCK != 99 & df$FGM == 1)/ sum(df$SHOT_CLOCK > i & df$SHOT_CLOCK <= i+1 & df$SHOT_CLOCK != 99))}
+for (i in (0:23)){
+  acc[i, "accuracy"] <- (sum(df$SHOT_CLOCK >= i & df$SHOT_CLOCK < i+1 & df$SHOT_CLOCK != 99 & df$FGM == 1)/           sum(df$SHOT_CLOCK >= i & df$SHOT_CLOCK < i+1 & df$SHOT_CLOCK != 99))}
 
-acc[, "SECONDS_LEFT"] <- c(1:23)
+ggplot(acc, aes(x = seconds.left, y = accuracy)) + geom_point() + geom_smooth()
 
-ggplot(acc, aes(x = SECONDS_LEFT, y = Accuracy)) + geom_point() + geom_smooth()
+#accuracy for players per second
 
-df[, "ACC_P_SEC"] <- as.numeric()
+##simplifying shot clock to seconds left
+
+df.acc.players <- cast(df, player_id + player_name + SHOT_CLOCK ~ FGM)
+
+for (i in (1:nrow(df.acc.players))){
+  for (j in (0:23)){
+    if (df.acc.players$SHOT_CLOCK[i] >= j & df.acc.players$SHOT_CLOCK[i] < j+1) df.acc.players[i, "seconds.left"] <- j
+    if (df.acc.players$SHOT_CLOCK[i] == 99) df.acc.players[i, "seconds.left"] <- 99}
+}
+
+df.acc.players[, "acc.per.sec"] <- as.numeric()
+
+for (i in (1:nrow(df.acc.players))){
+  df.acc.players[i, "acc.per.sec"] <- sum(df.acc.players$'0'[df.acc.players$seconds.left == df.acc.players$seconds.left[i] & df.acc.players$player_id == df.acc.players$players_id[i]])} ### doesnt work as I would like it
+
+ggplot(df.acc.players, aes(x = seconds.left, y = ))
 
 #Assign accuracy per second to main dataframe
+df[, "acc.per.sec"] <- as.numeric()
+
 for (i in 1:nrow(df)){
-  for (j in 1:23){
-    if (df$SHOT_CLOCK[i] >j & df$SHOT_CLOCK[i] <= j+1 & df$SHOT_CLOCK[i] != 99) df[i, "ACC_P_SEC"] <- acc[j, "Accuracy"]
+  for (j in 0:23){
+    if (df$SHOT_CLOCK[i] >j & df$SHOT_CLOCK[i] <= j+1 & df$SHOT_CLOCK[i] != 99) df[i, "acc.per.sec"] <- acc[j, "accuracy"]
   }
 }
-
-cov(df$ACC_P_SEC, df$SHOT_CLOCK)
-cor.test(df$ACC_P_SEC, df$SHOT_CLOCK)
 

### 2016-10-13 MSFGroupwork-NBAHotHand.Rmd
minor changes

@@
 
-# Find shooting percentage of each match
+# Find field goal percentage (FG percentage) for each match
 
@@
 
-game <- cast(data = df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + LOCATION + player_name + W + FINAL_MARGIN~ FGM)
+shot.stats <- cast(data = df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + LOCATION  + W + player_name + FINAL_MARGIN ~ FGM)
 
@@
 
-game[, "FGA"] <- NA
-for(i in 1:nrow(game)){
-  game[i, "FGA"] <- game[i, "1"] + game[i, "0"]
+shot.stats[, "FGA"] <- NA
+for(i in 1:nrow(shot.stats)){
+  shot.stats[i, "FGA"] <- shot.stats[i, "1"] + shot.stats[i, "0"]
 }
@@
 
-game[, "FG.percentage"] <- NA 
-for (i in 1:nrow(game)){
-  game[i, "FG.percentage"] <- round(((game[i, "1"]/game[i, "FGA"]) * 100), digits = 2)
+shot.stats[, "FG.percentage"] <- NA 
+for (i in 1:nrow(shot.stats)){
+  shot.stats[i, "FG.percentage"] <- round(((shot.stats[i, "1"]/shot.stats[i, "FGA"]) * 100), digits = 2)
 }
@@
 
-# Impact of Super Star Players
+# Impact of Top Scorers
 
@@
 
-super.star.list <- c("lebron james", "james harden", "russell westbrook", "stephen curry", "anthony davis", "blake griffin", "klay thompson", "lamarcus aldridge") # players picked out
-super.star <- game[game$player_name %in% super.star.list, c("player_name", "W", "LOCATION", "FINAL_MARGIN", "FGA", "FG.percentage")] # create a subset for these players
+top.scorers.list <- c("paul millsap", "isaiah thomas", "brook lopez", "kemba walker", "jimmy butler", "mnta ellis", "ty lawson", "reggie jackson", "lebron james", "james harden", "russell westbrook", "stephen curry", "anthony davis", "blake griffin", "lamarcus aldridge", "cj miles", "kobe bryant", "marc gasol", "dwayne wade", "brandon knight", "andrew wiggins", "carmelo anthony", "nikola vucevic", "michael carter-williams", "eric bledsoe", "demarcus cousins", "kawhi leonard", "kyle lowry", "gordon hayward", "john wall") # players picked out
+top.scorers <- shot.stats[shot.stats$player_name %in% top.scorers.list, c("player_name", "W", "LOCATION", "FINAL_MARGIN", "FGA", "FG.percentage")] # create a subset for these players
 
 # Replace "W" and "L" with 1 and 0 respectively, 
-super.star[, "W"] <- as.character(super.star[, "W"])
-for (i in 1:nrow(super.star)) {
-  if (super.star[i, "W"] == "W") super.star[i, "W"] <- 1
-  if (super.star[i, "W"] == "L") super.star[i, "W"] <- 0
+top.scorers[, "W"] <- as.character(top.scorers[, "W"])
+for (i in 1:nrow(top.scorers)) {
+  if (top.scorers[i, "W"] == "W") top.scorers[i, "W"] <- 1
+  if (top.scorers[i, "W"] == "L") top.scorers[i, "W"] <- 0
 }
-super.star[, "W"] <- as.numeric(super.star[, "W"])
+top.scorers[, "W"] <- as.numeric(top.scorers[, "W"])
 
-ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
-ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W))
-ggplot(super.star) + geom_point(aes(x = FINAL_MARGIN, y = FGA, colour = player_name))
-ggplot(super.star) + geom_boxplot(aes(x = LOCATION, y = FINAL_MARGIN))
+ggplot(top.scorers) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
+ggplot(top.scorers) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W))
+ggplot(top.scorers) + geom_point(aes(x = FINAL_MARGIN, y = FGA, colour = player_name))
+ggplot(top.scorers) + geom_boxplot(aes(x = LOCATION, y = FINAL_MARGIN))
 
-# t.test(super.star$FINAL_MARGIN[super.star$LOCATION == "H"], super.star$FINAL_MARGIN[super.star$LOCATION == "A"])
+# t.test(top.scorers$FINAL_MARGIN[top.scorers$LOCATION == "H"], top.scorers$FINAL_MARGIN[top.scorers$LOCATION == "A"])
 
 # Correlation test
-cor.test(super.star$FINAL_MARGIN, super.star$FGA)
-cor.test(super.star$FINAL_MARGIN, super.star$FG.percentage) # cor = 0.272
+cor.test(top.scorers$FINAL_MARGIN, top.scorers$FGA)
+cor.test(top.scorers$FINAL_MARGIN, top.scorers$FG.percentage) # cor = 0.272
 
 # Linear regression
-summary(lm(FINAL_MARGIN ~ FGA + LOCATION, data = super.star)) # Final margin ~ FGA and location
-summary(lm(FINAL_MARGIN ~ FG.percentage + LOCATION, data = super.star)) # Final margin ~ FG% and location
-summary(lm(FINAL_MARGIN ~ FG.percentage + FGA + LOCATION, data= super.star))
+summary(lm(FINAL_MARGIN ~ FGA + LOCATION, data = top.scorers)) # Final margin ~ FGA and location
+summary(lm(FINAL_MARGIN ~ FG.percentage + LOCATION, data = top.scorers)) # Final margin ~ FG% and location
+summary(lm(FINAL_MARGIN ~ FG.percentage + FGA + LOCATION, data= top.scorers))
 
-summary(lm(FINAL_MARGIN ~ LOCATION, data = super.star)) # Final margin ~ FGA and location
+summary(lm(FINAL_MARGIN ~ LOCATION, data = top.scorers)) # Final margin ~ FGA and location
 

### 2016-10-13 MSFGroupwork-NBAHotHand.Rmd
edit document layout

--- 2016-10-13 MSFGroupwork-NBAHotHand.Rmd
@@
 Data was taken from NBA's REST API provided by [Kaggle](https://www.kaggle.com/dansbecker/nba-shot-logs) and acquired at 30 September 2016.
->>>>>>> ffa732d5e8e2f51ae1f136d3572f36becfcf2f39
 
@@
 ```{r}
-#install.packages('reshape')
+install.packages('reshape')
 #install.packages('gridExtra')
@@
 for (i in 1:nrow(df)) {
-  if (LOCATION[i]=="H") {df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
+  if (LOCATION[i] == "H") {df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
   df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7]
@@
 
-restdays<-cast(df1, DATE+ HOME_TEAM + AWAY_TEAM + WINNER ~ FGM)
+restdays <- cast(df1, DATE + HOME_TEAM + AWAY_TEAM + WINNER ~ FGM, value = 'FGM')
 
@@
     if (restdays$HOME_TEAM[j]==i) aa3$DATES[aa3$TEAM==i]<- c(aa3$DATES[aa3$TEAM==i], restdays$DATE[j])}}
-
-```
-
-#we need to change the format of some of the columns to 'Character'
-#get rid off 'attach()'
-#get rid off column X in shot logs1010 file
 

### 2016-10-14 MSFGroupwork-NBAHotHand.Rmd
edit document layout
changed some code for accuracy per player

--- 2016-10-13 MSFGroupwork-NBAHotHand.Rmd
@@
 
-This area of anlysis focuses on investigation of whether playing at 'home' gives the team advantage comparing to playing 'away'.
+This area of analysis focuses on investigation of whether playing at 'home' gives the team advantage comparing to playing 'away'.
 
@@
 
+The potential effects of tiredness and stress on the accuracy of shots are investigated in this part of the analysis.
+
 ```{r}
@@
 
-# accuracy
+# accuracy per period
 
@@
 for (i in (0:23)){
-  acc[i, "accuracy"] <- (sum(df$SHOT_CLOCK >= i & df$SHOT_CLOCK < i+1 & df$SHOT_CLOCK != 99 & df$FGM == 1)/           sum(df$SHOT_CLOCK >= i & df$SHOT_CLOCK < i+1 & df$SHOT_CLOCK != 99))}
+  acc[i, "accuracy"] <- (sum(df$SHOT_CLOCK >= i & df$SHOT_CLOCK < i+1 & df$SHOT_CLOCK != 99 & df$FGM == 1)/sum(df$SHOT_CLOCK >= i & df$SHOT_CLOCK < i+1 & df$SHOT_CLOCK != 99))}
 
@@
 
+### doesnt work as I would like it:
+
 for (i in (1:nrow(df.acc.players))){
-  df.acc.players[i, "acc.per.sec"] <- sum(df.acc.players$'0'[df.acc.players$seconds.left == df.acc.players$seconds.left[i] & df.acc.players$player_id == df.acc.players$players_id[i]])} ### doesnt work as I would like it
+  for (j in (0:23)){
+    df.acc.players[i, "acc.per.sec"] <- sum(df.acc.players$'1'[df.acc.players$seconds.left == j])}
+}
 

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
add line for each sections and add analysis explanations

--- 2016-10-13 MSFGroupwork-NBAHotHand.Rmd
@@
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 ---
@@
 
+***
+
 #Report structure
@@
 ##Introduction
-###What is your theme or research question?
-    a. Hot Hand Theory (Artemis)
-    b. Defender Proximity Analysis (Reza)
-    c. Shot Clock Pressure  (Alex)
-    d. Close Game Effect - Small Margin affecting the scoring rate 
-    e. Home Advantage (Alex is working in this point)  
-    f. Fatigue Effect (Quarter Heat Rate)  
-    g. Shooting Percentage (James)
-    h. Fatigue Effect (Number of Rest Day)  
-    i. Regression Analysis (Alex)
-
-###Introduction/Summary
-
-The purpose of this report is to analyse a chosen dataset and to produce meaningful results using statistical and mathematical analysis.
+
+The purpose of this report is to analyse the chosen dataset and to produce meaningful results using statistical and mathematical analysis.
 These analyses have been conducted using RStudio software.
 This report will focused on analysis of a number of areas, demonstrate how the analysis has been conducted, discusses theory used, demonstrate and evaluate produced results.
-In addition this report will highlight potential areas of future research.
-
-The chosen dataset for this report is a summary of shots made during NBA season 2015-2016. The following areas of research have been selected:
-
-#Report structure
-
-##Introduction
-I   What is your theme or research question?
-a. Hot Hand Theory (Artemis)
-b. Defender Proximity Analysis (Reza)
-c. Shot Clock Pressure  (Alex)
-d. Close Game Effect - Small Margin affecting the scoring rate 
-e. Home Advantage (Alex is working in this point)  
-f. Fatigue Effect (Quarter Heat Rate)  
-g. Shooting Percentage (James)
-h. Fatigue Effect (Number of Rest Day)  
-i. Regression Analysis (Alex)
-
-##Theory
-I   What do you expect to find?
-a.
-
-##Methods
+In addition, where relevant, this report will highlight potential areas of future research.
+
+The chosen dataset for this report is a summary of shots made during NBA season 2014-2015. The following areas of research have been selected:
+
+**1. Team related**
+  a. Home Advantage analysis
+  b. Effect of rest on teams' performance
+
+**2. Player related**
+  a. Effects on Accuracy 
+  b. Shot Clock Pressure
+  c. Fatigue Effect (Quarter Accuracy Rate)  
+  d. Game result impact by field goal attempts
+  e. Defender Proximity Analysis
+  f. Shooting Distance vs Shot Accuracy 
+  g. Hot Hand Theory
+
 ###Dataset Description:
@@
   Data on shots taken during the NBA 2014-2015 season for regular matches throughout the year.
-The data comprise of 128,609 rows and 22 columns with 16 MB data size with several value in some column is missing.
+The data comprise of 128,609 rows and 22 columns with 16 MB data size with several value in some column missing.
 The column titles are generally self-explanatory.
 It has the following data:
+
+Data  contains full set of shot attempts by each team/player during the NBA 2014-2015 season for regular matches throughout the year.
+The data comprise of 128,609 rows and 22 columns with.
+The column titles (variables) are as follows:
     1.  Match ID  
-    2.  Match date  
-    3.  Which team plays as home and away  
+    2.  Match date and names of contestants  
+    3.  Location: home and away  
     4.  Match result  
     5.  Final score difference  
-    6.  Every shots attempt per game for each player  
-    7.  Shot result whether it is success or not  
-    8.  Type of shoot attempt (2 points or 3 points)  
-    9.  Quarter number data (1 to 4)  
-    10. Game clock for each quarter  
-    11. Time left for a shot (shot clock)  
-    12. Who took the shot  
+    6.  shot number  
+    7.  FGM: Shot result: success or miss 
+    8.  Type of shot attempt (2 points or 3 points)  
+    9.  Quarter number (1 to 4, or higher for overtime)  
+    10. Game clock for each quarter when the shot has been made 
+    11. Shot clock (time left for a shot)  
+    12. Name and ID of a player who took the shot  
     13. Shot distance from basket  
-    14. Who was the nearest defender  
+    14. Name and ID of the nearest defender  
     15. Distance from the nearest defender
 
-  The dataset is useful for evaluating who the best shooter is, who the best defender is, the hot-hand hypothesis, affect of home playing team, close game effect, shooting accuracy from best players, and defender distance effect on shooting accuracy.
-Data was taken from NBA's REST API provided by [Kaggle](https://www.kaggle.com/dansbecker/nba-shot-logs) and acquired at 30 September 2016.
+ The data has been acquired from Kaggle website **(https://www.kaggle.com/dansbecker/nba-shot-logs)** on 30 September 2016.
 
@@
 
-    The dataset comprise of comprehensive observational data to start to work with.
-In general, the dataset only have small percentage of missing value (0.19%) so we do not need to remove or assign many data points to default values.
-And since the column name is not ambiguous it really helps us to understand dataset from the beginning.
+    The dataset comprises of comprehensive observational data to start to work with.
+In general, the dataset has only a small percentage of missing value (0.19%) so there was not need to remove or assign many data points to default values.
+Since the column name is not ambiguous it was really helpful to understand the dataset right from the beginning.
+
+    The dataset comprises of comprehensive observational data which enables efficient analysis without extensive 'clean up' of the data.
+The dataset only have small percentage of missing value (0.19%), therefore there is no need to remove or assign numerous data points to default values.
+The fact that initial dataset meant that only a small number of additional columns had to be added (see data clean-up part of the code for details.
 
 ###Dataset Limitations  
-  There are several constraints we need to mentioned in order to provide same understandings for this group work analysis as follows:  
-    1. The dataset consists only for 1 year NBA season (2014-2015)  
-    2. The dataset comprise only with regular match (82 matches per team for both home and away matches)  
-    3. The playoff and other matches during the year is not included
-    4. The dataset does not include free throw shot and results
+  There are several limitations of the dataset:  
+    1. The dataset consists only 1 year NBA season stats (2014-2015)  
+    2. The dataset comprise only the data for regular season matches  
+    3. The playoff and other matches during the year are not included
+    4. The dataset does not include information on free throw shots
+    5. The dataset does not include final scores of the games
+
+***
 
@@
 
-### Installing libraries
-
-```{r}
+```{r, eval=FALSE, include=FALSE}
+# Installing libraries
 install.packages('reshape')
-#install.packages('gridExtra')
-```
-
-### Calling libraries
-```{r}
+install.packages('gridExtra')
+
+```
+
+```{r, include=FALSE}
+# Calling libraries
 library(reshape)
@@
 library(gridExtra)
-```
-
+library(plyr)
+
+```
+
+```{r, include=FALSE}
 ###Data clean-up
 
-```{r}
-df<- read.csv("shot_logs.csv")  #assigning the dataframe to 'df' variables
-attach(df) #telling R that this is the dataframe we'll be working with (eliminating the need to use df$ for variables)
-
-df1<- read.csv("shot_logs1010.csv")
+df <- read.csv("shot_logs.csv")  #assigning the dataframe to 'df' variables
+ #telling R that this is the dataframe we'll be working with (eliminating the need to use df$ for variables)
+
+df1 <- read.csv("shot_logs1010.csv")
 
 # replacing W/L in column "W" with 1/0 (1 for win, 0 for loss)
-df[,"W"] <- as.character(df[,"W"])
+df[, "W"] <- as.character(df[, "W"])
 
 for (i in 1:nrow(df)) {
-  if (df[i,"W"]=="W") df[i,"W"]<- 1
-  if (df[i,"W"]=="L") df[i,"W"]<- 0}
-
-df[,"W"] <- as.numeric(df[,"W"])
-```
-
-```{r}
+  if (df[i, "W"] == "W") df[i, "W"] <- 1
+  if (df[i, "W"] == "L") df[i, "W"] <- 0
+}
+
+df[, "W"] <- as.numeric(df[, "W"])
+```
+
+```{r, include=FALSE}
 #adding columns for names of 'home' and 'away' teams. In addition, WINNER column tells which team won
-df[,"MATCHUP"] <- as.character(df[,"MATCHUP"])
-df[,"W"] <- as.character(df[,"W"])
+df[, "MATCHUP"] <- as.character(df[, "MATCHUP"])
+df[, "W"] <- as.character(df[, "W"])
 
 for (i in 1:nrow(df)) {
-  if (LOCATION[i] == "H") {df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
-  df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7]
-  if (W[i]=="W") df$WINNER[i] <- "HOME" 
-  if (W[i]=="L") df$WINNER[i] <- "AWAY"}
-  if (LOCATION[i]=="A") {df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[7] 
-  df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i]," "))[5]
-  if (W[i]=="L") df$WINNER[i] <- "HOME"  
-  if (W[i]=="W") df$WINNER[i] <- "AWAY"}}
+  if (df$LOCATION[i] == "H") {
+    df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i], " "))[5]
+    df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i], " "))[7]
+    if (df$W[i] == "W") df$WINNER[i] <- "HOME" 
+    if (df$W[i] == "L") df$WINNER[i] <- "AWAY"
+  }
+  if (df$LOCATION[i] == "A") {
+    df$HOME_TEAM [i] <- unlist(strsplit(df$MATCHUP[i], " "))[7] 
+    df$AWAY_TEAM [i] <- unlist(strsplit(df$MATCHUP[i], " "))[5]
+    if (df$W[i] == "L") df$WINNER[i] <- "HOME"  
+    if (df$W[i] == "W") df$WINNER[i] <- "AWAY"
+  }
+}
 
 #adding a column 'DATE'
-df1[,"MATCHUP"] <- as.character(df1[,"MATCHUP"])
-for (i in 1:nrow(df1)) df1$DATE[i] <- as.character(as.Date(unlist(strsplit(df1$MATCHUP[i]," -"))[1], format= "%B %d, %Y"))
-
-restdays <- cast(df1, DATE + HOME_TEAM + AWAY_TEAM + WINNER ~ FGM, value = 'FGM')
-
-#this code is supposed to produce a list of dates when teams play at HOME in DATES column. for some reason it is not working.
-
-for (i in aa3$TEAM) {
-  for (j in 1:nrow(restdays)) {
-    aa3$DATES<- as.list(aa3$DATES)
-    if (restdays$HOME_TEAM[j]==i) aa3$DATES[aa3$TEAM==i]<- c(aa3$DATES[aa3$TEAM==i], restdays$DATE[j])}}
-
-```
+df1[, "MATCHUP"] <- as.character(df1[, "MATCHUP"])
+for (i in 1:nrow(df1)) df1$DATE[i] <- as.character(as.Date(unlist(strsplit(df1$MATCHUP[i], " -"))[1], format = "%B %d, %Y"))
+
+#creating a column for 'TEAM'. This column relates to the team whose player made a shot described by the dataframe
+for (i in 1:nrow(df1)) df1$TEAM[i] <-unlist(strsplit(df1$MATCHUP[i], " "))[5]
+
+#preparing dataframe for rest days calculation
+df1$BLANK2<-""
+restdays2 <- cast(df1, TEAM + DATE + LOCATION + W + FINAL_MARGIN ~ FGM)
+restdays2$DATE<- as.Date(restdays2$DATE)
+
+```
+
+***
 
@@
 
-This area of anlysis focuses on investigation of whether playing at 'home' gives the team advantage comparing to playing 'away'.
-
-Literature review demonstrates that
-
-The NULL Hypothesis is:
+This area of analysis focuses on investigation of whether playing at 'home' gives the team advantage comparing to when playing 'away'.
+Esssentially, this analysis focuses on effect of the game lovation (home/away) on winning ratio.
+
+The NULL Hypothesis: Teams playing at home is more likely to win then when playing away.
+
+###Methods
+Steps to test the hypothesis are as follows:  
+    1. Calculate a total number of games, which has been won by a home or away team
+    2. Compare the total win numbers to identify whether home advantage is evident using means of the variables
+    3. Extend the analysis to review of the home/away wins on a team level 
+    4. Plot the team level data 
+    5. Discuss the findings of the analysis
+
+```{r, include=FALSE}
+#preparing a dataframe for home advantage analysis
+aa <- cast(df1, GAME_ID1 + HOME_TEAM + AWAY_TEAM + WINNER ~ W)
+
+#table showing a number of games won by home and away teams in the whole season
+table(aa$WINNER)
+
+#home wins:
+hh <- sum(aa$WINNER == "HOME")
+#Away wins:
+aw <- sum(aa$WINNER == "AWAY")
+
+#percentage of home team winning: 
+round(hh / (hh + aw), 4)
+round(aw / (hh + aw), 4)
+
+```
+
+```{r, include=FALSE}
+#preparing a dataframe for analysis of games when a team played at home 
+aa1 <- cast(aa, HOME_TEAM ~ WINNER); names(aa1)[1] <- "TEAM"
+aa1$WIN_PCT_HOME <- round(aa1$HOME / (aa1$HOME + aa1$AWAY), 2)
+
+#preparing a dataframe for analysis of games when a team played at away 
+aa2 <- cast(aa, AWAY_TEAM ~ WINNER); names(aa2)[1] <- "TEAM"
+aa2$WIN_PCT_AWAY <- round(aa2$AWAY / (aa2$HOME + aa2$AWAY), 2)
+
+#merging two dataframes together to enable the effective analysis of win ratios for teams playing at home and away.
+aa3 <- merge (aa1, aa2, by ="TEAM")
+
+#ordering the dataframe by win percentage at home.
+aa3 <- aa3[order(aa3$WIN_PCT_HOME, decreasing = TRUE),]
+
+#ordering factors of the dataframe by win percentage at home.
+aa3$TEAM <- factor(aa3$TEAM, levels = aa3$TEAM[order(-aa3$WIN_PCT_HOME)])
+
+#creating a column which shows a difference between home and away winning percentage rate
+aa3$WIN_PCT_DIFF<- aa3$WIN_PCT_HOME - aa3$WIN_PCT_AWAY
+
+```
+
+###Findings
+Table 1: Proportion of total games split by Location win factor
+
+X     | HOME WIN | AWAY WIN
+----- | -------- | ---
+Game  | 506      | 398
+Perc  | 55.97%   | 44.03%
+
+Tables 2 & 3: Winning ratio of teams playing at home vs games playing away
+
+TEAM        | GSW | ATL | POR | MEM | SAS | CLE | HOU | LAC | OKC | DAL | NOP | WAS | TOR | CHI | MIL  
+------------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----
+Win % HOME  |   93| 88  | 81  | 77  | 75  | 72  | 72  | 72  | 71  | 68  | 68  | 68  | 66  | 59  | 59 
+Win % AWAY  |   69| 71  | 54  | 67  | 50  | 53  | 61  | 57  | 42  | 61  | 40  | 43  | 59  | 66  | 45 
+Win % Diff  |   24| 17  | 27  | 10  | 25  | 19  | 11  | 15  | 29  | 7   | 28  | 25  | 7   | -7  | 14
+
+TEAM        | PHX | IND | BOS | CHA | UTA | DEN | BKN | SAC | MIA | DET | LAL | ORL | PHI | NYK |MI 
+------------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----
+Win % HOME  | 57  | 52  | 48  | 47  | 45  | 42  | 41  | 41  | 39  | 38  | 31  | 30  | 28  | 27  | 26 
+Win % AWAY  | 47  | 35  | 32  | 41  | 35  | 30  | 44  | 30  | 48  | 39  | 23  | 31  | 16  | 14  | 18 
+Win % Diff  | 10  | 17  | 16  | 6   | 10  | 12  |-3   | 11  |-9   |-1   |    8|   -1|   12|   13| 8
 
 ```{r}
-aa<- cast(df1, GAME_ID1 + HOME_TEAM + AWAY_TEAM + WINNER ~ W)
-
-table(aa$WINNER)
-#home wins:
-hh<-sum(aa$WINNER=="HOME")
-#Away wins:
-aw<-sum(aa$WINNER=="AWAY")
-#percentage of home team winning: 
-round(hh/(hh+aw),2)
-round(aw/(hh+aw),2)
-
-```
-
+#plot showing home and away win ratio per team
+ggplot(aa3, aes(x = TEAM)) + geom_point(aes(y = WIN_PCT_HOME, col = "HOME")) + geom_point(aes(y = WIN_PCT_AWAY, col = "AWAY")) 
+```
+
+The plot shows that out of 30 teams:
+  25 teams have won more games at home than away
+  3 teams have won more games away than at home
+  2 teams home and away win ratios are almost equal (difference of less than or equal to 1%)
+
+###Variance test Analysis
 ```{r}
-head(aa)
-aa1<- cast(aa, HOME_TEAM  ~ WINNER); names(aa1)[1]<-"TEAM"
-aa1$WIN_PCT_HOME<-round(aa1$HOME/(aa1$HOME+aa1$AWAY),2)
-
-aa2<- cast(aa, AWAY_TEAM ~ WINNER); names(aa2)[1]<-"TEAM"
-aa2$WIN_PCT_AWAY<-round(aa2$AWAY/(aa2$HOME+aa2$AWAY),2)
-
-aa3<- merge (aa1,aa2, by ="TEAM")
-
-for (i in 1:nrow(aa3)) {if(aa3$WIN_PCT_HOME[i]>aa3$WIN_PCT_AWAY[i]) aa3$HOME_ADV[i]<-"YES" else aa3$HOME_ADV[i]<-"NO"}
-head(aa3)
-
-aa3<- aa3[order(aa3$WIN_PCT_HOME,decreasing=TRUE),]
-rownames(aa4)<-NULL
-
-aa3$TEAM <- factor(aa3$TEAM, levels = aa3$TEAM[order(-aa3$WIN_PCT_HOME)])
-
-ggplot(aa3,aes(x=TEAM)) + geom_point(aes(y= WIN_PCT_HOME, col= "HOME"))+ geom_point(aes( y= WIN_PCT_AWAY, col= "AWAY"))
-
-```
-
-#Close Game Effect (Justin)
-
-```{r}
-# create new variable to seperate overtime
-
-df[, "overtime"] <- as.numeric()
-
-for (i in 1:nrow(df)) {
-  if (df[i, "PERIOD"] >= 5) df[i, "overtime"] <- 1
-  if (df[i, "PERIOD"] <= 4) df[i, "overtime"] <- 0
-}
-
-# accuracy
+
+```
+**JUSTING TO SEE IF THERE IS ANYTHING WE CAN INCLUDE HERE
+
+###regression analysis
+
+```{r, include=FALSE}
+margreg<- lm(FINAL_MARGIN ~ LOCATION, restdays2)
+summary(margreg)
+```
+
+Table 4: Regression Analysis
+
+  X         |Estimate |Std. Error| t value| Pr(>t)    
+------------|---------|--|--|------------------------
+(Intercept) |-2.12919 |   0.51589 | -4.127 |3.84e-05 ***
+LOCATIONH   |4.28556  |  0.63881  | 6.709 |2.63e-11 ***
+rest        |-0.01129 |   0.25992 | -0.043 |   0.965
+
+Adjusted R-squared = 0.02392
+
+###Discussion
+
+The analysis shown that there is a clear evidence that a team playing at home is more likely to win with an average win probability of 55.97%.
+The same conclusion is evident from the analysis conducted on a team level where 25 teams are winning more games at home than away.
+The result shown that the NULL hypothesis should be rejected as a team playing at home is more likely to win then an away team.
+Regression analysis shown that playing at home gives teams on average additional 4.29 margin points.
+
+The result of this research also confirms recent trend of decline in home advantage over the years: probability of a home team winning has reduced from around 65% in 1975-1992 to average of 60.3% in 1993-2011 and to around 58.5% for 2011-15.
+Home win ratio was 57.4% win in 2014-2015 season.
+The analysis demonstrated that the home win ratio went down even further to 56.0% in 2015-2016 season.
+Based on the literature review this trend could be potentially explained by the following factors (Economist, 2015):
+
+  1.
+Improvement of travel conditions.
+This means that when playing away, teams are now more likely to stay at better hotels and fly charterd planes more often than in the past.
+These changes result in teams being less tired as a consequence of travel and being better physically prepared for the games.
+
+  2.
+Change in style of play in NBA.
+The game now is more open with less physical contact, which reduces number of fouls and potential impact of 'notorious home-team bias' by referees (Economist, 2015).
+
+##Effect of rest on teams' performance
+
+###Introduction
+
+This area of analysis focuses on investigation of whether a number of rest days between games affect performance and more importantly winning ratio of a team.
+
+The Null Hypothesis is: By having longer rest a team is likely to have a higher winning ratio.
+
+###Methods
+The steps to calculate this measurement are as follows:  
+    1. Calculating number of rest days between games for all teams
+    2. Calculating ratios for effects of rest days on win/loss
+    3. Calculating ratios for effects of rest days on win/loss form home/awayCalculating ratios for effects of rest days on win/loss
+    4. Discuss the findings of the data
+
+```{r, include=FALSE}
+
+#calulating number of rest days between teams
+for (i in 2: nrow (restdays2)) {
+  if (restdays2$TEAM[i]==restdays2$TEAM[i-1])
+    restdays2$rest[i]<- as.numeric(restdays2$DATE[i] - restdays2$DATE[i-1]-1)
+  else 
+    (restdays2$rest[i]<- "NA")  
+}
+restdays2$BLANK<-""
+
+restdays2$rest <- as.numeric (restdays2$rest)
+
+#regression analysis
+margreg<- lm(FINAL_MARGIN ~ LOCATION + rest, restdays2)
+summary(margreg)
+
+#creating dataframe for total rest days
+restdays4<- cast (restdays2, rest ~  W)
+colnames(restdays4)<- c( "rest", "L","W")
+
+#creating a column for total restdays for a team when lost and won to enable calculation of ratios
+restdays4$total<-restdays4$L + restdays4$W
+restdays4$dist <-   round (restdays4$total * 100 /sum (restdays4$total) ,2)
+
+#plotting distribution of rest days to establish which of the rows should be analysed
+restdays4[,c(1,4,5)]
+#The distribution of rest days show that 92% of games are covered by 0,1 or 2 days of rest and therefore further analysis will only focus on these 3 sections
+
+restdays4$ L.pct <-   round(restdays4$L*100/(restdays4$L+restdays4$W),2) 
+restdays4$ W.pct <-   round(restdays4$W*100/(restdays4$L+restdays4$W),2)
+
+print (restdays4[1:3,c(1,3,7)])
+
+#creating a new dataframe for summary of home/away win losses based on number of rest days.
+restdays3<- cast (restdays2, rest ~ LOCATION + W) ;
+
+#renaming the columns
+colnames(restdays3)<- c( "rest", "AL","AW","HL", "HW")
+
+#caluclating ratios
+restdays3$ AL.pct <-   round(restdays3$AL*100/(restdays3$AL+restdays3$AW),2) 
+restdays3$ AW.pct <-   round(restdays3$AW*100/(restdays3$AL+restdays3$AW),2)
+restdays3$ HL.pct <-   round(restdays3$HL*100/(restdays3$HL+restdays3$HW),2)
+restdays3$ HW.pct <-   round(restdays3$HW*100/(restdays3$HL+restdays3$HW),2)
+
+print (restdays3[1:3,c(1,7,9)])
+
+#Testing of rest data
+
+#t.test(x=restdays2$W[restdays2$rest==0], y= restdays2$W[restdays2$rest==1])
+#chisq.test(x=restdays2$W[restdays2$rest==0], y= restdays2$W[restdays2$rest==1])
+
+#restdays2$W[restdays2$rest==0,]
+
+```
+
+###Findings
+
+Table 5: Distribution of number of games per rest days
+
+ Rest Days| Games | Distribution
+----------|-------|---
+    0  | 427 | 23.62
+    1  | 990 | 54.76
+    2  | 254 | 14.05
+    3  |  59 |  3.26
+    4  |  13 |  0.72
+    5  |   3 | 0.17
+    7  |   6 | 0.33
+    8  |  20 | 1.11
+    9  |   3 | 0.17
+    10 |   3 | 0.17
+   1st game of season  |  30|  1.66
+
+The distribution of rest days shows that 92% of games are covered by 0,1 or 2 days of rest and therefore further analysis will only focus on these 3 sections.
+
+Table 6: Win Rate split by number of rest days between games
+
+  rest days| number of wins| win percentage
+-|-|-
+    0| 201| 47.07
+    1| 496| 50.10
+    2| 137| 53.94
+
+Table 7: Win Rate split by number of rest days between games and location
+
+  rest days| home wins percentage | away wins percentage|
+-|-|-
+    0|   53.54 |44.33 
+    1 | 54.68 |  44.24
+    2  | 59.57 | 46.90
+
+###Discussion
+
+The analysis demonstrated that overall winning ratio improves with increase in  rest days: back-to-back games had 47.07% winning ratio, 1 day rest increased winning ratio to 50.10% and with 2 days rest the winning ration has increased even further to 53.94%.
+
+By extending the analysis to the team level, the results have demonstrated similar tendency for home  games where winning ratio has increased from 53.54% to 54.68% and 59.57% respectively with 0, 1 and 2 days of rest.
+Somewhat interestingly the analysis of effect of rest days on winning ratio of away games demonstrated that there is hardly any difference between 0 and 1 days of rest (44.33% and 44.24%).
+However with 2 days of rest teams' average winning ratio have increased to 46.90% (2.66% increase).
+This perhaps could be explained by the fact that teams were able to minimise effect of travelling when having a bigger window between games.
+
+Based on the above, the analysis shown that the null hypothesis should be rejected as there is a clear evidence that the higher rest days correspond to higher winning ratio
+
+###Further research
+
+To better understand effects of rest and home/away advantage the future research should focus on distances the teams have to travel between the games.
+For example if a team is travelling from one coast to another- it is likely to affect them more comparing to a team from New York 'travelling' to play an away gamein New Jersey.
+
+***
+
+##Effects on accuracy
+
+The potential effects of fatigue and stress on the accuracy of shots are investigated in this part of the analysis.
+
+Fatigue in basketball is a broadly analysed topic in physiological research.
+Latest literature suggests that elite basketball players cope well with the effects of fatigue so it does not affect their shooting kinematics.
+This results in no difference for their shot accuracy even if the players are fatigued (Erculj and Supej, 2009).
+This research paper analyses whether these previous findings can be confirmed for the NBA season 2014-2015.
+
+Regarding the effect of psychological stress on shot accuracy, previous research is not as decisive as for the effects of fatigue.
+While Ahart(1973, in P.C.
+Kendall & S.D Hollon (1979) *Cognitive-Behavioral Interventions: Theory, Research, and Procedures*) concluded that stressful events like small or big margins for the game score have an effect on players free-throw accuracy, latest research introducing more controlled conditions states that there is no effect of stress on players shot accuracy (Mascret et al., 2016).
+The indecisiveness of previous literature led to a deeper analysis in this research paper.
+
+Hypotheses for this area of analysis are:
+
+Fatigue: The Null hypothesis states that there is no significant difference (p < .05) for the shot accuracy in later periods of the game.
+
+Stress: The Null hypothesis states that there is no significant difference (p < .05) for the shot accuracy in more stressful events.
+
+###Methods
+
+In this analysis fatigue was measured as the period of the game, as it can be assumed that the later in the game the higher the fatigue of the players.
+
+The steps to test the first hypothesis were as follows:
+  1. The overall accuracy for each period was calculated, based on the number of shots made (FGM = 1) and the number of shots taken for the specific period (FGM = 1 & FGM = 0)
+  2. This results in a percentage value of shots made for each of the periods
+  3. A variable called "Overtime" was introduced to distinguish between periods in the regular game time and overtime
+  4. The average accuracy for regular time and overtime was calculated
+  5. A T-test was used to examine whether or not these values were significant different at a level of significance of 5%.
+
+Psychological stress was measured through the remaining time on the shot clock for each shot taken. It was assumed that a player would experience more stress the lower the shot clock goes.
+
+The steps to test the second hypothesis were as follows:
+  1. As the shot clock variable contained 5567 missing values, those were excluded from the analysis
+  2. To simplify the shot clock variable, decimals were removed from the data so every shot taken was assigned to 0 to 24 seconds left on the shot clock
+  3. The overall accuracy for each second was calculated using the FGM variable
+
+```{r, include=FALSE}
+
+### FATIGUe
+
+# accuracy per period
 
@@
 for (i in (1:7)){
-  df.acc[i, "accuracy"] <- sum(df$PERIOD == i & df$FGM == 1)/sum(df$PERIOD == i)
+  df.acc[i, "accuracy"] <- sum(df$PERIOD == i & df$FGM == 1) / sum(df$PERIOD == i)
   df.acc[i, "period"] <- i
@@
 }
+
+regular <- sum(df.acc$accuracy[df.acc$period <= 4])/4
+overtime <- sum(df.acc$accuracy[df.acc$period >=5])/3
 
@@
 t.test(df$FGM[df$PERIOD == 2], df$FGM[df$PERIOD == 4])
-
-# accuracy per player per period
-
-df.period <- cast(df, player_id + player_name + PERIOD ~ FGM)
-
-df.period[, "accuracy"] <- as.numeric()
-
-for (i in 1:nrow(df.period)){
-  df.period[i, "accuracy"] <- df.period[i, "1"] / sum(df.period[i, "1"], df.period[i, "0"] ) }
-
-# do analysis for top players per team in regards of accuracy per period
-
+t.test(df$FGM[df$PERIOD == c(1:4)], df$FGM[df$PERIOD == c(5:7)])
+
+### STRESS
 # accuracy for Shot Clock
@@
 # dealing with missing values
-df[is.na(df)] <- 99
-
-acc <- data.frame(seconds.left = c(0:23))
+df1[is.na(df1)] <- 99
+
+acc <- data.frame(seconds.left = c(0:24))
 
 #accuracy for every second
-for (i in (0:23)){
-  acc[i, "accuracy"] <- (sum(df$SHOT_CLOCK >= i & df$SHOT_CLOCK < i+1 & df$SHOT_CLOCK != 99 & df$FGM == 1)/           sum(df$SHOT_CLOCK >= i & df$SHOT_CLOCK < i+1 & df$SHOT_CLOCK != 99))}
+for (i in (0:24)){
+  acc[i, "accuracy"] <- (sum(df1$SHOT_CLOCK >= i & df1$SHOT_CLOCK < i + 1 & df1$SHOT_CLOCK != 99 & df1$FGM == 1) / sum(df1$SHOT_CLOCK >= i & df1$SHOT_CLOCK < i + 1 & df1$SHOT_CLOCK != 99))
+}
 
@@
 
-df.acc.players <- cast(df, player_id + player_name + SHOT_CLOCK ~ FGM)
+df.acc.players <- cast(df1, player_id + player_name + SHOT_CLOCK ~ FGM)
 
 for (i in (1:nrow(df.acc.players))){
-  for (j in (0:23)){
-    if (df.acc.players$SHOT_CLOCK[i] >= j & df.acc.players$SHOT_CLOCK[i] < j+1) df.acc.players[i, "seconds.left"] <- j
-    if (df.acc.players$SHOT_CLOCK[i] == 99) df.acc.players[i, "seconds.left"] <- 99}
-}
-
-df.acc.players[, "acc.per.sec"] <- as.numeric()
-
-for (i in (1:nrow(df.acc.players))){
-  df.acc.players[i, "acc.per.sec"] <- sum(df.acc.players$'0'[df.acc.players$seconds.left == df.acc.players$seconds.left[i] & df.acc.players$player_id == df.acc.players$players_id[i]])} ### doesnt work as I would like it
-
-ggplot(df.acc.players, aes(x = seconds.left, y = ))
-
-#Assign accuracy per second to main dataframe
-df[, "acc.per.sec"] <- as.numeric()
-
-for (i in 1:nrow(df)){
-  for (j in 0:23){
-    if (df$SHOT_CLOCK[i] >j & df$SHOT_CLOCK[i] <= j+1 & df$SHOT_CLOCK[i] != 99) df[i, "acc.per.sec"] <- acc[j, "accuracy"]
-  }
-}
-
-```
-
-# Shooting Percentage Analysis (James)
+  for (j in (0:24)){
+    if (df.acc.players$SHOT_CLOCK[i] >= j & df.acc.players$SHOT_CLOCK[i] < j + 1) df.acc.players[i, "seconds.left"] <- j
+    if (df.acc.players$SHOT_CLOCK[i] == 99) df.acc.players[i, "seconds.left"] <- 99
+  }
+}
+
+df.acc.secondsplayers<- df.acc.players
+
+df.acc.secondsplayers$SHOT_CLOCK <- NULL
+
+colnames(df.acc.secondsplayers)<- c( "player_id", "player_name","X0", "X1","seconds.left")
+
+df.acc.secondsplayers2 <- ddply(df.acc.secondsplayers, c("player_id", "player_name","seconds.left"), summarise,
+X0= sum(X0),
+X1= sum(X1))
+
+df.acc.secondsplayers2$acc <-  df.acc.secondsplayers2$X1/(df.acc.secondsplayers2$X0+df.acc.secondsplayers2$X1)
+
+df.acc.secondsplayers2[df.acc.secondsplayers2 == 99] <- NA
+
+df.acc.secondsplayers3 <- data.frame(row.names = c(0:24))
+df.acc.secondsplayers3$seconds.left <- c(0:24)
+
+for (i in df.acc.secondsplayers3$seconds.left){
+  df.acc.secondsplayers3$acc.top[df.acc.secondsplayers3$seconds.left == i] <- sum(df.acc.secondsplayers2$acc[df.acc.secondsplayers2$seconds.left == i & df.acc.secondsplayers2$player_name == "james harden"][!is.na(df.acc.secondsplayers2$acc[df.acc.secondsplayers2$seconds.left == i & df.acc.secondsplayers2$player_name == "james harden"])], df.acc.secondsplayers2$acc[df.acc.secondsplayers2$seconds.left == i & df.acc.secondsplayers2$player_name == "lebron james"][!is.na(df.acc.secondsplayers2$acc[df.acc.secondsplayers2$seconds.left == i & df.acc.secondsplayers2$player_name == "lebron james"])], df.acc.secondsplayers2$acc[df.acc.secondsplayers2$seconds.left == i & df.acc.secondsplayers2$player_name == "russell westbrook"][!is.na(df.acc.secondsplayers2$acc[df.acc.secondsplayers2$seconds.left == i & df.acc.secondsplayers2$player_name == "russell westbrook"])])/3}
+
+ggplot(NULL) + geom_smooth(data = df.acc.secondsplayers3, aes(x = seconds.left, y = acc.top, color = "Top Players")) + geom_smooth(data = acc, aes(x = seconds.left, y = accuracy, color = "Average"))
+
+t.test(acc$accuracy, df.acc.secondsplayers3$acc.top)
+
+```
+
+###Results
+
+The accuracies for each period are detailed in the table below.
+
+Period  | Accuracy
+------- | ---------
+1       | 46%
+2       | 45%
+3       | 46%
+4       | 44%
+5       | 39%
+6       | 43%
+7       | 37%
+regular | 45%
+overtime | 40%
+
+The T-tests for period 1 and period 2 was significant, t(65301) = 2.42, p = .015, as well as the T-test for period 3 and 4, t(60757) = 4.24, p < .01.
+Further the T-test for period 2 and 3 was not significant, t(63843) = -1.53, p = .126, as well as the T-test for period 1 and 3, t(65989) = 0.87, p = .38.
+The T-test for regular time and overtime was highly significant, t(383) = 2.77, p < .01.
+
+The changes in the overall accuracy per second are shown in the graph below.
 
 ```{r}
+graph.acc <- ggplot(acc, aes(x = seconds.left, y = accuracy)) + geom_point() + geom_smooth()
+
+graph.acc + labs(title = "Overall Accuracy per second left on shot clock") + xlab("Seconds left on shot clock") + ylab("Overall Accuracy")
+```
+
+### Discussion
+
+The results of the T-tests show a significant difference for the overall accuracy between period 1 and period 2, between period 3 and period 4 and between regular game time and overtime.
+This means that player have a higher accuracy in periods 1 and 3 than in the following one.
+As the table shows, the changes in accuracy are minor (from 46% to 44% over the regular game time) and often located on digit-level, so although the changes are significant they are negligible.
+After all, the Null hypothesis must be rejected as the statistical methods used show differences for accuracy in later periods of the game, but as they are negligible in the regular game time but it can be concluded that accuracy does not suffer for later game periods as long as the game ends in regular time.
+For games that reach overtime a highly significant drop in the accuracy of over 5% (45% to below 40%) is observale, stating that players are less likely to score in overtime periods that in the regular game time.
+This means that elite basketball players cope really well with any occurance of fatigue regarding their shot accuracy on a regular basis.
+As soon as the game reaches an intensity which is beyond the regular level effects of fatigue can be found leading to lower shot accuracy.
+
+The effects of psychological stress are highly visible as shown in the graph above.
+While the accuracy in the first 10 seconds of the shot clock is exceeding the overall accuracy in the regular game time (60-45% vs 45% in regular game time), it declines quite fast in the last 5 seconds of the shot clock (from 42% to 32%), leading to the rejection of the Null hypothesis, that there are no differences in the accuracy for more stressful events.
+These results imply that as the shot clock runs out, so as the psychological pressure on the shooter increases, their shot performance suffers.
+
+An ad hoc data analysis was undertaken to investigate whether this pattern of stress-inducted accuracy declination can be found as well for "superstar" players.
+The players were Lebron James, James Harden and Russel Westbrook, chosen because they were the top 3 shooters in the season.
+Their combined average accuracy per second can be seen in the graphic below.
+
+```{r}
+
+graph2 <- ggplot(NULL) + geom_smooth(data = df.acc.secondsplayers3, aes(x = seconds.left, y = acc.top, color = "Top Players")) + geom_smooth(data = acc, aes(x = seconds.left, y = accuracy, color = "Average"))
+graph2 + labs(title = "Comparison - Accuracy per second on shot clock left for Top Players vs Overall Average") + xlab("Seconds left on shot clock") + ylab("Accuracy")
+
+```
+
+The graph shows many similarities for top players to the average graph.
+The most visual differences are the slightly lower accuracy around the 15 seconds mark for the top players and the much lower decrease below 5 seconds left.
+Explanations for these differences could be that around the 15 second mark top players feel more obligated to shoot than the average player, even when they are not in the optimal position to do so - leading to lower accuracies.
+Further it seems like Top players dont get affected by last seconds stress as much as average players.
+They keep up a quite high accuracy even for the last second of the shot clock (40%).
+However, as there are many similarities for the accuracy of Top players and average players there must be other things than the accuracy that seperates those two groups.
+
+### Future research
+
+Future research could focus on this difference between Top players and average players.
+While this analysis revealed a difference for shot clock pressure it would be interesting to look at more events that induce stress like the margin of the game or the placement in the standings of the own and the enemy team.
+Further an investigation of the reasons that lead to the higher accuracy would be interesting.
+Do top players keep up their usual shot kinematics when under pressure while average players dont?
+Is it a different mindset that helps them to cope better with stress?
+Or do they simply have a higher stress tolerance as the result of continuous high expectations top players have to endure?
+
+Additionally, further reasons that lead to distinction between top players and average players should be investigated.
+
+###Limitations
+
+The shot clock variable had missing values for 5567 shots, equal to 4.34% of all shots on the dataset, limiting the meaningfulness of the findings in the accuracy per second part.
+
+***
+
+## Shooting Percentage Analysis
+###Introduction
+
+The aim of this analysis is to examine whether the top scorer of a team can help his team to win the match when he has a more outstanding performance in terms of both number of shooting attempts and accuracy.
+The Null hypothesis is that there is no difference regarding the likelyhood of winning if the top scorer of a team has more field goal attempts.
+The second Null hypothesis is there is no difference regarding the likelyhood of winning if the top scorer of a team has a higher shot percentage than on average.
+As this is such a niche area research a literature review did not reveal sufficient results so this analysis aims to fill this gap.
+
+###Methods
+
+The measurements are calculated in the following steps:
+
+  1. “Made” and “missed” in “FGM” column were replaced by 1 and 0 respectively.
+  2. A new subset was created so that each row contained the player’s name, match details and shooting performance for that player.
+  3. The number of attempts and field goal percentage of all players were calculated according to the shooting data. 
+  4. The players with most outstanding performances were picked out and another subset was generated for these players.
+  5. The correlation (p = .05) between the results of the matches and the performances of top scorers was investigated. 
+  6. A logistic regression with the outcome of the match as dependent variable and FGA and percentage of FG made as independent variables. 
+  7. Another logistic regression test was run in order to examine three special players: Russell Westbrook, Kawhi Leonard and Gordon Hayward.
+
+```{r, include=FALSE}
 # Replace "made" and "missed" with 1 and 0 respectively
@@
 
-# Find shooting percentage of each match
-
-```{r}
+### Find field goal percentage (FG percentage) for each match
+
+```{r, include=FALSE}
 # Reshape the dataframe
 
-game <- cast(data = df, GAME_ID1 + HOME_TEAM + AWAY_TEAM + LOCATION + player_name + W + FINAL_MARGIN~ FGM)
+shot.stats <- cast(data = df1, GAME_ID1 + HOME_TEAM + AWAY_TEAM + LOCATION  + W + player_name + FINAL_MARGIN ~ FGM)
 
@@
 
-game[, "FG.percentage"] <- NA 
-for (i in 1:nrow(game)){
-  game[i, "FG.percentage"] <- round(((game[i, "1"]/game[i, "FGA"]) * 100), digits = 2)
-}
-```
-
-# Impact of Super Star Players
-
-```{r}
-# Pick out 8 of the best players in that regular season and create a new subset. Analyse how shooting percentage of these players can influence the result of a match.
-
-super.star.list <- c("lebron james", "james harden", "russell westbrook", "stephen curry", "anthony davis", "blake griffin", "klay thompson", "lamarcus aldridge") # players picked out
-super.star <- game[game$player_name %in% super.star.list, c("player_name", "W", "LOCATION", "FINAL_MARGIN", "FGA", "FG.percentage")] # create a subset for these players
-
-# Replace "W" and "L" with 1 and 0 respectively, 
-super.star[, "W"] <- as.character(super.star[, "W"])
-for (i in 1:nrow(super.star)) {
-  if (super.star[i, "W"] == "W") super.star[i, "W"] <- 1
-  if (super.star[i, "W"] == "L") super.star[i, "W"] <- 0
-}
-super.star[, "W"] <- as.numeric(super.star[, "W"])
-
-ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
-ggplot(super.star) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W))
-ggplot(super.star) + geom_point(aes(x = FINAL_MARGIN, y = FGA, colour = player_name))
-ggplot(super.star) + geom_boxplot(aes(x = LOCATION, y = FINAL_MARGIN))
-
-# t.test(super.star$FINAL_MARGIN[super.star$LOCATION == "H"], super.star$FINAL_MARGIN[super.star$LOCATION == "A"])
+shot.stats[, "FG.percentage"] <- NA 
+for (i in 1:nrow(shot.stats)){
+  shot.stats[i, "FG.percentage"] <- round(((shot.stats[i, "1"]/shot.stats[i, "FGA"]) * 100), digits = 2)
+}
+```
+
+### Impact of Top Scorers
+
+```{r, include=FALSE}
+# Pick out 15 of the best players in that regular season and create a new subset. Analyse how shooting percentage of these players can influence the result of a match.
+
+top.scorers.list <- c("mnta ellis", "lebron james", "james harden", "russell westbrook", "stephen curry", "anthony davis", "blake griffin", "lamarcus aldridge", "dwayne wade", "carmelo anthony", "eric bledsoe", "demarcus cousins", "kawhi leonard", "gordon hayward", "john wall") # 15 players picked out from all players
+top.scorers <- shot.stats[shot.stats$player_name %in% top.scorers.list, c("player_name", "W", "LOCATION", "FINAL_MARGIN", "FGA", "FG.percentage")] # create a subset for these players
+top.scorers2 <- shot.stats[shot.stats$player_name %in% top.scorers.list, c("player_name", "W", "LOCATION", "FINAL_MARGIN", "FGA", "FG.percentage")] # create a subset for these players
+
+# Replace 1 and 0 with "W" and "L" respectively, 
+top.scorers2[, "W"] <- as.numeric(top.scorers2[, "W"])
+for (i in 1:nrow(top.scorers)) {
+  if (top.scorers2[i, "W"] == 1) top.scorers2[i, "W"] <- "W"
+  if (top.scorers2[i, "W"] == 0) top.scorers2[i, "W"] <- "L"
+}
+top.scorers2[, "W"] <- as.character(top.scorers2[, "W"])
 
 # Correlation test
-cor.test(super.star$FINAL_MARGIN, super.star$FGA)
-cor.test(super.star$FINAL_MARGIN, super.star$FG.percentage) # cor = 0.272
+cor.test(top.scorers$W, top.scorers$FGA)
+cor.test(top.scorers$W, top.scorers$FG.percentage)
 
 # Linear regression
-summary(lm(FINAL_MARGIN ~ FGA + LOCATION, data = super.star)) # Final margin ~ FGA and location
-summary(lm(FINAL_MARGIN ~ FG.percentage + LOCATION, data = super.star)) # Final margin ~ FG% and location
-summary(lm(FINAL_MARGIN ~ FG.percentage + FGA + LOCATION, data= super.star))
-
-summary(lm(FINAL_MARGIN ~ LOCATION, data = super.star)) # Final margin ~ FGA and location
-
-```
+# summary(lm(FINAL_MARGIN ~ FGA + LOCATION, data = top.scorers)) # Final margin ~ FGA and location
+# summary(lm(FINAL_MARGIN ~ FG.percentage + LOCATION, data = top.scorers)) # Final margin ~ FG% and location
+# summary(lm(FINAL_MARGIN ~ FG.percentage + FGA + LOCATION, data = top.scorers))
+
+# Logistic regression
+model <- glm(W ~ FGA + FG.percentage, data = top.scorers)
+summary(model)
+anova(model)
+
+```
+##a. Field Goal Attempts (FGA)
+
+There was no correlation found between the result of a match and the number of attempts of the player made in that match.
+The Pearson’s correlation coefficient for the relationship between the variables W and FGA was r(832) < .01, p = .794, failing to reach level of significance (p = .05).
+
+##b.  Field Goal Percentage (FG%)
+
+There was no correlation found between the result of a match and the field goal percentage of the player in that match.
+The Pearson’s correlation coefficient for the relationship between the variables W and FGA was r(832) < .01, p = .177, failing to reach level of significance (p = .05).
+
+```{r, include=FALSE}
+# Find out "special" players by t-test
+# Russell Westbrook
+t.test(top.scorers$FGA[top.scorers$player_name == "russell westbrook" & top.scorers$W == "1"], top.scorers$FGA[top.scorers$player_name == "russell westbrook" & top.scorers$W == "0"])
+t.test(top.scorers$FG.percentage[top.scorers$player_name == "russell westbrook" & top.scorers$W == "1"], top.scorers$FG.percentage[top.scorers$player_name == "russell westbrook" & top.scorers$W == "0"])
+# Gordon Heyward
+t.test(top.scorers$FGA[top.scorers$player_name == "gordon hayward" & top.scorers$W == "1"], top.scorers$FGA[top.scorers$player_name == "gordon hayward" & top.scorers$W == "0"])
+t.test(top.scorers$FG.percentage[top.scorers$player_name == "gordon hayward" & top.scorers$W == "1"], top.scorers$FG.percentage[top.scorers$player_name == "gordon hayward" & top.scorers$W == "0"])
+# Kawhi Leonard
+t.test(top.scorers$FGA[top.scorers$player_name == "kawhi leonard" & top.scorers$W == "1"], top.scorers$FGA[top.scorers$player_name == "kawhi leonard" & top.scorers$W == "0"])
+t.test(top.scorers$FG.percentage[top.scorers$player_name == "kawhi leonard" & top.scorers$W == "1"], top.scorers$FG.percentage[top.scorers$player_name == "kawhi leonard" & top.scorers$W == "0"])
+# Mnta Ellis
+t.test(top.scorers$FGA[top.scorers$player_name == "mnta ellis" & top.scorers$W == "1"], top.scorers$FGA[top.scorers$player_name == "mnta ellis" & top.scorers$W == "0"])
+t.test(top.scorers$FG.percentage[top.scorers$player_name == "mnta ellis" & top.scorers$W == "1"], top.scorers$FG.percentage[top.scorers$player_name == "mnta ellis" & top.scorers$W == "0"])
+# Lebron James
+t.test(top.scorers$FGA[top.scorers$player_name == "lebron james" & top.scorers$W == "1"], top.scorers$FGA[top.scorers$player_name == "lebron james" & top.scorers$W == "0"])
+t.test(top.scorers$FG.percentage[top.scorers$player_name == "lebron james" & top.scorers$W == "1"], top.scorers$FG.percentage[top.scorers$player_name == "lebron james" & top.scorers$W == "0"])
+
+special.list <- c("gordon hayward", "russell westbrook", "kawhi leonard")
+special <- shot.stats[shot.stats$player_name %in% special.list, c("player_name", "W", "LOCATION", "FINAL_MARGIN", "FGA", "FG.percentage")]
+special2 <- shot.stats[shot.stats$player_name %in% special.list, c("player_name", "W", "LOCATION", "FINAL_MARGIN", "FGA", "FG.percentage")]
+special2[, "W"] <- as.numeric(special2[, "W"])
+for (i in 1:nrow(special2)) {
+  if (special2[i, "W"] == 1) special2[i, "W"] <- "W"
+  if (special2[i, "W"] == 0) special2[i, "W"] <- "L"
+}
+
+ggplot(special2) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
+ggplot(special2) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W))
+
+special.model <- glm(W ~ FGA + FG.percentage, data = special)
+summary(special.model)
+anova(special.model)
+
+```
+Table xxx: The logistic regression for FGA and FG% of 15 players.
+
+   |   |Estimate | Std. Error | t value | Pr(>t)   
+|-|-|-|- | -             
+|(Intercept)  | 0.042259|   0.164816 |  0.256  |  0.798    
+|FGA   |-0.002679|   0.006468 | -0.414  |  0.679    
+|FG.percentage | 0.012329|   0.002736 |  4.506  |1.35e-05 ***
+
+Table xxx: The logistic regression for FGA and FG% of the 3 special players.
+| | Estimate | Std. Error | t value | Pr(>t)   
+|-|-|-|-| - 
+|(Intercept)   | 0.042259 |  0.164816 |  0.256  |  0.798    
+|FGA          | -0.002679 |  0.006468|  -0.414  |    0.679    
+|FG.percentage|  0.012329|   0.002736|   4.506  |1.35e-05 ***
+
+The box plots above shows the performances the three special players in terms of FGA and FG% respectively. The red (blue) boxes contain the data when the team lost (won) the match.
+
+###Discussion
+
+From the correlation tests between the match results and players’ performances, it can be noticed that the increase in number of field goal attempts from top scorers is unlikely to make any impact on the results of the game.
+The correlation coefficient between the results of the game and FG% is sightly higher but these two factors are still not correlated since r(832) < .01.
+Meanwhile, the p-values of both correlation tests are too higher to reach the default level of significance (p = .05).
+
+From the logistic regression test for FGA and FG%.
+It was found that FGA did not have any impact on the result of the match.
+FG% had significant correlation with the result of match.
+However, the anova test showed that the deviances of these two factors were both quite large and thus further investigations must be done in order to improve this model.
+
+There was no correlation found between field goal attempts of top scorers and game results.
+A possible explanation is that it is easier for the opponents to focus on a certain player on defensive side.
+Thus, over-using the scoring ability of a leading player is not likely to be a good approach if the team wants to increase the likelyhood of winning a game.
+Meanwhile, the accuracy of these players has a stronger impact on the outcome of the game.
+This is due to fact that once a player is being efficient on the offensive side, that player can help the team by not only keeping shooting but also creating more opportunities for his teammates.
+
+Based on results above, an additional research has been done regarding to the performances of three special players: Russell Westbrook, Kawhi Leonard and Gordon Hayward.
+These three players were chosen because their teams are more likely to take the victory when they score more in single game according to the bar plot and t-test.
+
+For these three players, the p-values still did not reach the level of significance (p = .05) for FGA.
+However, all p-values are smaller than 0.05 when their FG%s were tested.
+Based on these findings, another logistic regression was done with these three players.
+It was shown that p < .001 and residual deviance from anova dropped by 4.498 which is about 12.3% when FG% was considered.
+In comparison, the deviance was above 200 when 15 players were tested.
+
+Therefore, these three players are somehow different from others as they can contribute more to the win rates of their teams by their increasing shooting accuracy.
+There could be many reasons behind this, including their team compositions and the roles of the players.
+Further research has to be done to if one wants to find out the differences of these players and others.
+
+###Limitations
+
+The data set only contains 904 out of 1230 matches in the entire regular season.
+So there might be a chance that the highlight performances of certain players are missed out in our analysis.
+Meanwhile, it would be ideal if the contributions of all players can be analysed in more aspects including number of rebounds and assists.
+
+***
 
@@
 ####Introduction  
-This analysis would calculate whether there was any significant effect from closest defender distance from a shooter with the shot outcome.
-The hypothesis is the further a defender from the shooter, then it will be more likely to increase the chance of scoring.
-This is because the shooter is practically in "free" condition without any physical interference.
+This analysis focuses on whether there is a significant effect from closest defender distance to the shot outcome.
+The analysis is aimed to establish whether being away from the defender, i.e.
+having more time to prepare and execute a shot enables the shooter to score with a higher probability.
+Opposite effect will be when defender is close to the shooter it doesn't leave the shooter enough time to prepare and execute the shot properly.
+
+The Null Hypothesis is that the distance to a closest defender does not have an impact on accuracy of shooting.
 
@@
 The steps to calculate this measurement are as follows:  
-  1. Assign shot result to determine whether the shot is succeed (1) or missed (0)  
-  2. Select distinct of **all** closest defender distance  
+  1. Assign shot result to determine whether the shot is scored (1) or missed (0)  
+  2. Select and analyse subset of **all** closest defender distance values  
   3. Count the occurence fOr every distance  
-  4. Calculate how many succesfull shot scored from every distance point  
-  5. Calculate accuracy by dividing successfull shot scored with distance occurence for every defender distance point  
-  6. Plot the result
-```{r}
+  4. Calculate a number of uccessful shots for every distance point  
+  5. Calculate accuracy ratio for every defender distance value  
+  6. Plot and discuss the result
+```{r, include=FALSE}
 #Calculate accuracy against defender distance
 #Group the defender distance level
-distance_level <- as.data.frame(table(factor(ClosestDefenderDistance)))
+distance_level <- as.data.frame(table(factor(df[, "CLOSE_DEF_DIST"])))
 head(distance_level)
@@
 distance_level$Accuracy <- round(distance_level$SCORED / distance_level$Freq, 4)
-
+```
+
+```{r, echo=FALSE}
 #Plotting the result
@@
   scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
-  scale_x_discrete("Closest Defender Distance", breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24)) +
-  geom_smooth() +
+  scale_x_discrete("Closest Defender Distance", limits = 0:24) +
   ylab("Shooting Accuracy (%)") +
   ggtitle("Closest Defender Distance vs Shooting Accuracy")
-
-```
+```
+
 ###Plot Result
   The explanation from the plot above are as follows:   
-  1. If the defender stands really close to the shooter (0 to 2 feet), the accuracy varied between **42%** to **50%**  
-  2. If the defender distance range from 2 to 4 feet, the accuracy rate for this distance range from **44%** to **49%**  
-  3. If the defender distance range from 4 to 6 feet, the accuracy rate for this distance range from **41%** to **46%**
-  4. If the defender distance range from 6 to 8 feet, the accuracy rate for this distance range from **37%** to **42%**
-  5. If the defender distance range from 8 to 10 feet, the accuracy rate for this distance range from **37%** to **54%**
-  6. For defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from **33%** to **60%**.  
-  7. For the distance greater than 16 feet, the data varies between **0%** (no shots scored from that distance) to **100%** (certain successfull shot from that distance)
-
+  1. Defender distance range from 0 to 2 feet, the accuracy varied between **42%** to **50%**  
+  2. Defender distance range from 2 to 4 feet, the accuracy rate for this distance range from **44%** to **49%**  
+  3. Defender distance range from 4 to 6 feet, the accuracy rate for this distance range from **41%** to **46%**
+  4. Defender distance range from 6 to 8 feet, the accuracy rate for this distance range from **37%** to **42%**
+  5. Defender distance range from 8 to 10 feet, the accuracy rate for this distance range from **37%** to **54%**
+  6. Defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from **33%** to **60%**.  
+  7. Distance greater than 16 feet, the data varies between **0%** (no shots scored from that distance) to **100%** (certain successfull shot from that distance)
+
+###Correlation Test
   Next analysis will be calculated from correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:
-```{r}
+
+```{r, include=FALSE}
 #change distance level to numeric type for correlation test function
@@
 dfZerotoAll <- data.frame(distanceZerotoAll, accuracyDist1)
-
+```
+
+```{r, echo=FALSE}
 #calculate correlation
-cor.test(dfZerotoFive$distanceZerotoAll, dfZerotoFive$accuracyDist1)
-```
+cor.test(dfZerotoAll$distanceZerotoAll, dfZerotoAll$accuracyDist1)
+```
+
 ###Correlation Test Result  
-  The correlation test result produced slightly high number of correlation between defender distance with shooting accuracy for all shot distance attempt in the last seasons's NBA (0.59).
+  The correlation test result produced slightly higher number of correlation between defender distance with shooting accuracy for all shot distance attempts (0.59).
 If the findings made only based on this number only, there would be obvious that further defender distance will **automatically** increase shooting accuracy.
-However, after the analysis went into smaller piece of distance, ther result will be difference.
+However, when analysing distance ranges the result was somewhat different.
 For example, for distance range between 0 (zero) to 8 feet, the correlation measurement will be as follows:
 
-```{r}
+```{r, include=FALSE}
 #populate data from data frame
@@
 dfZerotoEight <- data.frame(distanceZerotoEight, accuracyDist2)
-
+```
+
+```{r, echo=FALSE}
 #calculate correlation
@@
 ```
-  As the analysis shown, the correlation measurement between 0 (zero) to 8 feet would be negative (-0.40). This result was different with the previous calculation done for all data points (0.59).
+
+  The correlation measurement between 0 (zero) to 8 feet would be negative (-0.40). This result was different with the previous calculation done for all data points (0.59).
 
 ###Discussion  
-  As the analysis shown, the accuracy rate of every player in the last season was somehow **affected** by the defender distance for certain range.
-For defender distance starting from 0 (zero) to 8 (eight) feet For the distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance).
-It is also shown from the dataset that only small attempt of shots were made from further defender distance affecting the accuracy calculation result.
-However, from the correlation perspective, the result would be different if the analysis included all of the data points versus only included short distance data points.
-This was arguably because further defender distance did not implied shooter distance from the basket is nearer to improve the accuracy and also there were not many players to perform shooting attempt automatically although their opponent was standing far from the shooter
-
-###Limitations  
+  The analysis demonstrated that the accuracy rate for every player was **affected** by the defender distance for certain ranges.
+For defender distance starting from 0 (zero) to 8 (eight) feet, the shooting accuracy had negative correlations.
+For the distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance), which means that this part of the analysis should not be taken into account as the results were insignificant.
+It also shown from the dataset that only small attempt of shots were made from further defender distance.
+Therefore, it would affect the accuracy calculation result.
+However, from the correlation perspective, the result would be different if the analysis included all of the data points versus only short distance data points.
+This was arguably due to further defender distance did not mean the shooter distance from the basket was closer.
+Therefore, we **reject** the Null Hypothesis, since the defender distance affected shooting accuracy.
+
+###Future Research
   As the dataset did not hold many data points for more than 20 feet defender distance, the accuracy analysis for these distances are limited.
-The analysis could be enhanced further by including additional season data.
-That way, there will be more data point for further distance and possibly would produce stronger pattern as the short distance shown.
+The future analysis could be enhanced further by including data for additional seasons.
 
@@
 ####Introduction  
-This analysis would calculate whether was there any significant effect from shooting distance from basket with the shot outcome.
-The hypothesis is the further a shooter from the basket, then the success rate would be decreased.
-This was because when the shooter was practically in closer shooting range, the shooter shooting's result would be more accurate.
+This analysis focuses on whether there is any significant effect resulting from shooting distance from basket on the shot outcome.
+
+The NULL hypothesis is that the distance to the basket does not influence outcome of a shot.
 
@@
 The steps to calculate this measurement are as follows:  
-  1. Assign shot result to determine whether the shot is succeed (1) or missed (0)  
-  2. Select distinct of **all** shooting distance  
-  3. Count the occurence fOr every distance  
-  4. Calculate how many succesfull shot scored from every distance point  
-  5. Calculate accuracy by dividing successfull shot scored with particular distance occurence for every shooting distance point  
-  6. Plot the result
-```{r}
+  1. Assign shot result to determine whether the shot is scored (1) or missed (0)  
+  2. Select subsets of **all** shooting distances from the basket  
+  3. Count the occurence fOr every distance to basket value  
+  4. Allocate shot success rate to every distance from basket value  
+  5. Plot and discuss the result
+```{r, include=FALSE}
 #Calculate accuracy against shooting distance
 #Group the shooting distance level
-shooting_distance_level <- as.data.frame(table(factor(ShotDistance)))
+shooting_distance_level <- as.data.frame(table(factor(df[, "SHOT_DIST"])))
 head(shooting_distance_level)
@@
 shooting_distance_level$Accuracy <- round(shooting_distance_level$SCORED / shooting_distance_level$Freq, 4)
-
+```
+
+```{r, echo=FALSE}
 #Plotting the result
@@
   scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
-  scale_x_discrete("Shooting Distance", breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24, 26, 28, 30, 32, 34, 36, 38, 40, 42, 44, 46, 48)) +
+  scale_x_discrete("Shooting Distance", limits = 0:43) + 
   geom_smooth() + 
@@
   ggtitle("Shooting Distance vs Shooting Accuracy")
-
 ```
@@
 ###Plot Result
-  The explanation from the plot above are as follows:   
-  1.  If the shooter stands really close to the basket (0 to 2 feet), the accuracy varied between **62%** to **68%**  
-  2.  If the shooting distance range from 2 to 4 feet, the accuracy rate for this distance range from **61%** to **67%**  
-  3.  If the shooting distance range from 4 to 6 feet, the accuracy rate for this distance range from **44%** to **58%**  
-  4.  If the shooting distance range from 6 to 8 feet, the accuracy rate for this distance range from **38%** to **51%**  
-  5.  If the shooting distance range from 8 to 10 feet, the accuracy rate for this distance range from **35%** to **42%**  
-  6.  If the shooting distance range from 10 to 12 feet, the accuracy rate for this distance range from **34%** to **42%**  
-  7.  If the shooting distance range from 12 to 14 feet, the accuracy rate for this distance range from **37%** to **47%**  
-  8.  If the shooting distance range from 14 to 16 feet, the accuracy rate for this distance range from **33%** to **51%**  
-  9.  If the shooting distance range from 16 to 18 feet, the accuracy rate for this distance range from **37%** to **48%**  
-  10. If the shooting distance range from 18 to 24 feet, the accuracy rate for this distance range from **33%** to **43%**  
-  11. If the shooting distance range from 24 to 28 feet, the accuracy rate for this distance range from **18%** to **38%**  
-  12. For shooting distance more than 28 feet, the accuracy varies greater than previous distance range, starting from **0%** (no shots scored from that distance) to **50%**
-
-  Next analysis will be calculated from correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:  
-```{r}
+  The explanation from the plot above are as follows. The shooter distance ranges and accuracy:   
+  1.  0 to 2 feet: the accuracy varied between **62%** to **68%**  
+  2.  2 to 4 feet: the accuracy rate for this distance range from **61%** to **67%**  
+  3.  4 to 6 feet: the accuracy rate for this distance range from **44%** to **58%**  
+  4.  6 to 8 feet: the accuracy rate for this distance range from **38%** to **51%**  
+  5.  8 to 10 feet: the accuracy rate for this distance range from **35%** to **42%**  
+  6.  10 to 12 feet: the accuracy rate for this distance range from **34%** to **42%**  
+  7.  12 to 14 feet: the accuracy rate for this distance range from **37%** to **47%**  
+  8.  14 to 16 feet: the accuracy rate for this distance range from **33%** to **51%**  
+  9.  16 to 18 feet: the accuracy rate for this distance range from **37%** to **48%**  
+  10. 18 to 24 feet: the accuracy rate for this distance range from **33%** to **43%**  
+  11. 24 to 28 feet: the accuracy rate for this distance range from **18%** to **38%**  
+  12. more than 28 feet: the accuracy varies greater than previous distance range, starting from **0%** (no shots scored from that distance) to **50%**
+
+  Analysis has also produced a correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:  
+```{r, include=FALSE}
 #change distance level to numeric type for correlation test function
@@
 dfShootingZerotoAll <- data.frame(shootingDistanceZerotoAll, shootingAccuracyDist1)
+```
+
+```{r, echo=FALSE}
 cor.test(dfShootingZerotoAll$shootingDistanceZerotoAll, dfShootingZerotoAll$shootingAccuracyDist1)
 ```
+
 ###Correlation Test Result  
   The correlation test result shown strong negative correlation between shooting distance and shooting accuracy (-0.87).
-This meant that for all data points, the further a shooter from the basket, it was unlikely the shooter's attempt was success, regardless who was the shooter.
-However, if the correlation test taken for certain range distance based on the diagram, for example range from 10 feet to 20 feet, the correlation result would be different as follows:
-
-```{r}
-#populate data for shooting distance in range 10 to 24 feet
+This meant that for all data points, the further a shooter from the basket, the shooter's accuracy would decrease.
+However, if the correlation test taken for certain range distance based on the diagram, the correlation result would be different as follows:
+
+```{r, include=FALSE}
+#populate data for shooting distance in range 10 to 20 feet
 shootingDistance10to20 <- c()
@@
 dfShooting10to20 <- data.frame(shootingDistance10to20, shootingAccuracyDist2)
+```
+
+```{r, echo=FALSE}
 cor.test(dfShooting10to20$shootingDistance10to20, dfShooting10to20$shootingAccuracyDist2)
 ```
+
   The second correlation test produced a small positive correlation between range 10 to 20 feet (0.11). This result was different with previous measurement (-0.87)
@@
 ###Discussion  
-  Based on the diagram and correlation, it can be concluded that although the shooting distance would directly affect shooting accuracy, but the relationship between these variables are not perfectly linear.
-This was shown in the diagram where there was a relatively **similar** shooting accuracy for shots taken from 10 feet until 20 feet, regardless the shooter's performance.
-And for the success rate in longer distances, it could be arguably inferred as the shooters tried to attempt in closing game time (buzzer beater) when the shot clock or game clock nearly reach zero.
-
-###Limitations  
- Since the dataset mostly return 0% for shooting distance greater than 30 feet, it would better if the dataset could be enriched with previous season statistics, or even the last 10 seasons to exercise in more detail about the pattern for long distance shots attempt.
-```{r}
-#Summary of total points
-ptssum<- cast(df1, GAME_ID1+ player_id + player_name  ~ PTS); head(ptssum)
-#creating "total points per player, per game" column
-ptssum$totalpts<- ptssum[,"2"]*2 + ptssum[,"3"]*3
-
-pttsum2<-ptssum[order(ptssum$totalpts,decreasing=TRUE),]
-head(pttsum2,10)
-```
+  Based on the diagram and correlation above, it could be concluded that although the shooting distance directly affects shooting accuracy, the relationship between these variables are not perfectly linear.
+This was shown in the diagram where there is a relatively **similar** shooting accuracy for shots taken from 10 feet to 20 feet, regardless the shooter's performance.
+The success rate for longer distances, it could be arguably inferred that the shooters tried to attempt a shot in closing game time (buzzer beater) when the shot clock or game clock nearly reach zero- hence a longer distances.
+Therefore, the analysis would reject the  Null Hypothesis.
+
+###Future Research
+ The dataset mostly returns 0% for shooting distance greater than 30 feet. Therefore the future research could be enriched with statistics from other seasons to see if these findings hold.
+
+***
 
@@
 
-```{r}
+```{r, include=FALSE}
 
@@
 
+#Summary of total points
+ptssum<- cast(df1, GAME_ID1+ player_id + player_name  ~ PTS); head(ptssum)
+#creating "total points per player, per game" column
+ptssum$totalpts<- ptssum[,"2"]*2 + ptssum[,"3"]*3
+
+pttsum2<-ptssum[order(ptssum$totalpts,decreasing=TRUE),]
+
 #Set a new data frame for the 3 players with best performances(higher points)
-best_players_df <- head(pttsum2,3)
+best_players_df <- head(pttsum2, 3)
 
@@
 player_and_game <- function(x, y) {
-   player <- df[df[ ,'player_id'] == x & df[ ,'GAME_ID1'] == y, 'FGM']
+   player <- df[df[, 'player_id'] == x & df[,'GAME_ID1'] == y, 'FGM']
    number_of_streaks(as.vector(player))
@@
 for(i in best_players_df$player_id) {
-   streak_per_player[[j]] <- player_and_game(i, best_players_df[ourtable[ ,'player_id'] == i, 'GAME_ID1'])
+   streak_per_player[[j]] <- player_and_game(i, best_players_df[best_players_df[, 'player_id'] == i, 'GAME_ID1'])
    j <- j + 1
@@
 ###Barplots
-```{r}
+```{r, include=FALSE}
 
@@
 player3_df <- data.frame('streaks' = streak_per_player[[3]])
-best_players_df
+
 #Show distribution for each player's streak lengths
@@
 
+####Barplot result
   The distribution of the three players is unimodal and left skewed. There are shown some extremely long shooting streaks made(i.e. Klay Thompson with 13 shots in a row).
@@
 ###Boxplots
-####Boxplot for Kyrie Irnving
-```{r}
+
+```{r, include=FALSE}
 #Create a boxplot for each player's streak lengths 
-boxplot(player1_df$streaks) 
-summary(player1_df$streaks) #Check that boxplot is valid
-IQR(player1_df$streaks) #Check that boxplot is valid
-```
+attach(mtcars)
+par(mfrow=c(1, 3))
+boxplot(player1_df$streaks, main = 'Kyrie Irnving')
+boxplot(player2_df$streaks, main = 'Mo Williams')
+boxplot(player3_df$streaks, main = 'Klay Thompson')
+```
+
+```{r, include=FALSE}
+#Kyrie Irnving
+summary(player1_df$streaks)
+IQR(player1_df$streaks) 
+```
+#### Outcome for Kyrie Irnving:
 
@@
 
-####Boxplot for Mo Williams
-```{r}
-boxplot(player2_df$streaks)
+```{r, include=FALSE}
+#Mo Williams
 summary(player2_df$streaks)
@@
 ```
+
+#### Outcome for Mo Williams
 
@@
 
-####Boxplot for Klay Thompson
-```{r}
-boxplot(player3_df$streaks)
+```{r, include=FALSE}
+#Klay Thompson
 summary(player3_df$streaks)
@@
 
+#### Outcome for Klay Thompson
+
   1. The typical length of a streak is 8 (median = 8)
   2. The Interquartile Range is 5
-  3. Streak length of 13 is unsually high compared to the rest of the distribution
+  3. There is a streak of length 13 (max = 13)
 
@@
 
-```{r}
+```{r, include=FALSE}
 #Find the shooting percentage for each of the 3 players
@@
 for(i in best_players_df$player_id) {
-  shooting_percentage[j] <- sum(df[df[ ,'player_id'] == i, 'FGM'])/ nrow(df[df[ ,'player_id'] == i, ])
-  j <- j+1
+  shooting_percentage[j] <- sum(df[df[, 'player_id'] == i, 'FGM']) / nrow(df[df[, 'player_id'] == i, ])
+  j <- j + 1
 }
@@
 ###Kyrie Irving
-```{r}
+```{r, include=FALSE}
 ###PLAYER 1 VS SAMPLE PLAYER###
@@
 simul_shooter <- as.list(simul_shooter)
-for(i in 1:length(simul_shooter)) {
-   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
-   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0
+for (i in 1:length(simul_shooter)) {
+   if (simul_shooter[i] == 'made') {
+     simul_shooter[i] <- 1
+   } else if (simul_shooter[i] == 'missed') {
+     simul_shooter[i] <- 0
+   }
 }
@@
 #If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
-g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar() + labs(title = 'sample shooter')
-g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar() + labs(title = 'Kyrie Irving')
-grid.arrange(g1, g2, ncol = 2
-```
-
-  The streak length distributions are quite similar, so we can conclude that Kyrie Irving doesn't have a hot hand.
+g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar(fill = 'blue') + labs(title = 'sample shooter')
+g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar(fill = 'red') + labs(title = 'Kyrie Irving')
+grid.arrange(g1, g2, ncol = 2)
+```
+
+####Plot Result
+  After running many times the sample shooter, there are times that the distributions of sample and real shooter are quite similar.
+As a consequence, it can be concluded that Mo Williams does not have a hot hand.
 
 ### Mo Williams
-```{r}
+```{r, include=FALSE}
 ###PLAYER 2 VS SAMPLE PLAYER###
@@
 shot_results <- c('made', 'missed')
-simul_shooter <- sample(shot_results, size = length(streak_per_player[[2]]), replace = TRUE, prob = c(shooting_percentage[[2]], 1 - shooting_percentage[[2]]))
+simul_shooter <- sample(shot_results, size = length(streak_per_player[[2]]), replace = TRUE, 
+                        prob = c(shooting_percentage[[2]], 1 - shooting_percentage[[2]]))
 simul_shooter
@@
 #If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
-g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar() + labs(title = 'sample shooter')
-g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar() + labs(title = 'Mo Williams')
+g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar(fill = 'blue') + labs(title = 'sample shooter')
+g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar(fill = 'green') + labs(title = 'Mo Williams')
 grid.arrange(g1, g2, ncol = 2)
@@
 
-  The streak length distributions are quite similar, so we can conclude that Mo Williams doesn't have a hot hand.
+####Plot Result
+
+  After running many times the sample shooter, there are times that the distributions of sample and real shooter are quite similar.
+As a consequence, it can be concluded that Mo Williams does not have a hot hand.
 
 ###Klay Thompson
-```{r}
+```{r, include=FALSE}
 ###PLAYER 3 VS SAMPLE PLAYER###
@@
 shot_results <- c('made', 'missed')
-simul_shooter <- sample(shot_results, size = length(streak_per_player[[3]]), replace = TRUE, prob = c(shooting_percentage[[3]], 1 - shooting_percentage[[3]]))
+simul_shooter <- sample(shot_results, size = length(streak_per_player[[3]]), replace = TRUE, 
+                        prob = c(shooting_percentage[[3]], 1 - shooting_percentage[[3]]))
 simul_shooter
@@
 #If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
-g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar() + labs(title = 'sample shooter')
-g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar() + labs(title = 'Klay Thompson')
+g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar(fill = 'blue') + labs(title = 'sample shooter')
+g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar(fill = 'brown') + labs(title = 'Klay Thompson')
 grid.arrange(g1, g2, ncol = 2)
@@
 
+####Plot Result
+
   Klay Thompson is an exception making 13 shots in a row.
-There is no such distribution in our sample, so it coulb be assumed that he is likely to have a hot hand.
+There is no such distribution in the sample, so it could be assumed that he is likely to have a hot hand.
 Further statistical analysis (logistic regression) should be done in order to reach a conclusion.
 
-```{r}
+```{r, include=FALSE}
 #Regression Analysis
 
-df1$GAME_CLOCK <-as.numeric(df1$GAME_CLOCK)
+df1$GAME_CLOCK <- as.numeric(df1$GAME_CLOCK)
 
@@
 
-linreg_fgm <-lm(FGM ~ LOCATION+W+FINAL_MARGIN+SHOT_NUMBER+ PERIOD+GAME_CLOCK+  SHOT_CLOCK+  DRIBBLES+ TOUCH_TIME+SHOT_DIST+PTS_TYPE+ CLOSEST_DEFENDER_PLAYER_ID + CLOSE_DEF_DIST+ player_id+ WINNER, df1 )
+linreg_fgm <- lm(FGM ~ LOCATION + W + FINAL_MARGIN + SHOT_NUMBER + PERIOD + GAME_CLOCK + SHOT_CLOCK+ DRIBBLES + TOUCH_TIME + SHOT_DIST + PTS_TYPE + CLOSEST_DEFENDER_PLAYER_ID + CLOSE_DEF_DIST + player_id + WINNER, df1)
 
@@
 
-linreg_fgm2 <-lm(FGM ~ FINAL_MARGIN+SHOT_NUMBER+ PERIOD +  SHOT_CLOCK+  DRIBBLES+ TOUCH_TIME+SHOT_DIST+PTS_TYPE+ CLOSE_DEF_DIST+ player_id, df1 )
+linreg_fgm2 <- lm(FGM ~ FINAL_MARGIN + SHOT_NUMBER + PERIOD + SHOT_CLOCK + DRIBBLES + TOUCH_TIME + SHOT_DIST + PTS_TYPE + CLOSE_DEF_DIST + player_id, df1 )
 
@@
 
-##Discussion  
-###What results did you find?
-
-###Why is this interesting?
-
-###What would be the next steps?
----
+#Conclusion
+
+#References
+
+  Ahart (1973). In Kendall, P. and Hollon, S. (1979). Cognitive-behavioral interventions. New York: Academic Press.
+  Erculj, F.
+and Supej, M.
+(2009).
+Impact of Fatigue on the Position of the Release Arm and Shoulder Girdle over a Longer Shooting Distance for an Elite Basketball Player.
+Journal of Strength and Conditioning Research, 23(3), pp.1029-1036.
+  Mascret, N., Ibáñez-Gijón, J., Bréjard, V., Buekers, M., Casanova, R., Marqueste, T., Montagne, G., Rao, G., Roux, Y.
+and Cury, F.
+(2016).
+The Influence of the ‘Trier Social Stress Test’ on Free Throw Performance in Basketball: An Interdisciplinary Study.
+PLOS ONE, 11(6), p.e0157215.
+
+ECONOMIST. As sweet as ever (2015) Available at: http://www.economist.com/blogs/gametheory/2015/06/home-advantage-basketball (Accessed: 16 October 2016).

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
debugs

@@
 ```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo=TRUE)
 knitr::opts_chunk$set(echo = TRUE)
 ```
-Report produced by: Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo  
-Date submitted: 16/10/2016
+##Report produced by: Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo  
+##Date submitted: 16/10/2016
 
 ***
-
-#Report structure
 
@@
 The chosen dataset for this report is a summary of shots made during NBA season 2014-2015. The following areas of research have been selected:
+
+** TO GET UPDATED AS A FINAL STEP**
 
@@
 
- The data has been acquired from Kaggle website **(https://www.kaggle.com/dansbecker/nba-shot-logs)** on 30 September 2016.
+ The data has been acquired from [Kaggle](https://www.kaggle.com/dansbecker/nba-shot-logs) website on 30 September 2016.
 
@@
 
-***
-
-##Analysis
+#Analysis
 
@@
 
-***
-
 ## Home Advantage analysis
@@
 This area of analysis focuses on investigation of whether playing at 'home' gives the team advantage comparing to when playing 'away'.
-Esssentially, this analysis focuses on effect of the game lovation (home/away) on winning ratio.
+Esssentially, this analysis focuses on effect of the game location (home/away) on winning ratio.
 
 The NULL Hypothesis: Teams playing at home is more likely to win then when playing away. 
+The NULL Hypothesis: Analysis of means for home and away wins does not show any significant difference
 
@@
     4. Plot the team level data 
-    5. Discuss the findings of the analysis
+    5. Perform a t-test (p .05) to examine whether there is a difference in the means between teams winning at home and teams winning away
+    6. Linear regression with Final margin as dependent variable and Location as independent variable to examine whether playing at home results predicts a higher final margin
+
+The graph below shows descriptive statistics for home/away wins:
 
@@
 #preparing a dataframe for home advantage analysis
-aa <- cast(df1, GAME_ID1 + HOME_TEAM + AWAY_TEAM + WINNER ~ W)
+advdf <- cast(df1, GAME_ID1 + HOME_TEAM + AWAY_TEAM + WINNER ~ W)
 
 #table showing a number of games won by home and away teams in the whole season
-table(aa$WINNER)
+table(advdf$WINNER)
 
 #home wins:
-hh <- sum(aa$WINNER == "HOME")
+hh <- sum(advdf$WINNER == "HOME")
 #Away wins:
-aw <- sum(aa$WINNER == "AWAY")
+aw <- sum(advdf$WINNER == "AWAY")
 
@@
 round(aw / (hh + aw), 4)
-
 ```
@@
 #preparing a dataframe for analysis of games when a team played at home 
-aa1 <- cast(aa, HOME_TEAM ~ WINNER); names(aa1)[1] <- "TEAM"
-aa1$WIN_PCT_HOME <- round(aa1$HOME / (aa1$HOME + aa1$AWAY), 2)
+advdf1 <- cast(advdf, HOME_TEAM ~ WINNER); names(advdf1)[1] <- "TEAM"
+advdf1$WIN_PCT_HOME <- round(advdf1$HOME / (advdf1$HOME + advdf1$AWAY), 2)
 
 #preparing a dataframe for analysis of games when a team played at away 
-aa2 <- cast(aa, AWAY_TEAM ~ WINNER); names(aa2)[1] <- "TEAM"
-aa2$WIN_PCT_AWAY <- round(aa2$AWAY / (aa2$HOME + aa2$AWAY), 2)
+advdf2 <- cast(advdf, AWAY_TEAM ~ WINNER); names(advdf2)[1] <- "TEAM"
+advdf2$WIN_PCT_AWAY <- round(advdf2$AWAY / (advdf2$HOME + advdf2$AWAY), 2)
 
 #merging two dataframes together to enable the effective analysis of win ratios for teams playing at home and away.
-aa3 <- merge (aa1, aa2, by ="TEAM")
+advdf3 <- merge (advdf1, advdf2, by ="TEAM")
 
 #ordering the dataframe by win percentage at home.
-aa3 <- aa3[order(aa3$WIN_PCT_HOME, decreasing = TRUE),]
+advdf3 <- advdf3[order(advdf3$WIN_PCT_HOME, decreasing = TRUE),]
 
 #ordering factors of the dataframe by win percentage at home.
-aa3$TEAM <- factor(aa3$TEAM, levels = aa3$TEAM[order(-aa3$WIN_PCT_HOME)])
+advdf3$TEAM <- factor(advdf3$TEAM, levels = advdf3$TEAM[order(-advdf3$WIN_PCT_HOME)])
 
 #creating a column which shows a difference between home and away winning percentage rate
-aa3$WIN_PCT_DIFF<- aa3$WIN_PCT_HOME - aa3$WIN_PCT_AWAY
-
-```
+advdf3$WIN_PCT_DIFF<- advdf3$WIN_PCT_HOME - advdf3$WIN_PCT_AWAY
+
+#t tests
+
+t.test(advdf3$AWAY.y, advdf3$HOME.x)
+```
+
+```{r, echo=FALSE}
+#plot showing home and away win ratio per team
+ggplot(advdf, aes(x = advdf$WINNER)) + geom_bar(aes()) + ylab("game count") + xlab("winner")
+```
+
+###Results
 
 ###Findings
-Table 1: Proportion of total games split by Location win factor
-
-X     | HOME WIN | AWAY WIN
+**Table 1**: Proportion of total games split by Location win factor
+
+X     | Home Win | Away Win
 ----- | -------- | ---
@@
 
-Tables 2 & 3: Winning ratio of teams playing at home vs games playing away
-
-TEAM        | GSW | ATL | POR | MEM | SAS | CLE | HOU | LAC | OKC | DAL | NOP | WAS | TOR | CHI | MIL  
+**Tables 2 & 3**: Winning ratio of teams playing at home vs games playing away
+
+Team     | GSW | ATL | POR | MEM | SAS | CLE | HOU | LAC | OKC | DAL | NOP | WAS | TOR | CHI | MIL  
 ------------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----
@@
 
-TEAM        | PHX | IND | BOS | CHA | UTA | DEN | BKN | SAC | MIA | DET | LAL | ORL | PHI | NYK |MI 
+Team     | PHX | IND | BOS | CHA | UTA | DEN | BKN | SAC | MIA | DET | LAL | ORL | PHI | NYK |MI 
 ------------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----
@@
 
-```{r}
+```{r, echo=FALSE}
 #plot showing home and away win ratio per team
-ggplot(aa3, aes(x = TEAM)) + geom_point(aes(y = WIN_PCT_HOME, col = "HOME")) + geom_point(aes(y = WIN_PCT_AWAY, col = "AWAY")) 
+ggplot(advdf3, aes(x = TEAM)) + geom_point(aes(y = WIN_PCT_HOME, col = "wins at HOME")) + geom_point(aes(y = WIN_PCT_AWAY, col = "wins  AWAY")) + ylab("Win ratio") + xlab("Team Names")
 ```
@@
 
-###Variance test Analysis
-```{r}
-
-```
-**JUSTING TO SEE IF THERE IS ANYTHING WE CAN INCLUDE HERE
-
-###regression analysis
+The T-test used to verify the differences in the means seen in the plot reached significance, t(58) = -2.39, p = .02.
+
+###Regression analysis
 
@@
 
-Table 4: Regression Analysis
+**Table 4**: Regression analysis for the effect of location and rest days on final score margin
 
@@
 
-The analysis shown that there is a clear evidence that a team playing at home is more likely to win with an average win probability of 55.97%.
+The analysis shows that there is a clear evidence that a team playing at home is more likely to win with an average win probability of 55.97%.
 The same conclusion is evident from the analysis conducted on a team level where 25 teams are winning more games at home than away.
-The result shown that the NULL hypothesis should be rejected as a team playing at home is more likely to win then an away team.
-Regression analysis shown that playing at home gives teams on average additional 4.29 margin points.
-
-The result of this research also confirms recent trend of decline in home advantage over the years: probability of a home team winning has reduced from around 65% in 1975-1992 to average of 60.3% in 1993-2011 and to around 58.5% for 2011-15.
-Home win ratio was 57.4% win in 2014-2015 season.
-The analysis demonstrated that the home win ratio went down even further to 56.0% in 2015-2016 season.
+The results show that the NULL hypothesis should be rejected as a team playing at home is more likely to win than an away team.
+Linear regression analysis shows that playing at home gives teams on average additional 4.29 margin points.
+
+The result of this research also confirms the recent trend of decline in home advantage over the years: probability of a home team winning has reduced from around 65% in 1975-1992 to average of 60.3% in 1993-2011 and to around 58.5% for 2011-14.
+The analysis demonstrated that the home win ratio went down even further to 56.0% in 2014-2015 season.
 Based on the literature review this trend could be potentially explained by the following factors (Economist, 2015):
@@
 
+***
+
 ##Effect of rest on teams' performance
@@
 
-This area of analysis focuses on investigation of whether a number of rest days between games affect performance and more importantly winning ratio of a team.
-
-The Null Hypothesis is: By having longer rest a team is likely to have a higher winning ratio.
+This area of analysis focuses on investigation of whether a number of rest days between games affect the performance and more importantly winning ratio of a team.
+
+The NULL Hypothesis is: Teams' mean values of wins are not significantly different based on the number of rest days between games
 
 ###Methods
+
 The steps to calculate this measurement are as follows:  
     1. Calculating number of rest days between games for all teams
-    2. Calculating ratios for effects of rest days on win/loss
-    3. Calculating ratios for effects of rest days on win/loss form home/awayCalculating ratios for effects of rest days on win/loss
-    4. Discuss the findings of the data
-
-```{r, include=FALSE}
-
-#calulating number of rest days between teams
-for (i in 2: nrow (restdays2)) {
-  if (restdays2$TEAM[i]==restdays2$TEAM[i-1])
-    restdays2$rest[i]<- as.numeric(restdays2$DATE[i] - restdays2$DATE[i-1]-1)
-  else 
-    (restdays2$rest[i]<- "NA")  
-}
-restdays2$BLANK<-""
-
-restdays2$rest <- as.numeric (restdays2$rest)
-
-#regression analysis
-margreg<- lm(FINAL_MARGIN ~ LOCATION + rest, restdays2)
-summary(margreg)
-
-#creating dataframe for total rest days
-restdays4<- cast (restdays2, rest ~  W)
-colnames(restdays4)<- c( "rest", "L","W")
-
-#creating a column for total restdays for a team when lost and won to enable calculation of ratios
-restdays4$total<-restdays4$L + restdays4$W
-restdays4$dist <-   round (restdays4$total * 100 /sum (restdays4$total) ,2)
-
-#plotting distribution of rest days to establish which of the rows should be analysed
-restdays4[,c(1,4,5)]
-#The distribution of rest days show that 92% of games are covered by 0,1 or 2 days of rest and therefore further analysis will only focus on these 3 sections
-
-restdays4$ L.pct <-   round(restdays4$L*100/(restdays4$L+restdays4$W),2) 
-restdays4$ W.pct <-   round(restdays4$W*100/(restdays4$L+restdays4$W),2)
-
-print (restdays4[1:3,c(1,3,7)])
-
-#creating a new dataframe for summary of home/away win losses based on number of rest days.
-restdays3<- cast (restdays2, rest ~ LOCATION + W) ;
-
-#renaming the columns
-colnames(restdays3)<- c( "rest", "AL","AW","HL", "HW")
-
-#caluclating ratios
-restdays3$ AL.pct <-   round(restdays3$AL*100/(restdays3$AL+restdays3$AW),2) 
-restdays3$ AW.pct <-   round(restdays3$AW*100/(restdays3$AL+restdays3$AW),2)
-restdays3$ HL.pct <-   round(restdays3$HL*100/(restdays3$HL+restdays3$HW),2)
-restdays3$ HW.pct <-   round(restdays3$HW*100/(restdays3$HL+restdays3$HW),2)
-
-print (restdays3[1:3,c(1,7,9)])
-
-#Testing of rest data
-
-#t.test(x=restdays2$W[restdays2$rest==0], y= restdays2$W[restdays2$rest==1])
-#chisq.test(x=restdays2$W[restdays2$rest==0], y= restdays2$W[restdays2$rest==1])
-
-#restdays2$W[restdays2$rest==0,]
-
-```
-
-###Findings
-
-Table 5: Distribution of number of games per rest days
+    3. Calculating ratios for effects of rest days on win/loss form home/away
+    4. Perform a Chi-square test to examine whether there is a difference in the distributon for games won and games lost for the different restdays
+    5. Calculating ratios for effects of rest days on win/loss form home/away
+    6. Run a regression analysis
+    7. Discuss the findings of the data
+
+The table below shows descriptive statistics for Rest Days:
+
+**Table 5**: Distribution of number of games per rest days
 
@@
 
-Table 6: Win Rate split by number of rest days between games
-
-  rest days| number of wins| win percentage
+```{r, include=FALSE}
+
+#calulating number of rest days between teams
+for (i in 2: nrow (restdays2)) {
+  if (restdays2$TEAM[i]==restdays2$TEAM[i-1])
+    restdays2$rest[i]<- as.numeric(restdays2$DATE[i] - restdays2$DATE[i-1]-1)
+  else 
+    (restdays2$rest[i]<- "NA")  
+}
+restdays2$BLANK<-""
+
+restdays2$rest <- as.numeric (restdays2$rest)
+
+#regression analysis
+margreg<- lm(FINAL_MARGIN ~ LOCATION + rest, restdays2)
+summary(margreg)
+
+#creating dataframe for total rest days
+restdays4<- cast (restdays2, rest ~  W)
+colnames(restdays4)<- c( "rest", "L","W")
+
+#creating a column for total restdays for a team when lost and won to enable calculation of ratios
+restdays4$total<-restdays4$L + restdays4$W
+restdays4$dist <-   round (restdays4$total * 100 /sum (restdays4$total) ,2)
+
+#plotting distribution of rest days to establish which of the rows should be analysed
+restdays4[,c(1,4,5)]
+#The distribution of rest days show that 92% of games are covered by 0,1 or 2 days of rest and therefore further analysis will only focus on these 3 sections
+
+restdays4$ L.pct <-   round(restdays4$L*100/(restdays4$L+restdays4$W),2) 
+restdays4$ W.pct <-   round(restdays4$W*100/(restdays4$L+restdays4$W),2)
+
+print (restdays4[1:3,c(1,3,7)])
+
+#creating a new dataframe for summary of home/away win losses based on number of rest days.
+restdays3<- cast (restdays2, rest ~ LOCATION + W) ;
+
+#renaming the columns
+colnames(restdays3)<- c( "rest", "AL","AW","HL", "HW")
+
+#caluclating ratios
+restdays3$ AL.pct <-   round(restdays3$AL*100/(restdays3$AL+restdays3$AW),2) 
+restdays3$ AW.pct <-   round(restdays3$AW*100/(restdays3$AL+restdays3$AW),2)
+restdays3$ HL.pct <-   round(restdays3$HL*100/(restdays3$HL+restdays3$HW),2)
+restdays3$ HW.pct <-   round(restdays3$HW*100/(restdays3$HL+restdays3$HW),2)
+
+print (restdays3[1:3,c(1,7,9)])
+
+#Testing of rest data
+qqqq<-as.matrix(sapply(restdays4, as.numeric))
+
+chisq.test(qqqq[2,], qqqq[3,])
+
+```
+
+###Findings
+
+**Table 5**: Distribution of number of games per rest days
+
+ Rest Days| Games | Distribution
+----------|-------|---
+    0  | 427 | 23.62
+    1  | 990 | 54.76
+    2  | 254 | 14.05
+    3  |  59 |  3.26
+    4  |  13 |  0.72
+    5  |   3 | 0.17
+    7  |   6 | 0.33
+    8  |  20 | 1.11
+    9  |   3 | 0.17
+    10 |   3 | 0.17
+   1st game of season  |  30|  1.66
+
+The distribution of rest days shows that 92% of games are covered by 0,1 or 2 days of rest and therefore further analysis will only focus on these 3 sections.
+
+**Table 6**: Win rate split by number of rest days between games
+
+  Rest Days| Number of Wins|Win Percentage
 -|-|-
@@
 
-Table 7: Win Rate split by number of rest days between games and location
-
-  rest days| home wins percentage | away wins percentage|
+**Table 7**: Win rate split by number of rest days between games and location
+
+  Rest Days| Home Wins Percentage | Away Wins Percentage|
 -|-|-
@@
 
+The Pearson's Chi-Square test failed to reach the level of significance, X2 (36) = 42, p= .23.
+
 ###Discussion
 
-The analysis demonstrated that overall winning ratio improves with increase in  rest days: back-to-back games had 47.07% winning ratio, 1 day rest increased winning ratio to 50.10% and with 2 days rest the winning ration has increased even further to 53.94%.
-
-By extending the analysis to the team level, the results have demonstrated similar tendency for home  games where winning ratio has increased from 53.54% to 54.68% and 59.57% respectively with 0, 1 and 2 days of rest.
-Somewhat interestingly the analysis of effect of rest days on winning ratio of away games demonstrated that there is hardly any difference between 0 and 1 days of rest (44.33% and 44.24%).
-However with 2 days of rest teams' average winning ratio have increased to 46.90% (2.66% increase).
+The analysis demonstrates that overall the winning ratio stays the same with increase in  rest days.
+However, when looking at invididual restdays the findings are different: back-to-back games have 47.07% winning ratio, 1 day rest increases winning ratio to 50.1% and with 2 days rest the winning ration increases even further to 53.94%.
+
+By extending the analysis to the team level, the results demonstrate similar tendency for home games where winning ratio increased from 53.54% to 54.68% and 59.57% respectively with 0, 1 and 2 days of rest.
+Somewhat interestingly the analysis of effect of rest days on winning ratio of away games demonstrates that there is hardly any difference between 0 and 1 days of rest (44.33% and 44.24%).
+However with 2 days of rest teams' average winning ratio increases to 46.90% (2.66% increase).
 This perhaps could be explained by the fact that teams were able to minimise effect of travelling when having a bigger window between games.
 
-Based on the above, the analysis shown that the null hypothesis should be rejected as there is a clear evidence that the higher rest days correspond to higher winning ratio
+Based on the above, the analysis shows that the null hypothesis should be rejected as there is a clear evidence that the higher rest days correspond to higher winning ratio.
 
@@
 
-To better understand effects of rest and home/away advantage the future research should focus on distances the teams have to travel between the games.
-For example if a team is travelling from one coast to another- it is likely to affect them more comparing to a team from New York 'travelling' to play an away gamein New Jersey.
-
-***
+To better understand effects of rest and home/away advantage future research should focus on distances the teams have to travel between the games.
+For example if a team is travelling from one coast to another- it is likely to affect them more comparing to a team from New York 'travelling' to play an away game in New Jersey.
 
 ##Effects on accuracy
+
+###Introduction
 
@@
 In this analysis fatigue was measured as the period of the game, as it can be assumed that the later in the game the higher the fatigue of the players.
+The descriptive statistics of the shots per period can be seen below.
+
+```{r, echo=FALSE}
+bar.period <- ggplot(df) + geom_bar(aes(x = PERIOD))
+bar.period + labs(title = "Number of shots taken per period") + scale_x_discrete(name ="Period", limits=c("1","2","3","4", "5", "6", "7")) + ylab("Number of shots")
+```
 
@@
 The accuracies for each period are detailed in the table below.
+
+**Table 8**: Accuracy per period
 
@@
 
-```{r}
+```{r, echo=FALSE}
 graph.acc <- ggplot(acc, aes(x = seconds.left, y = accuracy)) + geom_point() + geom_smooth()
@@
 This means that player have a higher accuracy in periods 1 and 3 than in the following one.
-As the table shows, the changes in accuracy are minor (from 46% to 44% over the regular game time) and often located on digit-level, so although the changes are significant they are negligible.
+As the table 8 shows, the changes in accuracy are minor (from 46% to 44% over the regular game time) and often located on digit-level, so although the changes are significant they are negligible.
 After all, the Null hypothesis must be rejected as the statistical methods used show differences for accuracy in later periods of the game, but as they are negligible in the regular game time but it can be concluded that accuracy does not suffer for later game periods as long as the game ends in regular time.
@@
 
-```{r}
+```{r, echo=FALSE}
 
@@
 
+As there are much more shots taken in periods 1 to 4 than in 5 to 7 the differences in the accuracy have to be treat with caution.
 The shot clock variable had missing values for 5567 shots, equal to 4.34% of all shots on the dataset, limiting the meaningfulness of the findings in the accuracy per second part.
 
-***
-
-## Shooting Percentage Analysis
+## Shooting Percentage Analysis (James)
+
 ###Introduction
@@
   5. The correlation (p = .05) between the results of the matches and the performances of top scorers was investigated. 
-  6. A logistic regression with the outcome of the match as dependent variable and FGA and percentage of FG made as independent variables. 
-  7. Another logistic regression test was run in order to examine three special players: Russell Westbrook, Kawhi Leonard and Gordon Hayward.
+  6. A logistic regression analysis with the outcome of the match as dependent variable and FGA and percentage of FG made as independent variables was conducted.
 
@@
 ```
-
-### Find field goal percentage (FG percentage) for each match
 
@@
 ```
-##a. Field Goal Attempts (FGA)
-
-There was no correlation found between the result of a match and the number of attempts of the player made in that match.
-The Pearson’s correlation coefficient for the relationship between the variables W and FGA was r(832) < .01, p = .794, failing to reach level of significance (p = .05).
-
-##b.  Field Goal Percentage (FG%)
-
-There was no correlation found between the result of a match and the field goal percentage of the player in that match.
-The Pearson’s correlation coefficient for the relationship between the variables W and FGA was r(832) < .01, p = .177, failing to reach level of significance (p = .05).
+###Results
+
+####a. Field Goal Attempts (FGA)
+
+There was no correlation found between the outcome of the matches in this season and the number of attempts of the players made.
+The Pearson’s correlation coefficient for the relationship between the variables W and FGA was r(832) < .01, p = .794, failing to reach the level of significance (p = .05).
+
+####b.  Field Goal Percentage (FG%)
+
+There was no correlation found between the outcome of the matches in this seaon and the field goal percentage of the players.
+The Pearson’s correlation coefficient for the relationship between the variables W and FGA was r(832) < .01, p = .177, failing to reach the level of significance (p = .05).
 
@@
 ```
-Table xxx: The logistic regression for FGA and FG% of 15 players.
+**Table 9**: Result for the logistic regression analysis for FGA and FG% of 15 players
 
@@
 
-Table xxx: The logistic regression for FGA and FG% of the 3 special players.
+**Table 10**: Result for the anova for FGA and FG% for 15 players
+
+   | |           Df| Deviance| Resid. Df| Resid. Dev
+-|-|-|-|- 
+|NULL        |     |              | 777   |  188.73
+|FGA         |   1  | 0.0014   |    776  |   188.73
+|FG.percentage | 1 |  7.4664   |    775  |   181.26
+
+An ad hoc data analysis was used as 3 of the 15 players stood out in particular.
+To examine whether their data result in a better model, another logistic regression analysis was performed with the same dependent and independent variables for three special players: Russell Westbrook, Kawhi Leonard and Gordon Hayward.
+
+```{r, echo=FALSE}
+graph01 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
+graph01 <- graph01 + labs(title = "Differences in FGA for games lost/won of the three special players") + xlab("Player name") + ylab("FGA")
+graph02 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W)) 
+graph02 <- graph02 + labs(title = "Differences in FG% for games lost/won of the three special players") + xlab("Player name") + ylab("FG%")
+grid.arrange(graph01, graph02, ncol = 2)
+```
+
+The box plots above shows the performances of the three special players in terms of FGA and FG% respectively.
+
+**Table 11**: Result for the logistic regression analysis for FGA and FG% of the 3 special players
+
 | | Estimate | Std. Error | t value | Pr(>t)   
+-|-|-|-| - 
+|(Intercept)   | 0.042259 |  0.164816 |  0.256  |  0.798    
+|FGA          | -0.002679 |  0.006468 |  -0.414  |    0.679    
+|FG.percentage |  0.012329 |   0.002736 |   4.506  | 1.35e-05 ***
+
+**Table 12**: Result for the anova for FGA and FG% for the 3 special players
+
+| |         Df| Deviance| Resid. Df| Resid. Dev
 |-|-|-|-| - 
-|(Intercept)   | 0.042259 |  0.164816 |  0.256  |  0.798    
-|FGA          | -0.002679 |  0.006468|  -0.414  |    0.679    
-|FG.percentage|  0.012329|   0.002736|   4.506  |1.35e-05 ***
-
-The box plots above shows the performances the three special players in terms of FGA and FG% respectively. The red (blue) boxes contain the data when the team lost (won) the match.
+|NULL         |   |         |       147|     36.669
+|FGA          |  1|   0.0498|       146|     36.619
+|FG.percentage|  1|   4.4978|       145|     32.121
 
@@
 From the correlation tests between the match results and players’ performances, it can be noticed that the increase in number of field goal attempts from top scorers is unlikely to make any impact on the results of the game.
-The correlation coefficient between the results of the game and FG% is sightly higher but these two factors are still not correlated since r(832) < .01.
-Meanwhile, the p-values of both correlation tests are too higher to reach the default level of significance (p = .05).
-
-From the logistic regression test for FGA and FG%.
-It was found that FGA did not have any impact on the result of the match.
-FG% had significant correlation with the result of match.
-However, the anova test showed that the deviances of these two factors were both quite large and thus further investigations must be done in order to improve this model.
-
-There was no correlation found between field goal attempts of top scorers and game results.
-A possible explanation is that it is easier for the opponents to focus on a certain player on defensive side.
+The p-values of both correlation tests are too high to reach the level of significance (p = .05).
+
+From the logistic regression analysis for FGA and FG%, it was found that FGA did not have any impact on the result of the match.
+FG% is a significant predictor for the result of match.
+However, the anova showed that the residual deviances of these two factors were both quite large and thus further investigations must be done in order to improve this model.
+
+There was no correlation found between field goal attempts of top scorers and outcome of the matches.
+A possible explanation is that it is easier for the opponent to focus on a certain player on defensive side and then the overall efficiency of the team will start to decrease.
 Thus, over-using the scoring ability of a leading player is not likely to be a good approach if the team wants to increase the likelyhood of winning a game.
 Meanwhile, the accuracy of these players has a stronger impact on the outcome of the game.
-This is due to fact that once a player is being efficient on the offensive side, that player can help the team by not only keeping shooting but also creating more opportunities for his teammates.
-
-Based on results above, an additional research has been done regarding to the performances of three special players: Russell Westbrook, Kawhi Leonard and Gordon Hayward.
-These three players were chosen because their teams are more likely to take the victory when they score more in single game according to the bar plot and t-test.
+This could be explained by that once a player is being efficient on the offensive side, that player can help the team not only by shooting but also by creating more opportunities for his teammates.
+
+Based on the results above, an additional analysis was done regarding the performance of three particular players: Russell Westbrook, Kawhi Leonard and Gordon Hayward.
+These three players were chosen because their teams are more likely to take the victory when they score more in single game according to Diagram and t-test.
 
@@
 Based on these findings, another logistic regression was done with these three players.
-It was shown that p < .001 and residual deviance from anova dropped by 4.498 which is about 12.3% when FG% was considered.
-In comparison, the deviance was above 200 when 15 players were tested.
-
-Therefore, these three players are somehow different from others as they can contribute more to the win rates of their teams by their increasing shooting accuracy.
-There could be many reasons behind this, including their team compositions and the roles of the players.
-Further research has to be done to if one wants to find out the differences of these players and others.
+It was shown that p < .05 and residual deviance from anova dropped by 4.498 which is about 12.3% when FG% was considered.
+In comparison, the deviance was above 180 when 15 players were analysed.
+For these three players, their teams are more likely to win by 1.2% when their FG% increase by 1.
+
+Therefore, the analysis on these three players has shown that there is a difference between them and the others as they can contribute more to the win rates of their teams by their increasing shooting accuracy.
+There could be various reasons behind this observation, such as their team compositions and the roles of these players.
+Further research has to be done to in order to find out why the performances of these special players are more decisive on the outcome.
 
 ###Limitations
-
-The data set only contains 904 out of 1230 matches in the entire regular season.
-So there might be a chance that the highlight performances of certain players are missed out in our analysis.
-Meanwhile, it would be ideal if the contributions of all players can be analysed in more aspects including number of rebounds and assists.
-
-***
+The data set only covers 904 of 1230 matches in the regular season.
+So there might be a chance that the highlight performances of certain players are missed out in the analysis.
+Meanwhile, it would be ideal if more aspects of data, such as rebounds and assists, could be analysed.
 
@@
 distance_level$Accuracy <- round(distance_level$SCORED / distance_level$Freq, 4)
+
+library(ggplot2)
+library(scales)
+ggplot(data = df) + geom_bar(aes(x = df[, "CLOSE_DEF_DIST"])) +
+  scale_x_discrete("Closest Defender Distance", breaks = c (0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24), limits = 0:24) + 
+  ggtitle("Defender Distance Distribution")
+
 ```
@@
 library(scales)
-ggplot(data = distance_level, aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
+ggplot(data = distance_level) +
   geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
@@
   scale_x_discrete("Closest Defender Distance", limits = 0:24) +  
-  ylab("Shooting Accuracy (%)") +
-  ggtitle("Closest Defender Distance vs Shooting Accuracy")
+  ylab("Shooting Accuracy (%)") + ggtitle("Closest Defender Distance vs Shooting Accuracy")
 ```
@@
   The explanation from the plot above are as follows:   
-  1. Defender distance range from 0 to 2 feet, the accuracy varied between **42%** to **50%**  
-  2. Defender distance range from 2 to 4 feet, the accuracy rate for this distance range from **44%** to **49%**  
-  3. Defender distance range from 4 to 6 feet, the accuracy rate for this distance range from **41%** to **46%**
-  4. Defender distance range from 6 to 8 feet, the accuracy rate for this distance range from **37%** to **42%**
-  5. Defender distance range from 8 to 10 feet, the accuracy rate for this distance range from **37%** to **54%**
-  6. Defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from **33%** to **60%**.  
-  7. Distance greater than 16 feet, the data varies between **0%** (no shots scored from that distance) to **100%** (certain successfull shot from that distance)
+  1. Defender distance range from 0 to 2 feet, the accuracy varied between 42% to 50%  
+  2. Defender distance range from 2 to 4 feet, the accuracy rate for this distance range from 44% to 49%  
+  3. Defender distance range from 4 to 6 feet, the accuracy rate for this distance range from 41% to 46%
+  4. Defender distance range from 6 to 8 feet, the accuracy rate for this distance range from 37% to 42%
+  5. Defender distance range from 8 to 10 feet, the accuracy rate for this distance range from 37% to 54%
+  6. Defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from 33% to 60%.  
+  7. Distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance)
 
@@
   Next analysis will be calculated from correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:
+  There was medium to strong positive correlation found between the defender distance and shooting accuracy.
+The Pearson’s correlation coefficient for the relationship between the variables Defender Distance and Accuracy was r(297) = .59, p < .05, as part of level of significance (p = .05).
 
@@
 
-```{r, echo=FALSE}
+```{r, include=FALSE}
 #calculate correlation
@@
 For example, for distance range between 0 (zero) to 8 feet, the correlation measurement will be as follows:
-
+  There was medium negative correlation found between the defender distance and shooting accuracy in this range.
+The Pearson’s correlation coefficient for the relationship between the variables Defender Distance and Accuracy was r(297) = -0.40, p < .05, as part of level of significance (p = .05).
 ```{r, include=FALSE}
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 #Plotting the result
@@
   Analysis has also produced a correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:  
+  There was strong negative correlation found between the shooting distance and shooting accuracy.
+The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(446) = -0.87, p < .05, as part of level of significance (p = .05).
 ```{r, include=FALSE}
@@
 
-```{r, echo=FALSE}
+```{r, include=FALSE}
 cor.test(dfShootingZerotoAll$shootingDistanceZerotoAll, dfShootingZerotoAll$shootingAccuracyDist1)
@@
 However, if the correlation test taken for certain range distance based on the diagram, the correlation result would be different as follows:
-
+  There was low positive correlation found between the shooting distance and shooting accuracy between range 10 to 20 feet.
+The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(99) = .11, p > .05, fail to reach level of significance (p = .05).
 ```{r, include=FALSE}
@@
 ###Introduction
-  The analysis aims to show the effect of the 'hot hand theory' on NBA players.
-The analysis is based on the hypothesis that a shooter with a hot hand will have shots that are not independent of one another.
+  The specific analysis aims to show the effect of the 'hot hand theory' on NBA players.
+The analysis is aimed to establish whether there are  the hypothesis that a shooter with a hot hand will have shots that are dependent of one another.
 If the shooter makes his first shot, the probability of making his next shot is higher.
@@
   3. Extract zeros and keep only streaks above 1
-  4. Visualise streaks' length distribution - barrplots
+  4. Visualise streaks' length distribution - barplots
   5. Visualise streaks' length via boxplot
   6. Calculate the shooting percentage of the 3 players 
-  7. Create a sample independent shooter setting as shooting percentage the real player's shooting percentage
+  7. Create a sample independent shooter 
   8. Compare distributions of sample and real player
@@
 
-#From 'FGM' column we have that 1 represents a basket made and 0 represents a basket missed
-#Define the length of a shooting streak to be the number of consecutive baskets made until a miss occurs
-#Create a function that gives us a vector with the number of total streaks(from all players and all games)
+#From 'FGM' column: 1 represents a basket made and 0 represents a basket missed
+#Defining the length of a shooting streak to be the number of consecutive baskets made until a miss occurs
+#Creating a function that gives a vector with the number of total streaks(from all players and all games)
 number_of_streaks <- function(x) {
@@
 ptssum<- cast(df1, GAME_ID1+ player_id + player_name  ~ PTS); head(ptssum)
-#creating "total points per player, per game" column
+#Creating "total points per player, per game" column
 ptssum$totalpts<- ptssum[,"2"]*2 + ptssum[,"3"]*3
@@
 
-#Set a new data frame for the 3 players with best performances(higher points)
+#Setting a new data frame for the 3 players with best performances(higher points)
 best_players_df <- head(pttsum2, 3)
 
-#Create a function that gives us a vector of streaks for a specific player and game
+#Creating a function that gives a vector of streaks for a specific player and game
 player_and_game <- function(x, y) {
@@
 
-#Create a list that gives us the streaks of each player and game (5 best performances)
+#Creating a list that gives the streaks of each player and game (3 best performances)
 streak_per_player <- list()
@@
 
-#Remove 0 from streak_per_player in order to show only the streaks achieved
+#Removing zeros from streak_per_player in order to show only the streaks achieved
 streak_per_player[[1]] <- streak_per_player[[1]][streak_per_player[[1]] != 0]
@@
 ```{r, include=FALSE}
-
-#Create a data frame for the strikes of each one of the 3 players (cannot create a data frame including all of them, because each list has different length)
+#Creating a data frame for the strikes of each one of the 3 players (cannot create a data frame including all of them, because each list has different length)
 player1_df <- data.frame('streaks' = streak_per_player[[1]])
@@
 player3_df <- data.frame('streaks' = streak_per_player[[3]])
-
-#Show distribution for each player's streak lengths
+```
+
+```{r}
+#Showing distribution for each player's streak lengths
 gg1 <- ggplot(data = player1_df, aes(x = streaks)) + geom_bar(fill = 'brown') + labs(title = 'Kyrie Irnving')
@@
 ####Barplot result
-  The distribution of the three players is unimodal and left skewed. There are shown some extremely long shooting streaks made(i.e. Klay Thompson with 13 shots in a row).
+  The distribution of the three players is unimodal and left skewed. There are shown some extremely long shooting streaks made (i.e. Klay Thompson with 13 shots in a row).
 
@@
 
-```{r, include=FALSE}
+```{r}
 #Create a boxplot for each player's streak lengths 
@@
 
-  1. The typical length of a streak is 1 (median = 1)
+  1. The typical length of a streak is 1 
   2. The Interquartile Range is 1 
@@
 
-  1. The typical length of a streak is 1.5 (median = 1.5)
+  1. The typical length of a streak is 1.5
   2. The Interquartile Range is 1
@@
 
-  1. The typical length of a streak is 8 (median = 8)
+  1. The typical length of a streak is 8 
   2. The Interquartile Range is 5
-  3. There is a streak of length 13 (max = 13)
+  3. There is a streak of length 13
 
@@
 
-  The distribution of all players is left skewed and all of them (especially player3) have some long shooting streaks.
+  The distribution of all players is left skewed and all of them (especially Klay Thompson) have some long shooting streaks.
 In order to prove/disprove the hot hand theory,  the examination of the indipendence of the shots is required.
@@
 ###PLAYER 1 VS SAMPLE PLAYER###
-#Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
+#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
 set.seed(974)
@@
 
-#Set 1 for baskets made and 0 for baskets missed
+#Setting 1 for baskets made and 0 for baskets missed
 simul_shooter <- as.list(simul_shooter)
@@
 
-#Use previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
+#Using previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
 streaks_of_simulshooter <- number_of_streaks(simul_shooter)
 
-#Create new data frames for the graphs 
+#Creating new data frames for the graphs 
 simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
@@
 summary(realshooter_df$streaks)
-
-#Compare barplots of sample shooter and real shooter1
+```
+
+```{r}
+#Comparing barplots of sample shooter and real shooter1
 #If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
@@
 ####Plot Result
-  After running many times the sample shooter, there are times that the distributions of sample and real shooter are quite similar.
+  After running many times the sample, there are several times when the distributions of sample and real shooter are quite similar.
 As a consequence, it can be concluded that Mo Williams does not have a hot hand.
@@
 ###PLAYER 2 VS SAMPLE PLAYER###
-#Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
+#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter2 
 set.seed(572)
@@
 
-  After running many times the sample shooter, there are times that the distributions of sample and real shooter are quite similar.
+  After running many times the sample, there are several times when the distributions of sample and real shooter are quite similar.
 As a consequence, it can be concluded that Mo Williams does not have a hot hand.
@@
 ###PLAYER 3 VS SAMPLE PLAYER###
-#Simulation - create a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
+#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter3 
 set.seed(57)
@@
 There is no such distribution in the sample, so it could be assumed that he is likely to have a hot hand.
-Further statistical analysis (logistic regression) should be done in order to reach a conclusion.
+Further statistical analysis - logistic regression - should be done in order to reach a more accurate conclusion.
+
+###Discussion
+
+  As the analysis shown, there are some extremely high shot streaks achieved from the players.
+Kyrie Irnving achieved six shots in a row, Mo Williams 4 shots in a row - twice, while Klay Thompson 13 shots in a row.
+However, to assume that a player has a hot hand, it is required to prove that his continuous shots are dependent of one another.
+As the vectors of streaks do not contain continuous variables, a linear regression cannot be used.
+Creating a sample independent shooter with same number of attempts and same shooting percentage as the real's one, is the closest approach to the specific question.
+In the first two cases, there are several times when the distribution of streaks of the sample independent player is similar to the real shooter's distribution.
+As the sample shooter's distribution visualises independent shots, it can be concluded that Kyrie Irnving's and Mo Williams' continuous shots are independent of one another and as a result they have not a hot hand.
+On the other hand, Klay Thompson is likely to have a hot hand, as his shots length distribution diverges significantly from sample's distribution.
+
+###Limitations
+
+  Since the dataset is referred to only one season, the size of the vectors of streaks for each player is very small, a fact that can lead to misleading results.
+Moreover, streaks do not constitute continuous variables and as a result they cannot be used in a linear regression.
+
+###Future research
+
+  Since the dataset is reffered to only one season and the analysis is not able to eliminate the hot hand theory, it would be an interesting topic to be investigated in future seasons.
 
@@
 
+***
 #Conclusion
 
+This report has demonstrated results of mathematical and statistical analysis of the dataset containing information of shots from NBA season 2014-2015.
+The report specifically focused on number of research areas covering teams and players performance.
+The report described steps taken to produce the results, discussed and evaluated the produced results and suggested potential areas for future research.
+
+***
 #References
@@
   Ahart (1973). In Kendall, P. and Hollon, S. (1979). Cognitive-behavioral interventions. New York: Academic Press.
+  Economist: As sweet as ever (2015) Available at: http://www.economist.com/blogs/gametheory/2015/06/home-advantage-basketball (Accessed: 16 October 2016).
   Erculj, F.
@@
 PLOS ONE, 11(6), p.e0157215.
-
-ECONOMIST. As sweet as ever (2015) Available at: http://www.economist.com/blogs/gametheory/2015/06/home-advantage-basketball (Accessed: 16 October 2016).

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
Table 13 added

@@
 ###Plot Result
-  The explanation from the plot above are as follows. The shooter distance ranges and accuracy:   
-  1.  0 to 2 feet: the accuracy varied between **62%** to **68%**  
-  2.  2 to 4 feet: the accuracy rate for this distance range from **61%** to **67%**  
-  3.  4 to 6 feet: the accuracy rate for this distance range from **44%** to **58%**  
-  4.  6 to 8 feet: the accuracy rate for this distance range from **38%** to **51%**  
-  5.  8 to 10 feet: the accuracy rate for this distance range from **35%** to **42%**  
-  6.  10 to 12 feet: the accuracy rate for this distance range from **34%** to **42%**  
-  7.  12 to 14 feet: the accuracy rate for this distance range from **37%** to **47%**  
-  8.  14 to 16 feet: the accuracy rate for this distance range from **33%** to **51%**  
-  9.  16 to 18 feet: the accuracy rate for this distance range from **37%** to **48%**  
-  10. 18 to 24 feet: the accuracy rate for this distance range from **33%** to **43%**  
-  11. 24 to 28 feet: the accuracy rate for this distance range from **18%** to **38%**  
-  12. more than 28 feet: the accuracy varies greater than previous distance range, starting from **0%** (no shots scored from that distance) to **50%**
+
+**Table 13**: The relationship between the shooter distance ranges and accuracy
+
+  Distance (in feet) | Accuracy (range in %)
+  - | -
+  0 - 2 | 62 - 68
+  2 - 4 | 61 - 67
+  4 - 6 | 44 - 58
+  6 - 8 | 38 - 51
+  8 - 10 | 35 - 42
+  10 - 12 | 34 - 42
+  12 - 14 | 37 - 47
+  14 - 16 | 33 - 51
+  16 - 18 | 37 - 48
+  18 - 24 | 33 - 43
+  24 - 28 | 18 - 38
+  28+ | 0 - 50
 

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd

@@
 shot_results <- c('made', 'missed')
-simul_shooter <- sample(shot_results, size = length(streak_per_player[[1]]), replace = TRUE, prob = c(shooting_percentage[[1]], 1 - shooting_percentage[[1]]))
+simul_shooter <- sample(shot_results, size = sum(best_players_df[1,4], best_players_df[1,5], best_players_df[1,6]), replace = TRUE, prob = c(shooting_percentage[[1]], 1 - shooting_percentage[[1]]))
 simul_shooter
@@
 #Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter2 
-set.seed(572)
+set.seed(57)
 shot_results <- c('made', 'missed')
-simul_shooter <- sample(shot_results, size = length(streak_per_player[[2]]), replace = TRUE, 
-                        prob = c(shooting_percentage[[2]], 1 - shooting_percentage[[2]]))
+simul_shooter <- sample(shot_results, size = sum(best_players_df[2,4], best_players_df[2,5], best_players_df[2,6]), replace = TRUE, prob = c(shooting_percentage[[2]], 1 - shooting_percentage[[2]]))
 simul_shooter
@@
 shot_results <- c('made', 'missed')
-simul_shooter <- sample(shot_results, size = length(streak_per_player[[3]]), replace = TRUE, 
+simul_shooter <- sample(shot_results, size = sum(best_players_df[3,4], best_players_df[3,5], best_players_df[3, 6]), replace = TRUE, 
                         prob = c(shooting_percentage[[3]], 1 - shooting_percentage[[3]]))

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
minor changes

--- 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
@@
 knitr::opts_chunk$set(echo=TRUE)
-knitr::opts_chunk$set(echo = TRUE)
 ```
@@
 
-The NULL Hypothesis: Teams playing at home is more likely to win then when playing away. 
-The NULL Hypothesis: Analysis of means for home and away wins does not show any significant difference
+The NULL Hypothesis: Analysis of means for home and away wins does not show any significant difference in the means.
 
@@
     5. Perform a t-test (p .05) to examine whether there is a difference in the means between teams winning at home and teams winning away
-    6. Linear regression with Final margin as dependent variable and Location as independent variable to examine whether playing at home results predicts a higher final margin
-
-The graph below shows descriptive statistics for home/away wins:
+    6. Linear regression with "Final margin"" as dependent variable and "Location" as independent variable to examine whether playing at home predicts a higher final margin
+
+Table 1 below shows descriptive statistics for home/away wins:
+
+**Table 1**: Proportion of total games split by Location win factor
+
+X     | Home Win | Away Win
+----- | -------- | ---
+Game  | 506      | 398
+Perc  | 55.97%   | 44.03%
 
@@
 ###Results
-
-###Findings
-**Table 1**: Proportion of total games split by Location win factor
-
-X     | Home Win | Away Win
------ | -------- | ---
-Game  | 506      | 398
-Perc  | 55.97%   | 44.03%
 
@@
 
-The T-test used to verify the differences in the means seen in the plot reached significance, t(58) = -2.39, p = .02.
-
-###Regression analysis
+The T-test used to verify the differences in the means seen in the plot reached the level of significance, t(58) = -2.39, p = .02.
+
+####Regression analysis
 
@@
     5. Calculating ratios for effects of rest days on win/loss form home/away
-    6. Run a regression analysis
-    7. Discuss the findings of the data
 
@@
 
-Fatigue: The Null hypothesis states that there is no significant difference (p < .05) for the shot accuracy in later periods of the game.
-
-Stress: The Null hypothesis states that there is no significant difference (p < .05) for the shot accuracy in more stressful events.
+Fatigue: The Null hypothesis states that there is no significant difference (p < .05) in the mean of shot accuracy in later periods of the game.
+
+Stress: The Null hypothesis states that there is no significant difference (p < .05) in the means for the shot accuracy for lower seconds left on the shot clock.
 
@@
   4. The players with most outstanding performances were picked out and another subset was generated for these players.
-  5. The correlation (p = .05) between the results of the matches and the performances of top scorers was investigated. 
+  5. The correlation (p = .05) between the results of the matches and the performances regarding FGA and FG% of top scorers was investigated. 
   6. A logistic regression analysis with the outcome of the match as dependent variable and FGA and percentage of FG made as independent variables was conducted.

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd

@@
   1. Assign shot result to determine whether the shot is scored (1) or missed (0)  
-  2. Select and analyse subset of **all** closest defender distance values  
+  2. Select and analyse subset of all closest defender distance values  
   3. Count the occurence fOr every distance  
-  4. Calculate a number of uccessful shots for every distance point  
+  4. Calculate a number of successful shots for every distance point  
   5. Calculate accuracy ratio for every defender distance value
-  6. Plot and discuss the result
+
 ```{r, include=FALSE}
@@
 ###Correlation Test
-  Next analysis will be calculated from correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:
+  The next analysis calculates the correlation between both variables. If the analysis included all of the data points for every distance, the correlation result would be as follows:
   There was medium to strong positive correlation found between the defender distance and shooting accuracy.
-The Pearson’s correlation coefficient for the relationship between the variables Defender Distance and Accuracy was r(297) = .59, p < .05, as part of level of significance (p = .05).
+The Pearson’s correlation coefficient for the relationship between the variables "Defender Distance" and "Accuracy" was r(297) = .59, p < .05, as part of level of significance (p = .05).
 
@@
   The correlation test result produced slightly higher number of correlation between defender distance with shooting accuracy for all shot distance attempts (0.59).
-If the findings made only based on this number only, there would be obvious that further defender distance will **automatically** increase shooting accuracy.
+If the findings made only based on this number only, there would be obvious that further defender distance will automatically increase shooting accuracy.
 However, when analysing distance ranges the result was somewhat different.
@@
 ###Discussion  
-  The analysis demonstrated that the accuracy rate for every player was **affected** by the defender distance for certain ranges.
+  The analysis demonstrated that the accuracy rate for every player was affected by the defender distance for certain ranges.
 For defender distance starting from 0 (zero) to 8 (eight) feet, the shooting accuracy had negative correlations.

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
Table 13 and 14 added

@@
 ###Plot Result
-  The explanation from the plot above are as follows:   
-  1. Defender distance range from 0 to 2 feet, the accuracy varied between 42% to 50%  
-  2. Defender distance range from 2 to 4 feet, the accuracy rate for this distance range from 44% to 49%  
-  3. Defender distance range from 4 to 6 feet, the accuracy rate for this distance range from 41% to 46%
-  4. Defender distance range from 6 to 8 feet, the accuracy rate for this distance range from 37% to 42%
-  5. Defender distance range from 8 to 10 feet, the accuracy rate for this distance range from 37% to 54%
-  6. Defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from 33% to 60%.  
-  7. Distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance)
+  The explanation from the plot above is summerised in the following table:
+
+** Table 13**: Relationship between the defender distance and accuracy
+
+Defender Distance (range in feet) | accuracy (range in %) 
+- | -
+0 - 2 | 42 - 50
+2 - 4 | 44 - 49
+4 - 6 | 41 - 46
+6 - 8 | 37 - 42
+8 - 10 | 37 - 54
+10 - 16 | 33 - 60
+16+ | 0 - 100
 
@@
 
-**Table 13**: The relationship between the shooter distance ranges and accuracy
-
-  Distance (in feet) | Accuracy (range in %)
+**Table 14**: Relationship between the shooter distance ranges and accuracy
+
+  Shooter Distance (in feet) | Accuracy (range in %)
   - | -

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
Removed all "HEAD"

@@
 
+#change distance level to numeric type for correlation test function
+distance_level$Var1 <- as.numeric(distance_level$Var1)
+distance_level$Var1 <- distance_level$Var1 * 0.1 - 0.1 #to keep the distance point as it is
+
 library(ggplot2)
@@
 ```{r, include=FALSE}
-#change distance level to numeric type for correlation test function
-distance_level$Var1 <- as.numeric(distance_level$Var1)
-distance_level$Var1 <- distance_level$Var1 * 0.1 - 0.1 #to keep the distance point as it is
 
@@
 shooting_distance_level$Accuracy <- round(shooting_distance_level$SCORED / shooting_distance_level$Freq, 4)
+
+#change distance level to numeric type for correlation test function
+shooting_distance_level$Var1 <- as.numeric(shooting_distance_level$Var1)
+shooting_distance_level$Var1 <- shooting_distance_level$Var1 * 0.1 - 0.1 #to maintain distance data point as it is
 ```
@@
 ```{r, include=FALSE}
-#change distance level to numeric type for correlation test function
-shooting_distance_level$Var1 <- as.numeric(shooting_distance_level$Var1)
-shooting_distance_level$Var1 <- shooting_distance_level$Var1 * 0.1 - 0.1 #to maintain distance data point as it is
 
@@
 ###Introduction
+  The analysis aims to investigate whether the effect of the 'hot hand theory' is valid based on the given dataset.
+The 'hothand theory' states that if the shooter makes his first shot, the probability of making his next shot is higher.
+
   The specific analysis aims to show the effect of the 'hot hand theory' on NBA players.

### 2016-10-16 dataframe.R
Delete dataframe.R

@@ -1,31 +1,0 @@
-
-df <- read.csv("shot_logs.csv", header = TRUE)
-
-GameID <- df[,1] #Identification number of a game
-MatchUp <- df[,2] #Date of the game and names of playing teams
-Location <- df[,3] #Home or Away
-WinLoss <- df[,4] #Win or loss
-FinalMargin <- df[,5] #Final score margin
-ShotNumber <- df[,6] #shot number
-Period <- df[,7] #period/quarter the shot took place
-GameClock <- df[,8] #amount of time shown on the game clock
-ShotClock <- df[,9] #amount of time shown on the shot clock. A team has 24 seconds to complete a shot
-Dribbles <- df[,10] #number of dribbles a player commits before the shot
-TouchTime <- df[,11] #number of seconds the player had the ball before making the shot
-ShotDistance <- df[,12] #distance to the basket when the shot is made
-PointType <- df[,13] #amount of points the shot carries (two or three) 
-ShotResult <- df[,14] #result of ths shot (made/missed)
-ClosestDefenderName <- df[,15] #name of the defender closest to the player taking the shot
-ClosestDefenderID <- df[,16] #ID of the defender closest to the player taking the shot
-ClosestDefenderDistance <- df[,17] #distance to the closest defender 
-FieldGoalMade <- df[,18] #1-shot is made/scored; 2-shot is missed
-Points <- df[,19] #number of points
-PlayerName <- df[,20] #name of the player making the shot
-PlayerID <- df[,21] #id of the player making the shot
-
-summary(df)
-  
-
-
-
-

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
further minor changes in wording

@@
   5. Calculate accuracy ratio for every defender distance value
+  6. Seperate data in two ranges - 0 to 8 feet distance and above 8 feet distance
 
@@
 ###Correlation Test
-  The next analysis calculates the correlation between both variables. If the analysis included all of the data points for every distance, the correlation result would be as follows:
-  There was medium to strong positive correlation found between the defender distance and shooting accuracy.
-The Pearson’s correlation coefficient for the relationship between the variables "Defender Distance" and "Accuracy" was r(297) = .59, p < .05, as part of level of significance (p = .05).
+  The next analysis calculates the correlation between both variables. If the analysis included all data points for every distance, the correlation result would be as follows:
+  There is a medium to strong positive correlation between the defender distance and shooting accuracy.
+The Pearson’s correlation coefficient for the relationship between the variables "Defender Distance" and "Accuracy" was r(297) = .59, p < .01, with a level of significance of p = .05.
 
@@
 ###Correlation Test Result  
-  The correlation test result produced slightly higher number of correlation between defender distance with shooting accuracy for all shot distance attempts (0.59).
-If the findings made only based on this number only, there would be obvious that further defender distance will automatically increase shooting accuracy.
-However, when analysing distance ranges the result was somewhat different.
-For example, for distance range between 0 (zero) to 8 feet, the correlation measurement will be as follows:
-  There was medium negative correlation found between the defender distance and shooting accuracy in this range.
-The Pearson’s correlation coefficient for the relationship between the variables Defender Distance and Accuracy was r(297) = -0.40, p < .05, as part of level of significance (p = .05).
+  When analysing the relationship between distance and the accuracy for the range of 0 to 8 feet distance the result of the correlation is different:  
+  There was medium negative correlation between the defender distance and shooting accuracy in this range.
+The Pearson’s correlation coefficient for the relationship between the variables Defender Distance and Accuracy was r(297) = -0.40, p < .01, with the level of significance of p = .05.
 ```{r, include=FALSE}
@@
 
-  The correlation measurement between 0 (zero) to 8 feet would be negative (-0.40). This result was different with the previous calculation done for all data points (0.59).
-
 ###Discussion  
-  The analysis demonstrated that the accuracy rate for every player was affected by the defender distance for certain ranges.
-For defender distance starting from 0 (zero) to 8 (eight) feet, the shooting accuracy had negative correlations.
+  The analysis demonstrated that the accuracy rate was affected by the defender distance for certain ranges.
+For the defender distance starting from 0 to 8 feet, there is a negative correlation between distance and accuracy, implying that players are more likely to miss when the defender comes closer.
 For the distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance), which means that this part of the analysis should not be taken into account as the results were insignificant.
-It also shown from the dataset that only small attempt of shots were made from further defender distance.
+It also shows that only a small amount of shots were made for higher defender distances.
 Therefore, it would affect the accuracy calculation result.
 However, from the correlation perspective, the result would be different if the analysis included all of the data points versus only short distance data points.
-This was arguably due to further defender distance did not mean the shooter distance from the basket was closer.
-Therefore, we **reject** the Null Hypothesis, since the defender distance affected shooting accuracy.
+This was arguably due to the fact that the defender distance does not mean the shooter distance from the basket was closer.
+Therefore, we reject the Null Hypothesis, since the defender distance affected shooting accuracy.
 
@@
 ####Introduction
+
 This analysis focuses on whether there is any significant effect resulting from shooting distance from basket on the shot outcome.
@@
   1. Assign shot result to determine whether the shot is scored (1) or missed (0)  
-  2. Select subsets of **all** shooting distances from the basket  
+  2. Select subsets of all shooting distances from the basket  
   3. Count the occurence fOr every distance to basket value  

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
more wording changes

@@
   4. Allocate shot success rate to every distance from basket value 
-  5. Plot and discuss the result
+  5. Examine the relationship between distance to the basket and shot outcome with a correlation test
+  6. Seperate data into ranges
+
 ```{r, include=FALSE}
@@
 
-  Analysis has also produced a correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:  
-  There was strong negative correlation found between the shooting distance and shooting accuracy.
-The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(446) = -0.87, p < .05, as part of level of significance (p = .05).
+  There is a strong negative correlation between the shooting distance and shooting accuracy.
+The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(446) = -0.87, p < .05, with the level of significance of p = .05.
 ```{r, include=FALSE}
@@
 ###Correlation Test Result  
-  The correlation test result shown strong negative correlation between shooting distance and shooting accuracy (-0.87).
-This meant that for all data points, the further a shooter from the basket, the shooter's accuracy would decrease.
-However, if the correlation test taken for certain range distance based on the diagram, the correlation result would be different as follows:
-  There was low positive correlation found between the shooting distance and shooting accuracy between range 10 to 20 feet.
+  If the correlation test taken for a certain range of distance based on the diagram, the correlation result would be different as follows:
+  There was low positive correlation found between the shooting distance and shooting accuracy for the range 10 to 20 feet distance.
 The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(99) = .11, p > .05, fail to reach level of significance (p = .05).
@@
 
-  The second correlation test produced a small positive correlation between range 10 to 20 feet (0.11). This result was different with previous measurement (-0.87)
-
 ###Discussion  
-  Based on the diagram and correlation above, it could be concluded that although the shooting distance directly affects shooting accuracy, the relationship between these variables are not perfectly linear.
-This was shown in the diagram where there is a relatively **similar** shooting accuracy for shots taken from 10 feet to 20 feet, regardless the shooter's performance.
-The success rate for longer distances, it could be arguably inferred that the shooters tried to attempt a shot in closing game time (buzzer beater) when the shot clock or game clock nearly reach zero- hence a longer distances.
-Therefore, the analysis would reject the  Null Hypothesis.
+  Based on the diagram and correlation above, it could be concluded that although the shooting distance affects the shooting accuracy, the relationship between these variables is not perfectly linear.
+This was shown in the diagram where the accuracy for shots taken from 10 to 20 feet is relatively similar, regardless the shooter's performance.
+Regarding the low success rate for higher distances, it could be arguably inferred that the shooter tried to attempt a shot in closing game time (buzzer beater) or when the shot clock or game clock nearly reach zero - Therefore, the analysis would reject the Null Hypothesis.
 

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
final tidying up

--- 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
@@
 
-The NULL Hypothesis: Analysis of means for home and away wins does not show any significant difference in the means.
+The NULL Hypothesis: Analysis of means for home and away wins does not show any significant difference
 
@@
 
+***
+
 ##Effects on accuracy
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 graph.acc <- ggplot(acc, aes(x = seconds.left, y = accuracy)) + geom_point() + geom_smooth()
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE,warning=FALSE}
 
@@
 
-## Shooting Percentage Analysis (James)
+***
+
+## Game result impact by field goal attempt
 
@@
 graph01 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
-graph01 <- graph01 + labs(title = "Differences in FGA for games lost/won of the three special players") + xlab("Player name") + ylab("FGA")
+graph01 <- graph01 + labs(title = "Differences in FGA for games lost/won /n of the three special players") + xlab("Player name") + ylab("FGA")
 graph02 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W)) 
-graph02 <- graph02 + labs(title = "Differences in FG% for games lost/won of the three special players") + xlab("Player name") + ylab("FG%")
+graph02 <- graph02 + labs(title = "Differences in FG% for games lost/won /n of the three special players") + xlab("Player name") + ylab("FG%")
 grid.arrange(graph01, graph02, ncol = 2)
@@
 ###Plot Result
-  The explanation from the plot above is summerised in the following table:
-
-** Table 13**: Relationship between the defender distance and accuracy
-
-Defender Distance (range in feet) | accuracy (range in %) 
-- | -
-0 - 2 | 42 - 50
-2 - 4 | 44 - 49
-4 - 6 | 41 - 46
-6 - 8 | 37 - 42
-8 - 10 | 37 - 54
-10 - 16 | 33 - 60
-16+ | 0 - 100
+  The explanation from the plot above are as follows:   
+  1. Defender distance range from 0 to 2 feet, the accuracy varied between 42% to 50%  
+  2. Defender distance range from 2 to 4 feet, the accuracy rate for this distance range from 44% to 49%  
+  3. Defender distance range from 4 to 6 feet, the accuracy rate for this distance range from 41% to 46%
+  4. Defender distance range from 6 to 8 feet, the accuracy rate for this distance range from 37% to 42%
+  5. Defender distance range from 8 to 10 feet, the accuracy rate for this distance range from 37% to 54%
+  6. Defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from 33% to 60%.  
+  7. Distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance)
 
@@
 The future analysis could be enhanced further by including data for additional seasons.
+
+***
 
@@
 
-**Table 14**: Relationship between the shooter distance ranges and accuracy
-
-  Shooter Distance (in feet) | Accuracy (range in %)
+**Table 13**: The relationship between the shooter distance ranges and accuracy
+
+  Distance (in feet) | Accuracy (range in %)
   - | -
@@
 ###Introduction
+
   The analysis aims to investigate whether the effect of the 'hot hand theory' is valid based on the given dataset.
 The 'hothand theory' states that if the shooter makes his first shot, the probability of making his next shot is higher.
-
-  The specific analysis aims to show the effect of the 'hot hand theory' on NBA players.
-The analysis is aimed to establish whether there are  the hypothesis that a shooter with a hot hand will have shots that are dependent of one another.
-If the shooter makes his first shot, the probability of making his next shot is higher.
+(Gilovich, T., Vallone, R.
+and Tversky, A.
+(1985))
 
@@
 
-```{r}
+```{r, echo=FALSE}
 #Showing distribution for each player's streak lengths
@@
 
-```{r}
+```{r, include=FALSE}
 #Create a boxplot for each player's streak lengths 
@@
 
-```{r}
+```{r, echo=FALSE}
 #Comparing barplots of sample shooter and real shooter1
@@
 ####Plot Result
-  After running many times the sample, there are several times when the distributions of sample and real shooter are quite similar.
-As a consequence, it can be concluded that Mo Williams does not have a hot hand.
+  After producing a sample of shots for a player, the result shows distributions of sample and real shooter are quite similar.
+As a consequence, it can be concluded that Kyrie Irving does not have a hot hand.
 
@@
 shot_results <- c('made', 'missed')
-simul_shooter <- sample(shot_results, size = length(streak_per_player[[2]]), replace = TRUE, 
+simul_shooter <- sample(shot_results, size = sum(best_players_df[2,4], best_players_df[2,5], best_players_df[2,6]), replace = TRUE, 
                         prob = c(shooting_percentage[[2]], 1 - shooting_percentage[[2]]))
@@
 
-  After running many times the sample, there are several times when the distributions of sample and real shooter are quite similar.
+  After producing a sample of shots for a player, the result shows distributions of sample and real shooter are quite similar.
 As a consequence, it can be concluded that Mo Williams does not have a hot hand.
@@
 shot_results <- c('made', 'missed')
-simul_shooter <- sample(shot_results, size = length(streak_per_player[[3]]), replace = TRUE, 
+simul_shooter <- sample(shot_results, size = sum(best_players_df[3,4], best_players_df[3,5], best_players_df[3,6]), replace = TRUE, 
                         prob = c(shooting_percentage[[3]], 1 - shooting_percentage[[3]]))
@@
 There is no such distribution in the sample, so it could be assumed that he is likely to have a hot hand.
-Further statistical analysis - logistic regression - should be done in order to reach a more accurate conclusion.
+However, further analysis of Klay Thompson's (and other basketball players) shooting streaks in other games is required to see if such a rare occurence is repeated(this is out of scope for this analysis).
 
@@
 
-  As the analysis shown, there are some extremely high shot streaks achieved from the players.
-Kyrie Irnving achieved six shots in a row, Mo Williams 4 shots in a row - twice, while Klay Thompson 13 shots in a row.
-However, to assume that a player has a hot hand, it is required to prove that his continuous shots are dependent of one another.
-As the vectors of streaks do not contain continuous variables, a linear regression cannot be used.
-Creating a sample independent shooter with same number of attempts and same shooting percentage as the real's one, is the closest approach to the specific question.
-In the first two cases, there are several times when the distribution of streaks of the sample independent player is similar to the real shooter's distribution.
-As the sample shooter's distribution visualises independent shots, it can be concluded that Kyrie Irnving's and Mo Williams' continuous shots are independent of one another and as a result they have not a hot hand.
-On the other hand, Klay Thompson is likely to have a hot hand, as his shots length distribution diverges significantly from sample's distribution.
+  The analysis demonstrates that the streaks generated by the players (K.
+Irving & M.
+Williams) are similar to randomly generated streaks, which use these players shooting probabilities.
+As a result there is no evidence to suggest that there is a hot hand phenomenon present here.
+
+  In the case of K. Thompson, his streak is seem as somewhat anomaly and further research is required to see if any similar results have been achieved by the player.
 
@@
 
-  Since the dataset is reffered to only one season and the analysis is not able to eliminate the hot hand theory, it would be an interesting topic to be investigated in future seasons.
-
-```{r, include=FALSE}
-#Regression Analysis
-
-df1$GAME_CLOCK <- as.numeric(df1$GAME_CLOCK)
-
-df1$AWAY_TEAM <- as.character(df1$AWAY_TEAM)
-
-linreg_fgm <- lm(FGM ~ LOCATION + W + FINAL_MARGIN + SHOT_NUMBER + PERIOD + GAME_CLOCK + SHOT_CLOCK+ DRIBBLES + TOUCH_TIME + SHOT_DIST + PTS_TYPE + CLOSEST_DEFENDER_PLAYER_ID + CLOSE_DEF_DIST + player_id + WINNER, df1)
-
-summary(linreg_fgm)
-
-linreg_fgm2 <- lm(FGM ~ FINAL_MARGIN + SHOT_NUMBER + PERIOD + SHOT_CLOCK + DRIBBLES + TOUCH_TIME + SHOT_DIST + PTS_TYPE + CLOSE_DEF_DIST + player_id, df1 )
-
-summary(linreg_fgm2)
-
-```
+  Since the dataset is reffered to only one season and the analysis is not able to eliminate the hot hand theory, it would be an interesting topic to be investigated by using data from more than one season.
 
@@
 Journal of Strength and Conditioning Research, 23(3), pp.1029-1036.
+  Gilovich, T., Vallone, R. and Tversky, A. (1985) ‘The hot hand in basketball: On the misperception of random sequences’, Cognitive Psychology, 17(3), pp. 295–314. doi: 10.1016/0010-0285(85)90010-6.
   Mascret, N., Ibáñez-Gijón, J., Bréjard, V., Buekers, M., Casanova, R., Marqueste, T., Montagne, G., Rao, G., Roux, Y.

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
final changes

@@
   1. Assign shot result to determine whether the shot is scored (1) or missed (0)  
-  2. Select subsets of **all** shooting distances from the basket  
+  2. Select subsets of all shooting distances from the basket  
   3. Count the occurence fOr every distance to basket value  
   4. Allocate shot success rate to every distance from basket value 
-  5. Plot and discuss the result
+  5. Examine the relationship between distance to the basket and shot outcome with a correlation test
+  6. Seperate data into ranges
+
 ```{r, include=FALSE}
@@
 
-  Analysis has also produced a correlation test. If the analysis included **all** of the data points for every distance, the correlation result would be as follows:  
-  There was strong negative correlation found between the shooting distance and shooting accuracy.
-The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(446) = -0.87, p < .05, as part of level of significance (p = .05).
+  There is a strong negative correlation between the shooting distance and shooting accuracy.
+The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(446) = -0.87, p < .05, with the level of significance of p = .05.
+
 ```{r, include=FALSE}

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
further minor changes

@@
 ```
-##Report produced by: Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo  
-##Date submitted: 16/10/2016
+#####Report produced by: Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo  
+#####Date submitted: 16/10/2016
 
@@
 This area of analysis focuses on investigation of whether playing at 'home' gives the team advantage comparing to when playing 'away'.
-Esssentially, this analysis focuses on effect of the game location (home/away) on winning ratio.
+Essentially, this analysis focuses on effect of the game location (home/away) on winning ratio.
 
@@
 #plot showing home and away win ratio per team
-ggplot(advdf3, aes(x = TEAM)) + geom_point(aes(y = WIN_PCT_HOME, col = "wins at HOME")) + geom_point(aes(y = WIN_PCT_AWAY, col = "wins  AWAY")) + ylab("Win ratio") + xlab("Team Names")
+ggplot(advdf3, aes(x = TEAM)) + geom_point(aes(y = WIN_PCT_HOME, col = "wins HOME")) + geom_point(aes(y = WIN_PCT_AWAY, col = "wins  AWAY")) + ylab("Win ratio") + xlab("Team Names")
 ```
@@
 Improvement of travel conditions.
-This means that when playing away, teams are now more likely to stay at better hotels and fly charterd planes more often than in the past.
+This means that when playing away, teams are now more likely to stay at better hotels and fly chartered planes more often than in the past.
 These changes result in teams being less tired as a consequence of travel and being better physically prepared for the games.
@@
     3. Calculating ratios for effects of rest days on win/loss form home/away
-    4. Perform a Chi-square test to examine whether there is a difference in the distributon for games won and games lost for the different restdays
+    4. Perform a Chi-square test to examine whether there is a difference in the distribution for games won and games lost for the different restdays
     5. Calculating ratios for effects of rest days on win/loss form home/away
@@
 
-**Table 5**: Distribution of number of games per rest days
-
- Rest Days| Games | Distribution
-----------|-------|---
-    0  | 427 | 23.62
-    1  | 990 | 54.76
-    2  | 254 | 14.05
-    3  |  59 |  3.26
-    4  |  13 |  0.72
-    5  |   3 | 0.17
-    7  |   6 | 0.33
-    8  |  20 | 1.11
-    9  |   3 | 0.17
-    10 |   3 | 0.17
-   1st game of season  |  30|  1.66
-
-The distribution of rest days shows that 92% of games are covered by 0,1 or 2 days of rest and therefore further analysis will only focus on these 3 sections.
-
 **Table 6**: Win rate split by number of rest days between games
@@
 The analysis demonstrates that overall the winning ratio stays the same with increase in  rest days.
-However, when looking at invididual restdays the findings are different: back-to-back games have 47.07% winning ratio, 1 day rest increases winning ratio to 50.1% and with 2 days rest the winning ration increases even further to 53.94%.
+However, when looking at individual restdays the findings are different: back-to-back games have 47.07% winning ratio, 1 day rest increases winning ratio to 50.1% and with 2 days rest the winning ration increases even further to 53.94%.
 
@@
 After all, the Null hypothesis must be rejected as the statistical methods used show differences for accuracy in later periods of the game, but as they are negligible in the regular game time but it can be concluded that accuracy does not suffer for later game periods as long as the game ends in regular time.
-For games that reach overtime a highly significant drop in the accuracy of over 5% (45% to below 40%) is observale, stating that players are less likely to score in overtime periods that in the regular game time.
-This means that elite basketball players cope really well with any occurance of fatigue regarding their shot accuracy on a regular basis.
+For games that reach overtime a highly significant drop in the accuracy of over 5% (45% to below 40%) is observable, stating that players are less likely to score in overtime periods that in the regular game time.
+This means that elite basketball players cope really well with any occurrence of fatigue regarding their shot accuracy on a regular basis.
 As soon as the game reaches an intensity which is beyond the regular level effects of fatigue can be found leading to lower shot accuracy.
@@
 Explanations for these differences could be that around the 15 second mark top players feel more obligated to shoot than the average player, even when they are not in the optimal position to do so - leading to lower accuracies.
-Further it seems like Top players dont get affected by last seconds stress as much as average players.
+Further it seems like Top players don't get affected by last seconds stress as much as average players.
 They keep up a quite high accuracy even for the last second of the shot clock (40%).
-However, as there are many similarities for the accuracy of Top players and average players there must be other things than the accuracy that seperates those two groups.
+However, as there are many similarities for the accuracy of Top players and average players there must be other things than the accuracy that separates those two groups.
 
@@
 Further an investigation of the reasons that lead to the higher accuracy would be interesting.
-Do top players keep up their usual shot kinematics when under pressure while average players dont?
+Do top players keep up their usual shot kinematics when under pressure while average players don't?
 Is it a different mindset that helps them to cope better with stress?
@@
 The aim of this analysis is to examine whether the top scorer of a team can help his team to win the match when he has a more outstanding performance in terms of both number of shooting attempts and accuracy.
-The Null hypothesis is that there is no difference regarding the likelyhood of winning if the top scorer of a team has more field goal attempts.
-The second Null hypothesis is there is no difference regarding the likelyhood of winning if the top scorer of a team has a higher shot percentage than on average.
+The Null hypothesis is that there is no difference regarding the likelihood of winning if the top scorer of a team has more field goal attempts.
+The second Null hypothesis is there is no difference regarding the likelihood of winning if the top scorer of a team has a higher shot percentage than on average.
 As this is such a niche area research a literature review did not reveal sufficient results so this analysis aims to fill this gap.
@@
 
-There was no correlation found between the outcome of the matches in this seaon and the field goal percentage of the players.
+There was no correlation found between the outcome of the matches in this season and the field goal percentage of the players.
 The Pearson’s correlation coefficient for the relationship between the variables W and FGA was r(832) < .01, p = .177, failing to reach the level of significance (p = .05).
@@
 graph01 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
-graph01 <- graph01 + labs(title = "Differences in FGA for games lost/won /n of the three special players") + xlab("Player name") + ylab("FGA")
+graph01 <- graph01 + labs(title = "Differences in FGA for games lost/won for particular players") + xlab("Player name") + ylab("FGA")
 graph02 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W)) 
-graph02 <- graph02 + labs(title = "Differences in FG% for games lost/won /n of the three special players") + xlab("Player name") + ylab("FG%")
+graph02 <- graph02 + labs(title = "Differences in FG% for games lost/won for particular players") + xlab("Player name") + ylab("FG%")
 grid.arrange(graph01, graph02, ncol = 2)
@@
 FG% is a significant predictor for the result of match.
-However, the anova showed that the residual deviances of these two factors were both quite large and thus further investigations must be done in order to improve this model.
+However, the anova showed that the residual deviance of these two factors were both quite large and thus further investigations must be done in order to improve this model.
 
@@
 A possible explanation is that it is easier for the opponent to focus on a certain player on defensive side and then the overall efficiency of the team will start to decrease.
-Thus, over-using the scoring ability of a leading player is not likely to be a good approach if the team wants to increase the likelyhood of winning a game.
+Thus, over-using the scoring ability of a leading player is not likely to be a good approach if the team wants to increase the likelihood of winning a game.
 Meanwhile, the accuracy of these players has a stronger impact on the outcome of the game.
@@
   2. Select and analyse subset of all closest defender distance values  
-  3. Count the occurence fOr every distance  
+  3. Count the occurrence for every distance  
   4. Calculate a number of successful shots for every distance point  
   5. Calculate accuracy ratio for every defender distance value
-  6. Seperate data in two ranges - 0 to 8 feet distance and above 8 feet distance
+  6. Separate data in two ranges - 0 to 8 feet distance and above 8 feet distance
 
@@
 ###Plot Result
-  The explanation from the plot above are as follows:   
-  1. Defender distance range from 0 to 2 feet, the accuracy varied between 42% to 50%  
-  2. Defender distance range from 2 to 4 feet, the accuracy rate for this distance range from 44% to 49%  
-  3. Defender distance range from 4 to 6 feet, the accuracy rate for this distance range from 41% to 46%
-  4. Defender distance range from 6 to 8 feet, the accuracy rate for this distance range from 37% to 42%
-  5. Defender distance range from 8 to 10 feet, the accuracy rate for this distance range from 37% to 54%
-  6. Defender distance more than 10 feet up until 16 feet, the accuracy varies greater than previous distance range, starting from 33% to 60%.  
-  7. Distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance)
+
+** Table 13**: Relationship between the defender distance and accuracy
+
+Defender Distance (range in feet) | accuracy (range in %) 
+- | -
+0 - 2 | 42 - 50
+2 - 4 | 44 - 49
+4 - 6 | 41 - 46
+6 - 8 | 37 - 42
+8 - 10 | 37 - 54
+10 - 16 | 33 - 60
+16+ | 0 - 100
 
@@
 For the defender distance starting from 0 to 8 feet, there is a negative correlation between distance and accuracy, implying that players are more likely to miss when the defender comes closer.
-For the distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successfull shot from that distance), which means that this part of the analysis should not be taken into account as the results were insignificant.
+For the distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successful shot from that distance), which means that this part of the analysis should not be taken into account as the results were insignificant.
 It also shows that only a small amount of shots were made for higher defender distances.
@@
   2. Select subsets of all shooting distances from the basket  
-  3. Count the occurence fOr every distance to basket value  
+  3. Count the occurrence for every distance to basket value  
   4. Allocate shot success rate to every distance from basket value 
   5. Examine the relationship between distance to the basket and shot outcome with a correlation test
-  6. Seperate data into ranges
+  6. Separate data into ranges
 
@@
 
-**Table 13**: The relationship between the shooter distance ranges and accuracy
+**Table 14**: The relationship between the shooter distance ranges and accuracy
 
@@
   There was low positive correlation found between the shooting distance and shooting accuracy for the range 10 to 20 feet distance.
-The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(99) = .11, p > .05, fail to reach level of significance (p = .05).
+The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(99) = .11, p =.24, failing to  the level of significance (p = .05).
 ```{r, include=FALSE}
@@
   The analysis aims to investigate whether the effect of the 'hot hand theory' is valid based on the given dataset.
-The 'hothand theory' states that if the shooter makes his first shot, the probability of making his next shot is higher.
-(Gilovich, T., Vallone, R.
-and Tversky, A.
-(1985))
+The 'hot hand theory' states that if the shooter makes his first shot, the probability of making his next shot is higher.
+(Gilovich et al., 1985)
 
@@
   The distribution of all players is left skewed and all of them (especially Klay Thompson) have some long shooting streaks.
-In order to prove/disprove the hot hand theory,  the examination of the indipendence of the shots is required.
+In order to prove/disprove the hot hand theory,  the examination of the independence of the shots is required.
 If each shot that a player takes is independent of the next shot, then the player has the same probability of hitting each shot regardless of the previous shot.
 If each shot that a player takes is dependent of the next shot, then the probability of making the next shot is higher - hot hand.
-Sampling an independent shooter having the same shooting percentage with the real one and comparing his distribution with real player's distribution gives further insight in the specific aera.
+Sampling an independent shooter having the same shooting percentage with the real one and comparing his distribution with real player's distribution gives further insight in the specific area.
 
@@
 There is no such distribution in the sample, so it could be assumed that he is likely to have a hot hand.
-However, further analysis of Klay Thompson's (and other basketball players) shooting streaks in other games is required to see if such a rare occurence is repeated(this is out of scope for this analysis).
+However, further analysis of Klay Thompson's (and other basketball players) shooting streaks in other games is required to see if such a rare occurrence is repeated(this is out of scope for this analysis).
 
@@
 
-  Since the dataset is reffered to only one season and the analysis is not able to eliminate the hot hand theory, it would be an interesting topic to be investigated by using data from more than one season.
+  Since the dataset is referred to only one season and the analysis is not able to eliminate the hot hand theory, it would be an interesting topic to be investigated by using data from more than one season.
 
 ***
-#Conclusion
+##Conclusion
 
@@
 ***
-#References
+##References
 

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
final changes

@@
 ```
-#####Report produced by: Justin Leiendecker, Alexander Romanenko, Artmeis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo  
+#####Report produced by: Justin Leiendecker, Alexander Romanenko, Artemis Tomadaki-Balomenou, Zijun Wei, Reza Brianca Widodo  
 #####Date submitted: 16/10/2016
 
-***
-
-##Introduction
+---
+
+#Introduction
 
@@
 
-** TO GET UPDATED AS A FINAL STEP**
-
 **1. Team related**  
-  a. Home Advantage analysis  
-  b. Effect of rest on teams' performance
+  1.1 Home Advantage analysis  
+  1.2 Effect of rest on teams' performance
 
 **2. Player related**  
-  a. Effects on Accuracy  
-  b. Shot Clock Pressure  
-  c. Fatigue Effect (Quarter Accuracy Rate)    
-  d. Game result impact by field goal attempts  
-  e. Defender Proximity Analysis  
-  f. Shooting Distance vs Shot Accuracy   
-  g. Hot Hand Theory
-
-###Dataset Description:
+  2.1 Effects on Accuracy  
+    2.1.1 Fatigue Effect (Quarter Accuracy Rate) & Shot Clock Pressure     
+    2.1.2 Defender Proximity Analysis
+    2.1.3 Shooting Distance vs Shot Accuracy
+  2.2 Hot Hand Theory  
+  2.3 Game result impact by field goal attempts
+
+##Dataset Description:
 
@@
 
-###Dataset Strengths
+##Dataset Strengths
 
@@
 
-###Dataset Limitations  
+##Dataset Limitations  
   There are several limitations of the dataset:  
@@
 
-## Home Advantage analysis
+---
+
+## 1.1 Home Advantage analysis
 ###Introduction
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, width=50}
 #plot showing home and away win ratio per team
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, width=120}
 #plot showing home and away win ratio per team
-ggplot(advdf3, aes(x = TEAM)) + geom_point(aes(y = WIN_PCT_HOME, col = "wins HOME")) + geom_point(aes(y = WIN_PCT_AWAY, col = "wins  AWAY")) + ylab("Win ratio") + xlab("Team Names")
+ggplot(advdf3, aes(x = TEAM,)) + geom_point(aes(y = WIN_PCT_HOME, col = "wins HOME")) + geom_point(aes(y = WIN_PCT_AWAY, col = "wins  AWAY")) + ylab("Win ratio") + xlab("Team Names") 
 ```
@@
 
-***
-
-##Effect of rest on teams' performance
+---
+
+## 1.2 Effect of rest on teams' performance
 
@@
 
-###Findings
+###Results
 
@@
 
-***
-
-##Effects on accuracy
+---
+
+## 2.1 Effects on accuracy
+
+### 2.1.1 Fatigue Effect (Quarter Accuracy Rate) & Shot Clock Pressure
 
@@
 graph2 <- ggplot(NULL) + geom_smooth(data = df.acc.secondsplayers3, aes(x = seconds.left, y = acc.top, color = "Top Players")) + geom_smooth(data = acc, aes(x = seconds.left, y = accuracy, color = "Average"))
-graph2 + labs(title = "Comparison - Accuracy per second on shot clock left for Top Players vs Overall Average") + xlab("Seconds left on shot clock") + ylab("Accuracy")
+graph2 + labs(title = "Accuracy per second on shot clock left for Top Players vs Overall Average") + xlab("Seconds left on shot clock") + ylab("Accuracy")
 
@@
 
-***
-
-## Game result impact by field goal attempt
+---
+
+###2.1.2 Closest Defender Distance vs Shot Accuracy
+###Introduction  
+This analysis focuses on whether there is a significant effect from closest defender distance to the shot outcome.
+The analysis is aimed to establish whether being away from the defender, i.e.
+having more time to prepare and execute a shot enables the shooter to score with a higher probability.
+Opposite effect will be when defender is close to the shooter it doesn't leave the shooter enough time to prepare and execute the shot properly.
+
+The Null Hypothesis is that the distance to a closest defender does not have an impact on accuracy of shooting.
+
+###Methods  
+The steps to calculate this measurement are as follows:  
+  1. Assign shot result to determine whether the shot is scored (1) or missed (0)  
+  2. Select and analyse subset of all closest defender distance values  
+  3. Count the occurrence for every distance  
+  4. Calculate a number of successful shots for every distance point  
+  5. Calculate accuracy ratio for every defender distance value
+  6. Separate data in two ranges - 0 to 8 feet distance and above 8 feet distance
+
+```{r, include=FALSE}
+#Calculate accuracy against defender distance
+#Group the defender distance level
+distance_level <- as.data.frame(table(factor(df[, "CLOSE_DEF_DIST"])))
+head(distance_level)
+
+#Calculate how many times players scored in every defender distance
+for (i in 1:nrow(distance_level)){
+  distance_level[i, "SCORED"] <- sum(df[df[, "CLOSE_DEF_DIST"] == distance_level[i, "Var1"], "FGM"])
+}
+
+#Calculate shooting accuracy for each distance
+distance_level$Accuracy <- round(distance_level$SCORED / distance_level$Freq, 4)
+
+#change distance level to numeric type for correlation test function
+distance_level$Var1 <- as.numeric(distance_level$Var1)
+distance_level$Var1 <- distance_level$Var1 * 0.1 - 0.1 #to keep the distance point as it is
+
+library(ggplot2)
+library(scales)
+ggplot(data = df) + geom_bar(aes(x = df[, "CLOSE_DEF_DIST"])) +
+  scale_x_discrete("Closest Defender Distance", breaks = c (0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24), limits = 0:24) + 
+  ggtitle("Defender Distance Distribution")
+
+```
+
+```{r, echo=FALSE}
+#Plotting the result
+library(ggplot2)
+library(scales)
+ggplot(data = distance_level) +
+  geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
+  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
+  scale_x_discrete("Closest Defender Distance", limits = 0:24) +  
+  ylab("Shooting Accuracy (%)") + ggtitle("Closest Defender Distance vs Shooting Accuracy")
+```
+
+###Results
+
+####Plot Result
+
+** Table 9**: Relationship between the defender distance and accuracy
+
+Defender Distance (range in feet) | accuracy (range in %) 
+- | -
+0 - 2 | 42 - 50
+2 - 4 | 44 - 49
+4 - 6 | 41 - 46
+6 - 8 | 37 - 42
+8 - 10 | 37 - 54
+10 - 16 | 33 - 60
+16+ | 0 - 100
+
+####Correlation Test
+  The next analysis calculates the correlation between both variables. If the analysis included all data points for every distance, the correlation result would be as follows:
+  There is a medium to strong positive correlation between the defender distance and shooting accuracy.
+The Pearson’s correlation coefficient for the relationship between the variables "Defender Distance" and "Accuracy" was r(297) = .59, p < .01, with a level of significance of p = .05.
+
+```{r, include=FALSE}
+
+#populate data points from data frame
+distanceZerotoAll <- c()
+accuracyDist1 <- c()
+for (i in 0:nrow(distance_level)){
+  distanceZerotoAll <- c(distanceZerotoAll, distance_level[i, "Var1"])
+  accuracyDist1 <- c(accuracyDist1, distance_level[i, "Accuracy"])
+}
+
+#combine data frame
+dfZerotoAll <- data.frame(distanceZerotoAll, accuracyDist1)
+```
+
+```{r, include=FALSE}
+#calculate correlation
+cor.test(dfZerotoAll$distanceZerotoAll, dfZerotoAll$accuracyDist1)
+```
+
+####Correlation Test Result  
+  When analysing the relationship between distance and the accuracy for the range of 0 to 8 feet distance the result of the correlation is different:  
+  There was medium negative correlation between the defender distance and shooting accuracy in this range.
+The Pearson’s correlation coefficient for the relationship between the variables Defender Distance and Accuracy was r(297) = -0.40, p < .01, with the level of significance of p = .05.
+```{r, include=FALSE}
+#populate data from data frame
+distanceZerotoEight <- c()
+accuracyDist2 <- c()
+for (i in 0:80){
+  distanceZerotoEight <- c(distanceZerotoEight, distance_level[i, "Var1"])
+  accuracyDist2 <- c(accuracyDist2, distance_level[i, "Accuracy"])
+}
+
+#combine data frame
+dfZerotoEight <- data.frame(distanceZerotoEight, accuracyDist2)
+```
+
+```{r, include=FALSE}
+#calculate correlation
+cor.test(dfZerotoEight$distanceZerotoEight, dfZerotoEight$accuracyDist2)
+```
+
+###Discussion  
+  The analysis demonstrated that the accuracy rate was affected by the defender distance for certain ranges.
+For the defender distance starting from 0 to 8 feet, there is a negative correlation between distance and accuracy, implying that players are more likely to miss when the defender comes closer.
+For the distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successful shot from that distance), which means that this part of the analysis should not be taken into account as the results were insignificant.
+It also shows that only a small amount of shots were made for higher defender distances.
+Therefore, it would affect the accuracy calculation result.
+However, from the correlation perspective, the result would be different if the analysis included all of the data points versus only short distance data points.
+This was arguably due to the fact that the defender distance does not mean the shooter distance from the basket was closer.
+Therefore, we reject the Null Hypothesis, since the defender distance affected shooting accuracy.
+
+###Future Research
+  As the dataset did not hold many data points for more than 20 feet defender distance, the accuracy analysis for these distances are limited.
+The future analysis could be enhanced further by including data for additional seasons.
+
+---
+
+###2.1.3 Shooting Distance vs Shot Accuracy
+###Introduction
+
+This analysis focuses on whether there is any significant effect resulting from shooting distance from basket on the shot outcome.
+
+The NULL hypothesis is that the distance to the basket does not influence outcome of a shot.
+
+###Methods  
+The steps to calculate this measurement are as follows:  
+  1. Assign shot result to determine whether the shot is scored (1) or missed (0)  
+  2. Select subsets of all shooting distances from the basket  
+  3. Count the occurrence for every distance to basket value  
+  4. Allocate shot success rate to every distance from basket value 
+  5. Examine the relationship between distance to the basket and shot outcome with a correlation test
+  6. Separate data into ranges
+
+```{r, include=FALSE}
+#Calculate accuracy against shooting distance
+#Group the shooting distance level
+shooting_distance_level <- as.data.frame(table(factor(df[, "SHOT_DIST"])))
+head(shooting_distance_level)
+
+#Calculate how many times players scored in every shooting distance
+for (i in 1:nrow(shooting_distance_level)){
+  shooting_distance_level[i, "SCORED"] <- sum(df[df[, "SHOT_DIST"] == shooting_distance_level[i, "Var1"], "FGM"])
+}
+
+#Calculate shooting accuracy for each distance
+shooting_distance_level$Accuracy <- round(shooting_distance_level$SCORED / shooting_distance_level$Freq, 4)
+
+#change distance level to numeric type for correlation test function
+shooting_distance_level$Var1 <- as.numeric(shooting_distance_level$Var1)
+shooting_distance_level$Var1 <- shooting_distance_level$Var1 * 0.1 - 0.1 #to maintain distance data point as it is
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+#Plotting the result
+library(ggplot2)
+library(scales)
+ggplot(data = shooting_distance_level, aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) + 
+  geom_point(aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) + 
+  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
+  scale_x_discrete("Shooting Distance", limits = 0:43) + 
+  geom_smooth() + 
+  ylab("Shooting Accuracy (%)") + 
+  ggtitle("Shooting Distance vs Shooting Accuracy")
+```
+###Results
+
+####Plot Result
+
+**Table 10**: The relationship between the shooter distance ranges and accuracy
+
+  Distance (in feet) | Accuracy (range in %)
+  - | -
+  0 - 2 | 62 - 68
+  2 - 4 | 61 - 67
+  4 - 6 | 44 - 58
+  6 - 8 | 38 - 51
+  8 - 10 | 35 - 42
+  10 - 12 | 34 - 42
+  12 - 14 | 37 - 47
+  14 - 16 | 33 - 51
+  16 - 18 | 37 - 48
+  18 - 24 | 33 - 43
+  24 - 28 | 18 - 38
+  28+ | 0 - 50
+
+  There is a strong negative correlation between the shooting distance and shooting accuracy.
+The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(446) = -0.87, p < .05, with the level of significance of p = .05.
+
+```{r, include=FALSE}
+
+#populate data for every shooting distance
+shootingDistanceZerotoAll <- c()
+shootingAccuracyDist1 <- c()
+for (i in 0:nrow(shooting_distance_level)){
+  shootingDistanceZerotoAll <- c(shootingDistanceZerotoAll, shooting_distance_level[i, "Var1"])
+  shootingAccuracyDist1 <- c(shootingAccuracyDist1, shooting_distance_level[i, "Accuracy"])
+}
+
+dfShootingZerotoAll <- data.frame(shootingDistanceZerotoAll, shootingAccuracyDist1)
+```
+
+```{r, include=FALSE}
+cor.test(dfShootingZerotoAll$shootingDistanceZerotoAll, dfShootingZerotoAll$shootingAccuracyDist1)
+```
+
+####Correlation Test Result  
+  If the correlation test taken for a certain range of distance based on the diagram, the correlation result would be different as follows:
+  There was low positive correlation found between the shooting distance and shooting accuracy for the range 10 to 20 feet distance.
+The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(99) = .11, p =.24, failing to  the level of significance (p = .05).
+```{r, include=FALSE}
+#populate data for shooting distance in range 10 to 20 feet
+shootingDistance10to20 <- c()
+shootingAccuracyDist2 <- c()
+for (i in 100:200){
+  shootingDistance10to20 <- c(shootingDistance10to20, shooting_distance_level[i, "Var1"])
+  shootingAccuracyDist2 <- c(shootingAccuracyDist2, shooting_distance_level[i, "Accuracy"])
+}
+
+dfShooting10to20 <- data.frame(shootingDistance10to20, shootingAccuracyDist2)
+```
+
+```{r, include=FALSE}
+cor.test(dfShooting10to20$shootingDistance10to20, dfShooting10to20$shootingAccuracyDist2)
+```
+
+###Discussion  
+  Based on the diagram and correlation above, it could be concluded that although the shooting distance affects the shooting accuracy, the relationship between these variables is not perfectly linear.
+This was shown in the diagram where the accuracy for shots taken from 10 to 20 feet is relatively similar, regardless the shooter's performance.
+Regarding the low success rate for higher distances, it could be arguably inferred that the shooter tried to attempt a shot in closing game time (buzzer beater) or when the shot clock or game clock nearly reach zero - Therefore, the analysis would reject the Null Hypothesis.
+
+###Future Research
+ The dataset mostly returns 0% for shooting distance greater than 30 feet. Therefore the future research could be enriched with statistics from other seasons to see if these findings hold.
+
+---
+
+## 2.2 Hot Hand Theory
+###Introduction
+
+  The analysis aims to investigate whether the effect of the 'hot hand theory' is valid based on the given dataset.
+The 'hot hand theory' states that if the shooter makes his first shot, the probability of making his next shot is higher.
+(Gilovich et al., 1985)
+
+###Methods
+The steps to approximate this question are as follows:
+  1. Calculate streaks achieved in all games and from all players
+  2. Categorize streaks per player and game, focusing on the 3 players with the top 3 performances in terms of points
+  3. Extract zeros and keep only streaks above 1
+  4. Visualise streaks' length distribution - barplots
+  5. Visualise streaks' length via boxplot
+  6. Calculate the shooting percentage of the 3 players 
+  7. Create a sample independent shooter 
+  8. Compare distributions of sample and real player
+
+```{r, include=FALSE}
+
+#From 'FGM' column: 1 represents a basket made and 0 represents a basket missed
+#Defining the length of a shooting streak to be the number of consecutive baskets made until a miss occurs
+#Creating a function that gives a vector with the number of total streaks(from all players and all games)
+number_of_streaks <- function(x) {
+  x <- c(0, x, 0)
+  which_zero <- which(x == 0)
+  streak <- diff(which_zero) - 1
+  streak
+}
+
+number_of_streaks(df$FGM)
+
+#Summary of total points
+ptssum<- cast(df1, GAME_ID1+ player_id + player_name  ~ PTS); head(ptssum)
+#Creating "total points per player, per game" column
+ptssum$totalpts<- ptssum[,"2"]*2 + ptssum[,"3"]*3
+
+pttsum2<-ptssum[order(ptssum$totalpts,decreasing=TRUE),]
+
+#Setting a new data frame for the 3 players with best performances(higher points)
+best_players_df <- head(pttsum2, 3)
+
+#Creating a function that gives a vector of streaks for a specific player and game
+player_and_game <- function(x, y) {
+   player <- df[df[, 'player_id'] == x & df[,'GAME_ID1'] == y, 'FGM']
+   number_of_streaks(as.vector(player))
+}
+#For the first player and game try:
+#player_and_game('202681', '21400681')
+
+#Creating a list that gives the streaks of each player and game (3 best performances)
+streak_per_player <- list()
+j <- 1
+for(i in best_players_df$player_id) {
+   streak_per_player[[j]] <- player_and_game(i, best_players_df[best_players_df[, 'player_id'] == i, 'GAME_ID1'])
+   j <- j + 1
+}
+streak_per_player
+
+#Removing zeros from streak_per_player in order to show only the streaks achieved
+streak_per_player[[1]] <- streak_per_player[[1]][streak_per_player[[1]] != 0]
+for(i in 1:length(streak_per_player)) {
+  streak_per_player[[i]] <- streak_per_player[[i]][streak_per_player[[i]] != 0]
+}
+streak_per_player
+```
+
+###Results
+####Barplots
+```{r, include=FALSE}
+#Creating a data frame for the strikes of each one of the 3 players (cannot create a data frame including all of them, because each list has different length)
+player1_df <- data.frame('streaks' = streak_per_player[[1]])
+player2_df <- data.frame('streaks' = streak_per_player[[2]])
+player3_df <- data.frame('streaks' = streak_per_player[[3]])
+```
+
+```{r, echo=FALSE}
+#Showing distribution for each player's streak lengths
+gg1 <- ggplot(data = player1_df, aes(x = streaks)) + geom_bar(fill = 'brown') + labs(title = 'Kyrie Irnving')
+gg2 <- ggplot(data = player2_df, aes(x = streaks)) + geom_bar(fill = 'red') + labs(title = 'Mo Williams')
+gg3 <- ggplot(data = player3_df, aes(x = streaks)) + geom_bar(fill = 'blue') + labs(title = 'Klay Thompson')
+grid.arrange(gg1, gg2, gg3, ncol = 3)
+```
+
+####Barplot result
+  The distribution of the three players is unimodal and left skewed. There are shown some extremely long shooting streaks made (i.e. Klay Thompson with 13 shots in a row).
+
+####Boxplots
+
+```{r, include=FALSE}
+#Create a boxplot for each player's streak lengths 
+attach(mtcars)
+par(mfrow=c(1, 3))
+boxplot(player1_df$streaks, main = 'Kyrie Irnving')
+boxplot(player2_df$streaks, main = 'Mo Williams')
+boxplot(player3_df$streaks, main = 'Klay Thompson')
+```
+
+```{r, include=FALSE}
+#Kyrie Irnving
+summary(player1_df$streaks)
+IQR(player1_df$streaks) 
+```
+#### Outcome for Kyrie Irnving:
+
+  1. The typical length of a streak is 1 
+  2. The Interquartile Range is 1 
+  3. Streak length of 6 is unusually high compared to the rest of the distribution
+
+```{r, include=FALSE}
+#Mo Williams
+summary(player2_df$streaks)
+IQR(player2_df$streaks)
+```
+
+#### Outcome for Mo Williams
+
+  1. The typical length of a streak is 1.5
+  2. The Interquartile Range is 1
+  3. Streak length of 4 is unusually high compared to the rest of the distribution
+
+```{r, include=FALSE}
+#Klay Thompson
+summary(player3_df$streaks)
+IQR(player3_df$streaks)
+```
+
+#### Outcome for Klay Thompson
+
+  1. The typical length of a streak is 8 
+  2. The Interquartile Range is 5
+  3. There is a streak of length 13
+
+### Plots Result
+
+  The distribution of all players is left skewed and all of them (especially Klay Thompson) have some long shooting streaks.
+In order to prove/disprove the hot hand theory,  the examination of the independence of the shots is required.
+If each shot that a player takes is independent of the next shot, then the player has the same probability of hitting each shot regardless of the previous shot.
+If each shot that a player takes is dependent of the next shot, then the probability of making the next shot is higher - hot hand.
+Sampling an independent shooter having the same shooting percentage with the real one and comparing his distribution with real player's distribution gives further insight in the specific area.
+
+```{r, include=FALSE}
+#Find the shooting percentage for each of the 3 players
+j <- 1
+shooting_percentage <- list()
+for(i in best_players_df$player_id) {
+  shooting_percentage[j] <- sum(df[df[, 'player_id'] == i, 'FGM']) / nrow(df[df[, 'player_id'] == i, ])
+  j <- j + 1
+}
+shooting_percentage
+```
+
+####Kyrie Irving
+```{r, include=FALSE}
+###PLAYER 1 VS SAMPLE PLAYER###
+#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
+set.seed(974)
+shot_results <- c('made', 'missed')
+simul_shooter <- sample(shot_results, size = sum(best_players_df[1,4], best_players_df[1,5], best_players_df[1,6]), replace = TRUE, prob = c(shooting_percentage[[1]], 1 - shooting_percentage[[1]]))
+simul_shooter
+
+#Setting 1 for baskets made and 0 for baskets missed
+simul_shooter <- as.list(simul_shooter)
+for (i in 1:length(simul_shooter)) {
+   if (simul_shooter[i] == 'made') {
+     simul_shooter[i] <- 1
+   } else if (simul_shooter[i] == 'missed') {
+     simul_shooter[i] <- 0
+   }
+}
+simul_shooter
+
+#Using previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
+streaks_of_simulshooter <- number_of_streaks(simul_shooter)
+
+#Creating new data frames for the graphs 
+simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
+realshooter_df <- data.frame('streaks' = streak_per_player[[1]])
+
+#Comparisons of data between sample shooter and real shooters
+summary(simulshooter_df$streaks)
+summary(realshooter_df$streaks)
+```
+
+```{r, echo=FALSE}
+#Comparing barplots of sample shooter and real shooter1
+#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
+g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar(fill = 'blue') + labs(title = 'sample shooter')
+g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar(fill = 'red') + labs(title = 'Kyrie Irving')
+```
+
+####Plot Result
+  After producing a sample of shots for a player, the result shows distributions of sample and real shooter are quite similar.
+As a consequence, it can be concluded that Kyrie Irving does not have a hot hand.
+
+#### Mo Williams
+```{r, include=FALSE}
+###PLAYER 2 VS SAMPLE PLAYER###
+#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter2 
+set.seed(572)
+shot_results <- c('made', 'missed')
+simul_shooter <- sample(shot_results, size = sum(best_players_df[2,4], best_players_df[2,5], best_players_df[2,6]), replace = TRUE, 
+                        prob = c(shooting_percentage[[2]], 1 - shooting_percentage[[2]]))
+simul_shooter
+
+#Setting 1 for baskets made and 0 for baskets missed
+simul_shooter <- as.list(simul_shooter)
+for(i in 1:length(simul_shooter)) {
+   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
+   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0
+}
+simul_shooter
+
+#Using previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
+streaks_of_simulshooter <- number_of_streaks(simul_shooter)
+
+#Creating new data frames for the graphs 
+simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
+realshooter_df <- data.frame('streaks' = streak_per_player[[2]])
+
+#Comparisons of data between sample shooter and real shooters
+summary(simulshooter_df$streaks)
+summary(realshooter_df$streaks)
+```
+
+```{r, echo=FALSE}
+#Comparing barplots of sample shooter and real shooter1
+#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
+g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar(fill = 'blue') + labs(title = 'sample shooter')
+g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar(fill = 'green') + labs(title = 'Mo Williams')
+grid.arrange(g1, g2, ncol = 2)
+```
+
+####Plot Result
+
+  After producing a sample of shots for a player, the result shows distributions of sample and real shooter are quite similar.
+As a consequence, it can be concluded that Mo Williams does not have a hot hand.
+
+####Klay Thompson
+```{r, include=FALSE}
+###PLAYER 3 VS SAMPLE PLAYER###
+#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter3 
+set.seed(57)
+shot_results <- c('made', 'missed')
+simul_shooter <- sample(shot_results, size = sum(best_players_df[3,4], best_players_df[3,5], best_players_df[3,6]), replace = TRUE, 
+                        prob = c(shooting_percentage[[3]], 1 - shooting_percentage[[3]]))
+simul_shooter
+
+#Setting 1 for baskets made and 0 for baskets missed
+simul_shooter <- as.list(simul_shooter)
+for(i in 1:length(simul_shooter)) {
+   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
+   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0
+}
+simul_shooter
+
+#Using previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
+streaks_of_simulshooter <- number_of_streaks(simul_shooter)
+
+#Creating new data frames for the graphs 
+simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
+realshooter_df <- data.frame('streaks' = streak_per_player[[3]])
+
+#Comparisons of data between sample shooter and real shooters
+summary(simulshooter_df$streaks)
+summary(realshooter_df$streaks)
+```
+
+```{r, echo=FALSE}
+#Comparing barplots of sample shooter and real shooter1
+#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
+g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar(fill = 'blue') + labs(title = 'sample shooter')
+g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar(fill = 'brown') + labs(title = 'Klay Thompson')
+grid.arrange(g1, g2, ncol = 2)
+```
+
+####Plot Result
+
+  Klay Thompson is an exception making 13 shots in a row.
+There is no such distribution in the sample, so it could be assumed that he is likely to have a hot hand.
+However, further analysis of Klay Thompson's (and other basketball players) shooting streaks in other games is required to see if such a rare occurrence is repeated(this is out of scope for this analysis).
+
+###Discussion
+
+  The analysis demonstrates that the streaks generated by the players (K.
+Irving & M.
+Williams) are similar to randomly generated streaks, which use these players shooting probabilities.
+As a result there is no evidence to suggest that there is a hot hand phenomenon present here.
+
+  In the case of K. Thompson, his streak is seem as somewhat anomaly and further research is required to see if any similar results have been achieved by the player.
+
+###Limitations
+
+  Since the dataset is referred to only one season, the size of the vectors of streaks for each player is very small, a fact that can lead to misleading results.
+Moreover, streaks do not constitute continuous variables and as a result they cannot be used in a linear regression.
+
+###Future research
+
+  Since the dataset is referred to only one season and the analysis is not able to eliminate the hot hand theory, it would be an interesting topic to be investigated by using data from more than one season.
+
+---
+
+##2.3 Game result impact by field goal attempt
 
@@
 
-### Impact of Top Scorers
+#### Impact of Top Scorers
 
@@
 ```
-**Table 9**: Result for the logistic regression analysis for FGA and FG% of 15 players
+**Table 11**: Result for the logistic regression analysis for FGA and FG% of 15 players
 
@@
 
-**Table 10**: Result for the anova for FGA and FG% for 15 players
+**Table 12**: Result for the anova for FGA and FG% for 15 players
 
@@
 
-**Table 11**: Result for the logistic regression analysis for FGA and FG% of the 3 special players
+**Table 13**: Result for the logistic regression analysis for FGA and FG% of the 3 special players
 
@@
 
-**Table 12**: Result for the anova for FGA and FG% for the 3 special players
+**Table 14**: Result for the anova for FGA and FG% for the 3 special players
 
@@
 
-##Defender Proximity Analysis  
-###Closest Defender Distance vs Shot Accuracy
-####Introduction  
-This analysis focuses on whether there is a significant effect from closest defender distance to the shot outcome.
-The analysis is aimed to establish whether being away from the defender, i.e.
-having more time to prepare and execute a shot enables the shooter to score with a higher probability.
-Opposite effect will be when defender is close to the shooter it doesn't leave the shooter enough time to prepare and execute the shot properly.
-
-The Null Hypothesis is that the distance to a closest defender does not have an impact on accuracy of shooting.
-
-###Methods  
-The steps to calculate this measurement are as follows:  
-  1. Assign shot result to determine whether the shot is scored (1) or missed (0)  
-  2. Select and analyse subset of all closest defender distance values  
-  3. Count the occurrence for every distance  
-  4. Calculate a number of successful shots for every distance point  
-  5. Calculate accuracy ratio for every defender distance value
-  6. Separate data in two ranges - 0 to 8 feet distance and above 8 feet distance
-
-```{r, include=FALSE}
-#Calculate accuracy against defender distance
-#Group the defender distance level
-distance_level <- as.data.frame(table(factor(df[, "CLOSE_DEF_DIST"])))
-head(distance_level)
-
-#Calculate how many times players scored in every defender distance
-for (i in 1:nrow(distance_level)){
-  distance_level[i, "SCORED"] <- sum(df[df[, "CLOSE_DEF_DIST"] == distance_level[i, "Var1"], "FGM"])
-}
-
-#Calculate shooting accuracy for each distance
-distance_level$Accuracy <- round(distance_level$SCORED / distance_level$Freq, 4)
-
-#change distance level to numeric type for correlation test function
-distance_level$Var1 <- as.numeric(distance_level$Var1)
-distance_level$Var1 <- distance_level$Var1 * 0.1 - 0.1 #to keep the distance point as it is
-
-library(ggplot2)
-library(scales)
-ggplot(data = df) + geom_bar(aes(x = df[, "CLOSE_DEF_DIST"])) +
-  scale_x_discrete("Closest Defender Distance", breaks = c (0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24), limits = 0:24) + 
-  ggtitle("Defender Distance Distribution")
-
-```
-
-```{r, echo=FALSE}
-#Plotting the result
-library(ggplot2)
-library(scales)
-ggplot(data = distance_level) +
-  geom_point(aes(x = distance_level$Var1, y = distance_level$Accuracy)) +
-  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
-  scale_x_discrete("Closest Defender Distance", limits = 0:24) +  
-  ylab("Shooting Accuracy (%)") + ggtitle("Closest Defender Distance vs Shooting Accuracy")
-```
-
-###Plot Result
-
-** Table 13**: Relationship between the defender distance and accuracy
-
-Defender Distance (range in feet) | accuracy (range in %) 
-- | -
-0 - 2 | 42 - 50
-2 - 4 | 44 - 49
-4 - 6 | 41 - 46
-6 - 8 | 37 - 42
-8 - 10 | 37 - 54
-10 - 16 | 33 - 60
-16+ | 0 - 100
-
-###Correlation Test
-  The next analysis calculates the correlation between both variables. If the analysis included all data points for every distance, the correlation result would be as follows:
-  There is a medium to strong positive correlation between the defender distance and shooting accuracy.
-The Pearson’s correlation coefficient for the relationship between the variables "Defender Distance" and "Accuracy" was r(297) = .59, p < .01, with a level of significance of p = .05.
-
-```{r, include=FALSE}
-
-#populate data points from data frame
-distanceZerotoAll <- c()
-accuracyDist1 <- c()
-for (i in 0:nrow(distance_level)){
-  distanceZerotoAll <- c(distanceZerotoAll, distance_level[i, "Var1"])
-  accuracyDist1 <- c(accuracyDist1, distance_level[i, "Accuracy"])
-}
-
-#combine data frame
-dfZerotoAll <- data.frame(distanceZerotoAll, accuracyDist1)
-```
-
-```{r, include=FALSE}
-#calculate correlation
-cor.test(dfZerotoAll$distanceZerotoAll, dfZerotoAll$accuracyDist1)
-```
-
-###Correlation Test Result  
-  When analysing the relationship between distance and the accuracy for the range of 0 to 8 feet distance the result of the correlation is different:  
-  There was medium negative correlation between the defender distance and shooting accuracy in this range.
-The Pearson’s correlation coefficient for the relationship between the variables Defender Distance and Accuracy was r(297) = -0.40, p < .01, with the level of significance of p = .05.
-```{r, include=FALSE}
-#populate data from data frame
-distanceZerotoEight <- c()
-accuracyDist2 <- c()
-for (i in 0:80){
-  distanceZerotoEight <- c(distanceZerotoEight, distance_level[i, "Var1"])
-  accuracyDist2 <- c(accuracyDist2, distance_level[i, "Accuracy"])
-}
-
-#combine data frame
-dfZerotoEight <- data.frame(distanceZerotoEight, accuracyDist2)
-```
-
-```{r, include=FALSE}
-#calculate correlation
-cor.test(dfZerotoEight$distanceZerotoEight, dfZerotoEight$accuracyDist2)
-```
-
-###Discussion  
-  The analysis demonstrated that the accuracy rate was affected by the defender distance for certain ranges.
-For the defender distance starting from 0 to 8 feet, there is a negative correlation between distance and accuracy, implying that players are more likely to miss when the defender comes closer.
-For the distance greater than 16 feet, the data varies between 0% (no shots scored from that distance) to 100% (certain successful shot from that distance), which means that this part of the analysis should not be taken into account as the results were insignificant.
-It also shows that only a small amount of shots were made for higher defender distances.
-Therefore, it would affect the accuracy calculation result.
-However, from the correlation perspective, the result would be different if the analysis included all of the data points versus only short distance data points.
-This was arguably due to the fact that the defender distance does not mean the shooter distance from the basket was closer.
-Therefore, we reject the Null Hypothesis, since the defender distance affected shooting accuracy.
-
-###Future Research
-  As the dataset did not hold many data points for more than 20 feet defender distance, the accuracy analysis for these distances are limited.
-The future analysis could be enhanced further by including data for additional seasons.
-
-***
-
-###Shooting Distance vs Shot Accuracy
-####Introduction
-
-This analysis focuses on whether there is any significant effect resulting from shooting distance from basket on the shot outcome.
-
-The NULL hypothesis is that the distance to the basket does not influence outcome of a shot.
-
-###Methods  
-The steps to calculate this measurement are as follows:  
-  1. Assign shot result to determine whether the shot is scored (1) or missed (0)  
-  2. Select subsets of all shooting distances from the basket  
-  3. Count the occurrence for every distance to basket value  
-  4. Allocate shot success rate to every distance from basket value 
-  5. Examine the relationship between distance to the basket and shot outcome with a correlation test
-  6. Separate data into ranges
-
-```{r, include=FALSE}
-#Calculate accuracy against shooting distance
-#Group the shooting distance level
-shooting_distance_level <- as.data.frame(table(factor(df[, "SHOT_DIST"])))
-head(shooting_distance_level)
-
-#Calculate how many times players scored in every shooting distance
-for (i in 1:nrow(shooting_distance_level)){
-  shooting_distance_level[i, "SCORED"] <- sum(df[df[, "SHOT_DIST"] == shooting_distance_level[i, "Var1"], "FGM"])
-}
-
-#Calculate shooting accuracy for each distance
-shooting_distance_level$Accuracy <- round(shooting_distance_level$SCORED / shooting_distance_level$Freq, 4)
-
-#change distance level to numeric type for correlation test function
-shooting_distance_level$Var1 <- as.numeric(shooting_distance_level$Var1)
-shooting_distance_level$Var1 <- shooting_distance_level$Var1 * 0.1 - 0.1 #to maintain distance data point as it is
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-#Plotting the result
-library(ggplot2)
-library(scales)
-ggplot(data = shooting_distance_level, aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) + 
-  geom_point(aes(x = shooting_distance_level$Var1, y = shooting_distance_level$Accuracy)) + 
-  scale_y_continuous(labels = percent, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0), limits = c(0, 1)) +
-  scale_x_discrete("Shooting Distance", limits = 0:43) + 
-  geom_smooth() + 
-  ylab("Shooting Accuracy (%)") + 
-  ggtitle("Shooting Distance vs Shooting Accuracy")
-```
-
-###Plot Result
-
-**Table 14**: The relationship between the shooter distance ranges and accuracy
-
-  Distance (in feet) | Accuracy (range in %)
-  - | -
-  0 - 2 | 62 - 68
-  2 - 4 | 61 - 67
-  4 - 6 | 44 - 58
-  6 - 8 | 38 - 51
-  8 - 10 | 35 - 42
-  10 - 12 | 34 - 42
-  12 - 14 | 37 - 47
-  14 - 16 | 33 - 51
-  16 - 18 | 37 - 48
-  18 - 24 | 33 - 43
-  24 - 28 | 18 - 38
-  28+ | 0 - 50
-
-  There is a strong negative correlation between the shooting distance and shooting accuracy.
-The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(446) = -0.87, p < .05, with the level of significance of p = .05.
-
-```{r, include=FALSE}
-
-#populate data for every shooting distance
-shootingDistanceZerotoAll <- c()
-shootingAccuracyDist1 <- c()
-for (i in 0:nrow(shooting_distance_level)){
-  shootingDistanceZerotoAll <- c(shootingDistanceZerotoAll, shooting_distance_level[i, "Var1"])
-  shootingAccuracyDist1 <- c(shootingAccuracyDist1, shooting_distance_level[i, "Accuracy"])
-}
-
-dfShootingZerotoAll <- data.frame(shootingDistanceZerotoAll, shootingAccuracyDist1)
-```
-
-```{r, include=FALSE}
-cor.test(dfShootingZerotoAll$shootingDistanceZerotoAll, dfShootingZerotoAll$shootingAccuracyDist1)
-```
-
-###Correlation Test Result  
-  If the correlation test taken for a certain range of distance based on the diagram, the correlation result would be different as follows:
-  There was low positive correlation found between the shooting distance and shooting accuracy for the range 10 to 20 feet distance.
-The Pearson’s correlation coefficient for the relationship between the variables Shooting Distance and Shot Accuracy was r(99) = .11, p =.24, failing to  the level of significance (p = .05).
-```{r, include=FALSE}
-#populate data for shooting distance in range 10 to 20 feet
-shootingDistance10to20 <- c()
-shootingAccuracyDist2 <- c()
-for (i in 100:200){
-  shootingDistance10to20 <- c(shootingDistance10to20, shooting_distance_level[i, "Var1"])
-  shootingAccuracyDist2 <- c(shootingAccuracyDist2, shooting_distance_level[i, "Accuracy"])
-}
-
-dfShooting10to20 <- data.frame(shootingDistance10to20, shootingAccuracyDist2)
-```
-
-```{r, echo=FALSE}
-cor.test(dfShooting10to20$shootingDistance10to20, dfShooting10to20$shootingAccuracyDist2)
-```
-
-###Discussion  
-  Based on the diagram and correlation above, it could be concluded that although the shooting distance affects the shooting accuracy, the relationship between these variables is not perfectly linear.
-This was shown in the diagram where the accuracy for shots taken from 10 to 20 feet is relatively similar, regardless the shooter's performance.
-Regarding the low success rate for higher distances, it could be arguably inferred that the shooter tried to attempt a shot in closing game time (buzzer beater) or when the shot clock or game clock nearly reach zero - Therefore, the analysis would reject the Null Hypothesis.
-
-###Future Research
- The dataset mostly returns 0% for shooting distance greater than 30 feet. Therefore the future research could be enriched with statistics from other seasons to see if these findings hold.
-
-***
-
-##Hot Hand Theory
-###Introduction
-
-  The analysis aims to investigate whether the effect of the 'hot hand theory' is valid based on the given dataset.
-The 'hot hand theory' states that if the shooter makes his first shot, the probability of making his next shot is higher.
-(Gilovich et al., 1985)
-
-###Methods
-The steps to approximate this question are as follows:
-  1. Calculate streaks achieved in all games and from all players
-  2. Categorize streaks per player and game, focusing on the 3 players with the top 3 performances in terms of points
-  3. Extract zeros and keep only streaks above 1
-  4. Visualise streaks' length distribution - barplots
-  5. Visualise streaks' length via boxplot
-  6. Calculate the shooting percentage of the 3 players 
-  7. Create a sample independent shooter 
-  8. Compare distributions of sample and real player
-
-```{r, include=FALSE}
-
-#From 'FGM' column: 1 represents a basket made and 0 represents a basket missed
-#Defining the length of a shooting streak to be the number of consecutive baskets made until a miss occurs
-#Creating a function that gives a vector with the number of total streaks(from all players and all games)
-number_of_streaks <- function(x) {
-  x <- c(0, x, 0)
-  which_zero <- which(x == 0)
-  streak <- diff(which_zero) - 1
-  streak
-}
-
-number_of_streaks(df$FGM)
-
-#Summary of total points
-ptssum<- cast(df1, GAME_ID1+ player_id + player_name  ~ PTS); head(ptssum)
-#Creating "total points per player, per game" column
-ptssum$totalpts<- ptssum[,"2"]*2 + ptssum[,"3"]*3
-
-pttsum2<-ptssum[order(ptssum$totalpts,decreasing=TRUE),]
-
-#Setting a new data frame for the 3 players with best performances(higher points)
-best_players_df <- head(pttsum2, 3)
-
-#Creating a function that gives a vector of streaks for a specific player and game
-player_and_game <- function(x, y) {
-   player <- df[df[, 'player_id'] == x & df[,'GAME_ID1'] == y, 'FGM']
-   number_of_streaks(as.vector(player))
-}
-#For the first player and game try:
-#player_and_game('202681', '21400681')
-
-#Creating a list that gives the streaks of each player and game (3 best performances)
-streak_per_player <- list()
-j <- 1
-for(i in best_players_df$player_id) {
-   streak_per_player[[j]] <- player_and_game(i, best_players_df[best_players_df[, 'player_id'] == i, 'GAME_ID1'])
-   j <- j + 1
-}
-streak_per_player
-
-#Removing zeros from streak_per_player in order to show only the streaks achieved
-streak_per_player[[1]] <- streak_per_player[[1]][streak_per_player[[1]] != 0]
-for(i in 1:length(streak_per_player)) {
-  streak_per_player[[i]] <- streak_per_player[[i]][streak_per_player[[i]] != 0]
-}
-streak_per_player
-```
-
-###Barplots
-```{r, include=FALSE}
-#Creating a data frame for the strikes of each one of the 3 players (cannot create a data frame including all of them, because each list has different length)
-player1_df <- data.frame('streaks' = streak_per_player[[1]])
-player2_df <- data.frame('streaks' = streak_per_player[[2]])
-player3_df <- data.frame('streaks' = streak_per_player[[3]])
-```
-
-```{r, echo=FALSE}
-#Showing distribution for each player's streak lengths
-gg1 <- ggplot(data = player1_df, aes(x = streaks)) + geom_bar(fill = 'brown') + labs(title = 'Kyrie Irnving')
-gg2 <- ggplot(data = player2_df, aes(x = streaks)) + geom_bar(fill = 'red') + labs(title = 'Mo Williams')
-gg3 <- ggplot(data = player3_df, aes(x = streaks)) + geom_bar(fill = 'blue') + labs(title = 'Klay Thompson')
-grid.arrange(gg1, gg2, gg3, ncol = 3)
-```
-
-####Barplot result
-  The distribution of the three players is unimodal and left skewed. There are shown some extremely long shooting streaks made (i.e. Klay Thompson with 13 shots in a row).
-
-###Boxplots
-
-```{r, include=FALSE}
-#Create a boxplot for each player's streak lengths 
-attach(mtcars)
-par(mfrow=c(1, 3))
-boxplot(player1_df$streaks, main = 'Kyrie Irnving')
-boxplot(player2_df$streaks, main = 'Mo Williams')
-boxplot(player3_df$streaks, main = 'Klay Thompson')
-```
-
-```{r, include=FALSE}
-#Kyrie Irnving
-summary(player1_df$streaks)
-IQR(player1_df$streaks) 
-```
-#### Outcome for Kyrie Irnving:
-
-  1. The typical length of a streak is 1 
-  2. The Interquartile Range is 1 
-  3. Streak length of 6 is unusually high compared to the rest of the distribution
-
-```{r, include=FALSE}
-#Mo Williams
-summary(player2_df$streaks)
-IQR(player2_df$streaks)
-```
-
-#### Outcome for Mo Williams
-
-  1. The typical length of a streak is 1.5
-  2. The Interquartile Range is 1
-  3. Streak length of 4 is unusually high compared to the rest of the distribution
-
-```{r, include=FALSE}
-#Klay Thompson
-summary(player3_df$streaks)
-IQR(player3_df$streaks)
-```
-
-#### Outcome for Klay Thompson
-
-  1. The typical length of a streak is 8 
-  2. The Interquartile Range is 5
-  3. There is a streak of length 13
-
-### Plots Result
-
-  The distribution of all players is left skewed and all of them (especially Klay Thompson) have some long shooting streaks.
-In order to prove/disprove the hot hand theory,  the examination of the independence of the shots is required.
-If each shot that a player takes is independent of the next shot, then the player has the same probability of hitting each shot regardless of the previous shot.
-If each shot that a player takes is dependent of the next shot, then the probability of making the next shot is higher - hot hand.
-Sampling an independent shooter having the same shooting percentage with the real one and comparing his distribution with real player's distribution gives further insight in the specific area.
-
-```{r, include=FALSE}
-#Find the shooting percentage for each of the 3 players
-j <- 1
-shooting_percentage <- list()
-for(i in best_players_df$player_id) {
-  shooting_percentage[j] <- sum(df[df[, 'player_id'] == i, 'FGM']) / nrow(df[df[, 'player_id'] == i, ])
-  j <- j + 1
-}
-shooting_percentage
-```
-
-###Kyrie Irving
-```{r, include=FALSE}
-###PLAYER 1 VS SAMPLE PLAYER###
-#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter1 
-set.seed(974)
-shot_results <- c('made', 'missed')
-simul_shooter <- sample(shot_results, size = sum(best_players_df[1,4], best_players_df[1,5], best_players_df[1,6]), replace = TRUE, prob = c(shooting_percentage[[1]], 1 - shooting_percentage[[1]]))
-simul_shooter
-
-#Setting 1 for baskets made and 0 for baskets missed
-simul_shooter <- as.list(simul_shooter)
-for (i in 1:length(simul_shooter)) {
-   if (simul_shooter[i] == 'made') {
-     simul_shooter[i] <- 1
-   } else if (simul_shooter[i] == 'missed') {
-     simul_shooter[i] <- 0
-   }
-}
-simul_shooter
-
-#Using previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
-streaks_of_simulshooter <- number_of_streaks(simul_shooter)
-
-#Creating new data frames for the graphs 
-simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
-realshooter_df <- data.frame('streaks' = streak_per_player[[1]])
-
-#Comparisons of data between sample shooter and real shooters
-summary(simulshooter_df$streaks)
-summary(realshooter_df$streaks)
-```
-
-```{r, echo=FALSE}
-#Comparing barplots of sample shooter and real shooter1
-#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
-g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar(fill = 'blue') + labs(title = 'sample shooter')
-g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar(fill = 'red') + labs(title = 'Kyrie Irving')
-grid.arrange(g1, g2, ncol = 2)
-```
-
-####Plot Result
-  After producing a sample of shots for a player, the result shows distributions of sample and real shooter are quite similar.
-As a consequence, it can be concluded that Kyrie Irving does not have a hot hand.
-
-### Mo Williams
-```{r, include=FALSE}
-###PLAYER 2 VS SAMPLE PLAYER###
-#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter2 
-set.seed(572)
-shot_results <- c('made', 'missed')
-simul_shooter <- sample(shot_results, size = sum(best_players_df[2,4], best_players_df[2,5], best_players_df[2,6]), replace = TRUE, 
-                        prob = c(shooting_percentage[[2]], 1 - shooting_percentage[[2]]))
-simul_shooter
-
-#Setting 1 for baskets made and 0 for baskets missed
-simul_shooter <- as.list(simul_shooter)
-for(i in 1:length(simul_shooter)) {
-   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
-   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0
-}
-simul_shooter
-
-#Using previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
-streaks_of_simulshooter <- number_of_streaks(simul_shooter)
-
-#Creating new data frames for the graphs 
-simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
-realshooter_df <- data.frame('streaks' = streak_per_player[[2]])
-
-#Comparisons of data between sample shooter and real shooters
-summary(simulshooter_df$streaks)
-summary(realshooter_df$streaks)
-```
-
-```{r, echo=FALSE}
-#Comparing barplots of sample shooter and real shooter1
-#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
-g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar(fill = 'blue') + labs(title = 'sample shooter')
-g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar(fill = 'green') + labs(title = 'Mo Williams')
-grid.arrange(g1, g2, ncol = 2)
-```
-
-####Plot Result
-
-  After producing a sample of shots for a player, the result shows distributions of sample and real shooter are quite similar.
-As a consequence, it can be concluded that Mo Williams does not have a hot hand.
-
-###Klay Thompson
-```{r, include=FALSE}
-###PLAYER 3 VS SAMPLE PLAYER###
-#Simulation - creating a vector of shots('made', 'missed') for a sample shooter, setting as probability the shooting percentage of the real shooter3 
-set.seed(57)
-shot_results <- c('made', 'missed')
-simul_shooter <- sample(shot_results, size = sum(best_players_df[3,4], best_players_df[3,5], best_players_df[3,6]), replace = TRUE, 
-                        prob = c(shooting_percentage[[3]], 1 - shooting_percentage[[3]]))
-simul_shooter
-
-#Setting 1 for baskets made and 0 for baskets missed
-simul_shooter <- as.list(simul_shooter)
-for(i in 1:length(simul_shooter)) {
-   if(simul_shooter[i] == 'made') simul_shooter[i] <- 1
-   else if(simul_shooter[i] == 'missed') simul_shooter[i] <- 0
-}
-simul_shooter
-
-#Using previous function (number_of_streaks) to calculate the streaks made by the sample shooter 
-streaks_of_simulshooter <- number_of_streaks(simul_shooter)
-
-#Creating new data frames for the graphs 
-simulshooter_df <- data.frame('streaks' = streaks_of_simulshooter)
-realshooter_df <- data.frame('streaks' = streak_per_player[[3]])
-
-#Comparisons of data between sample shooter and real shooters
-summary(simulshooter_df$streaks)
-summary(realshooter_df$streaks)
-```
-
-```{r, echo=FALSE}
-#Comparing barplots of sample shooter and real shooter1
-#If real shooter's streaks diverge significantly from an independent shooter's streaks(sample shooter's streaks), we can conclude tha the real shooter has a 'hot hand'. 
-g1 <- ggplot(data = simulshooter_df, aes(x = streaks )) + geom_bar(fill = 'blue') + labs(title = 'sample shooter')
-g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar(fill = 'brown') + labs(title = 'Klay Thompson')
-grid.arrange(g1, g2, ncol = 2)
-```
-
-####Plot Result
-
-  Klay Thompson is an exception making 13 shots in a row.
-There is no such distribution in the sample, so it could be assumed that he is likely to have a hot hand.
-However, further analysis of Klay Thompson's (and other basketball players) shooting streaks in other games is required to see if such a rare occurrence is repeated(this is out of scope for this analysis).
-
-###Discussion
-
-  The analysis demonstrates that the streaks generated by the players (K.
-Irving & M.
-Williams) are similar to randomly generated streaks, which use these players shooting probabilities.
-As a result there is no evidence to suggest that there is a hot hand phenomenon present here.
-
-  In the case of K. Thompson, his streak is seem as somewhat anomaly and further research is required to see if any similar results have been achieved by the player.
-
-###Limitations
-
-  Since the dataset is referred to only one season, the size of the vectors of streaks for each player is very small, a fact that can lead to misleading results.
-Moreover, streaks do not constitute continuous variables and as a result they cannot be used in a linear regression.
-
-###Future research
-
-  Since the dataset is referred to only one season and the analysis is not able to eliminate the hot hand theory, it would be an interesting topic to be investigated by using data from more than one season.
-
-***
+---
+
 ##Conclusion
@@
 
-***
+---
+
 ##References
@@
 Journal of Strength and Conditioning Research, 23(3), pp.1029-1036.
-  Gilovich, T., Vallone, R. and Tversky, A. (1985) ‘The hot hand in basketball: On the misperception of random sequences’, Cognitive Psychology, 17(3), pp. 295–314. doi: 10.1016/0010-0285(85)90010-6.
+  Gilovich, T., Vallone, R.
+and Tversky, A.
+(1985) ‘The hot hand in basketball: On the misperception of random sequences’, Cognitive Psychology, 17(3), pp.
+295–314.
+doi: 10.1016/0010-0285(85)90010-6.
   Mascret, N., Ibáñez-Gijón, J., Bréjard, V., Buekers, M., Casanova, R., Marqueste, T., Montagne, G., Rao, G., Roux, Y.

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
almost there!
last graph changes

@@
 
-** Table 9**: Relationship between the defender distance and accuracy
+**Table 9**: Relationship between the defender distance and accuracy
 
@@
 ```{r, echo=FALSE}
-graph01 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))
-graph01 <- graph01 + labs(title = "Differences in FGA for games lost/won for particular players") + xlab("Player name") + ylab("FGA")
-graph02 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W)) 
-graph02 <- graph02 + labs(title = "Differences in FG% for games lost/won for particular players") + xlab("Player name") + ylab("FG%")
-grid.arrange(graph01, graph02, ncol = 2)
+graph01 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))+ graph01 + labs(title = "Differences in FGA for games lost/won for particular players") + xlab("Player name") + ylab("FGA")
+graph02 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W)) + labs(title = "Differences in FG% for games lost/won for particular players") + xlab("Player name") + ylab("FG%")
+graph01
+graph02
 ```

### 2016-10-16 MSFGroupwork-NBAHotHand.Rmd
final changes

@@
 g2 <- ggplot(data = realshooter_df, aes(x = streaks)) + geom_bar(fill = 'red') + labs(title = 'Kyrie Irving')
+grid.arrange(g1, g2, ncol = 2)
 ```
@@
 ```{r, echo=FALSE}
-graph01 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FGA, colour = W))+ graph01 + labs(title = "Differences in FGA for games lost/won for particular players") + xlab("Player name") + ylab("FGA")
+graph01 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FGA, colour = W)) + labs(title = "Differences in FGA for games lost/won for particular players") + xlab("Player name") + ylab("FGA")
 graph02 <- ggplot(special2) + geom_boxplot(aes(x = player_name, y = FG.percentage, colour = W)) + labs(title = "Differences in FG% for games lost/won for particular players") + xlab("Player name") + ylab("FG%")
@@
   Ahart (1973). In Kendall, P. and Hollon, S. (1979). Cognitive-behavioral interventions. New York: Academic Press.
+
   Economist: As sweet as ever (2015) Available at: http://www.economist.com/blogs/gametheory/2015/06/home-advantage-basketball (Accessed: 16 October 2016).
+
   Erculj, F.
@@
 Journal of Strength and Conditioning Research, 23(3), pp.1029-1036.
-  Gilovich, T., Vallone, R.
-and Tversky, A.
-(1985) ‘The hot hand in basketball: On the misperception of random sequences’, Cognitive Psychology, 17(3), pp.
-295–314.
-doi: 10.1016/0010-0285(85)90010-6.
+
+  Gilovich, T., Vallone, R. and Tversky, A. (1985) ‘The hot hand in basketball: On the misperception of random sequences’, Cognitive Psychology, 17(3), pp. 295–314. doi: 10.1016/0010-0285(85)90010-6.
+
   Mascret, N., Ibáñez-Gijón, J., Bréjard, V., Buekers, M., Casanova, R., Marqueste, T., Montagne, G., Rao, G., Roux, Y.
